### 2017-12-03 Report/paper.tex

\documentclass{article}

\usepackage{times}
\usepackage{graphicx} 
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{hyperref}
\usepackage{icml2016} 
\usepackage{amsmath}
\usepackage{subcaption}

\begin{document} 
\twocolumn[\icmltitle{Handwriting Recognition}
\icmlauthor{Jarre Knockaert}{Jarre.Knockaert@UGent.be}
\icmlauthor{Ruben Dedecker}{Ruben.Dedecker@UGent.be}

\vskip 0.3in
]

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Handwriting Recognition}

\subsection{Character Recognition}
The smallest distinguishable token in a text is a character. These handwritten characters are uniquely 
written by each individual and there's even variation when written by the same individual. These variations include the inclination of characters, i.e., the slant, orientation and size. These variations can be removed by preprocessing the image. We however take another, more general approach. Instead of adding rules to minimize variation we made a deep neural net which can recognise any character with a certain probability. The architecture of this deep neural network is described in \ref{sec:dnn}. In order to make the deep neural network recognise characters in images with these variations, we augmented the dataset to include images which have these variations. We discuss this in \ref{sec:preproc}. The dataset of handwritten character is described in \ref{sec:data}.

\subsubsection{Data}
\label{sec:data}
A good data dataset with many examples, enough variation, and which is balanced between classes is very important to train the neural network. We use the Chars74K dataset for this purpose. It has 62 classes (0-9, A-Z, a-z) and 3410 handwritten images. This dataset contains both very easily recognisable images of characters but also characters which are even hard to recognise for humans, as shown in figure \ref{fig:char}. 3410 input images is quite low for a neural network to be effectively trained. To overcome this problem we use some data augmentation techniques discussed in \ref{par:aug}.

\begin{figure}
\begin{subfigure}{.15\textwidth}
  \centering
  \includegraphics[width=\linewidth]{images/bad_char1}
  \caption{Picture of a 'z' which resembles a 7.}
\end{subfigure}
\begin{subfigure}{.15\textwidth}
  \centering
  \includegraphics[width=\linewidth]{images/bad_char2}
  \caption{Picture of a 'z' which resembles a 3.}
\end{subfigure}
\begin{subfigure}{0.15\textwidth}
  \centering
  \includegraphics[width=\linewidth]{images/good_char}
  \caption{Picture of a 'z' which resembles a 3.}
\end{subfigure}
\caption{Examples of written 'z' characters.}
\label{fig:char}
\end{figure}

\subsubsection{Preliminary steps}
\label{sec:preproc}
We make use of 2 types of preprocessing steps: normalization of the data (\ref{par:norm}) and augmentation of the data (\ref{par:aug}). Normalization of data adjusts the original data and data augmentation takes the data and uses augmentation techniques to create variations of the original data.

\paragraph{Normalizing data} 
\label{par:norm}
Theoretically feeding normalized data to the neural network returns the same output as before. 
In reality however, it's better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}. We normalize the data by reducing the size from 1200x900x3 to 32x32x1 (width x height x number of color channels).

%TODO BINARIZATION -> INCREASE CONTRAST BETWEEN BACKGROUND AND FOREGROUND
%This might solve our problem that handwriting isn't recognised.

%TODO INVERTING -> We still need to check influence of inverting color values. 
%This surely helps for augmenting the data.

% Size / Contrast
\paragraph{Augmenting data}
\label{par:aug}
Data augmentation serves 2 purposes. It makes the dataset more robust to different kinds of handwriting as we add variations of the original data to the input data. The second purpose is to increase the size of our dataset. As our dataset only contains images of about 3400 images with handwritten characters, we need more data for the neural network to function to its fullest extend. Increasing the size of our dataset has great impact on the performance of the neural network as discussed in \ref{sec:expres}. Now we will describe the data augmentation techniques we used. For these techniques we based us on \cite{DataAug}. A more thorough explanation can be found there.

\subparagraph{Adding noise to data}
We add a variant of each image which includes noise troughout the image with the purpose of reducing overfitting of the neural network \cite{DataNoise}. This noise is taken from the Gaussian distribution and summed with the original image.  
\subparagraph{Translations of data}
The dataset is also extended with variations in padding of characters in images. For this purpose we multiply every coordinate of the original image with the follow transformation matrix: 
\begin{equation}
        \begin{bmatrix}
                1 & 0 & t_x \\
                0 & 1 & t_y
        \end{bmatrix}
\end{equation}
where $t_x$ is the horizontal shift and $t_y$ is the vertical shift. 
\subparagraph{Rotations of data}
Rotations of the image are added to the dataset to make the network more robust to handwritings with different orientations. We rotate by -30 and 30 degrees. We use a builtin python function from the opencv2 library to create the transformation matrix. This matrix is equal to:
\begin{equation}
       \begin{bmatrix}
               \alpha & \beta & (1-\alpha)*center.x - \beta*center.y \\
               -\beta & \alpha & \beta*center.x + (1-\alpha)*center.y
       \end{bmatrix}
\end{equation}
where $\alpha = cos(angle)$ and $\beta = sin(angle)$. Center is the coordinate in our image around which is rotated and angle describes the amount of degrees to rotate in the clockwise direction. Translated images outside of the 32x32 frame are discarded.
\subparagraph{Scaling of data}
With the purpose of making the network more robust to different sizes of handwriting, we add different scalings of the data. We both scale in with and/or in height. The following transformation matrix scales the input image: 
\begin{equation}
       \begin{bmatrix}
               s_x & 0 & 0  \\
               0 & s_y & 0
       \end{bmatrix}
\end{equation}
where $s_x$ is the horizontal scaling factor and $s_y$ is the vertical scaling factor. 
\subsubsection{Deep neural network}
\label{sec:dnn}

\subsection{Tokenization}

\subsection{Natural language processing}

\subsubsection{Vocabularium}
\subsubsection{Language model}

\section{Experiments and results}
\label{sec:expres}
\section{Conclusion}

\section{Future work}

\bibliography{paper}
\bibliographystyle{icml2016}

\end{document}

### 2017-12-03 Tokenization/character_extraction_WIP.py
character recognition finished, maybe can be improved upon

@@ -43,6 +43,8 @@ def skeletonize_alternative(img):
             break
     return skel
 
+
+
 fileindex = 4;
 outputpath = os.path.join(dir, '../data/output/')
 filepath = os.path.join(dir, '../data/texts/')
@@ -56,7 +58,9 @@ for index in range(0, 1):# len(os.listdir(filepath))):
         img = cv2.imread(wordpath + word, 0)
         inverted = cv2.bitwise_not(img)
 
-        for angle in (-12, -6, 0, 6, 12):
+        angeled_splits = list()
+
+        for angle in (-7, -3, 0, 3, 7):
             rotatedimg = rotateImage(inverted, angle)
 
             height, width = rotatedimg.shape
@@ -68,10 +72,23 @@ for index in range(0, 1):# len(os.listdir(filepath))):
             #skel = skeletonize_alternative(thresh)
 
 
+
+            #inv = cv2.bitwise_not(thresh)
+            #im2, contours, hierarchy = cv2.findContours(inv,cv2.RETR_TREE,cv2.CHAIN_APPROX_NONE)
+            #res = cv2.drawContours(inv, contours, -1, (123,123,123), 1)
+            #print(contours)
+            #print(hierarchy)
+            #cv2.imshow ("contours", inv)
+            #cv2.waitKey(0)
+
+
+
+
+
             skel2 = skeletonize(thresh/255)
             result = list()
 
-            resultnpy = np.copy(skel)
+            resultnpy = np.copy(thresh)#skel)
             resultnpy.dtype = np.uint8
 
             index1 = 0
@@ -85,8 +102,8 @@ for index in range(0, 1):# len(os.listdir(filepath))):
                     index2 += 1
                 index1 += 1
 
-            for line in skel:
-                print(line)
+            #for line in skel:
+            #    print(line)
 
             for line in resultnpy:
                 print(line)
@@ -94,52 +111,16 @@ for index in range(0, 1):# len(os.listdir(filepath))):
             skel = resultnpy
 
 
-            #row_means = cv2.reduce(skel, 1, cv2.REDUCE_AVG)
-
-            #highest_average_row = 0
-            #highes_row_index = 0;
-            #for index, row_mean in row_means:
-            #    if highest_average_row < row_mean:
-            #        highest_average_row = row_mean
-            #        highes_row_index = index
-            #total = sum(row_means)
-
-            #currentcount = highest_average_row;
-            #currentlines = list()
-            #currentlines.append(index)
[…]

### 2017-12-03 Tokenization/character_extraction_main.py

@@ -1,0 +1,276 @@
+# https://github.com/bsdnoobz/zhang-suen-thinning
+# https://ac.els-cdn.com/S1877050913001464/1-s2.0-S1877050913001464-main.pdf?_tid=23580d56-d5c6-11e7-b3de-00000aab0f27&acdnat=1512043449_af48c0c94bee8350a664995a634d9317
+# https://gist.github.com/jsheedy/3913ab49d344fac4d02bcc887ba4277d
+
+
+import cv2
+import numpy as np
+import os
+import shutil
+import imutils
+from matplotlib import pyplot as plt
+
+#searched for a better skeletonizer and found this library
+from skimage.morphology import skeletonize
+
+dir = os.path.dirname(__file__)
+
+
+def rotateImage(image, angle):
+    #inverted = cv2.bitwise_not(image)
+    rotated = imutils.rotate_bound(image, angle)
+    #result = cv2.bitwise_not(rotated)
+    return rotated
+
+def skeletonize_alternative(img):
+    """ OpenCV function to return a skeletonized version of img, a Mat object"""
+
+    #  hat tip to http://felix.abecassis.me/2011/09/opencv-morphological-skeleton/
+
+    img = img.copy() # don't clobber original
+    skel = img.copy()
+
+    skel[:,:] = 0
+    kernel = cv2.getStructuringElement(cv2.MORPH_CROSS, (3,3))
+
+    while True:
+        eroded = cv2.morphologyEx(img, cv2.MORPH_ERODE, kernel)
+        temp = cv2.morphologyEx(eroded, cv2.MORPH_DILATE, kernel)
+        temp  = cv2.subtract(img, temp)
+        skel = cv2.bitwise_or(skel, temp)
+        img[:,:] = eroded[:,:]
+        if cv2.countNonZero(img) == 0:
+            break
+    return skel
+
+#fileindex = 4;
+#outputpath = os.path.join(dir, '../data/output/')
+#filepath = os.path.join(dir, '../data/texts/')
+#for index in range(0, 1):# len(os.listdir(filepath))):
+
+    #wordpath = os.path.join(outputpath, 'text' + str(fileindex).zfill(3) + '/words/')
+    #characterpath = os.path.join(outputpath, 'text' + str(fileindex).zfill(3) + '/characters/')
+    #print("#########################")
+    #print(wordpath)
+    #for word in sorted(os.listdir(wordpath)):
+
[…]

### 2017-12-03 Tokenization/word_extraction.py

@@ -70,9 +70,6 @@ def preprocess_image(img, file_index = 0):
 
 
     dir = os.path.dirname(__file__)
-
-
-    fileindex = 1;
     outputpath = os.path.join(dir, '../data/output/')
     filepath = os.path.join(dir, '../data/texts/')
     datapath = os.path.join(dir, '../data/')
@@ -86,6 +83,7 @@ def preprocess_image(img, file_index = 0):
 
     file_number = str(file_index).zfill(3)
 
+    print("processing file: " + str(file_number))
 
     #PREPROCESSING
 
@@ -287,14 +285,9 @@ def preprocess_image(img, file_index = 0):
                                 additions.append( newrectangle[0] )
                                 interrupted = True
 
-            print("removals")
-            print(removals)
-            print("additions")
-            print(additions)
+
 
             for element in removals:
-                print('removing')
-                print(element)
                 if element in line:
                     line.remove(element)
                 if element in rectangles_copy:
@@ -302,8 +295,6 @@ def preprocess_image(img, file_index = 0):
 
             for element in additions:
                 line.append(element)
-                print("appending")
-                print(element)
 
         sortedline = sorted(line, key=lambda tup: tup[0])
 
@@ -316,6 +307,9 @@ def preprocess_image(img, file_index = 0):
         linesleft = len(rectangles_copy) != 0
 
     # saving the found rectangles
+
+    extracted_words = list()
+
     ind = 0;
     for line in lines:
         for rectangle in line:
@@ -338,7 +332,7 @@ def preprocess_image(img, file_index = 0):
 
             #extracted_word2 = filledImg[ymin:ymax, xmin:xmax]
 
-            word_directory_path = os.path.join(dir, outputpath + 'text' + sfile_number + '/words/')
+            word_directory_path = os.path.join(dir, outputpath + 'text' + file_number + '/words/')
             wordpath = os.path.join(dir, outputpath + 'text' + file_number + '/words/word' + str(ind).zfill(4) + '.png')
             if not os.path.exists(word_directory_path):
                 os.makedirs(word_directory_path)
@@ -348,7 +342,7 @@ def preprocess_image(img, file_index = 0):
 
             #cv2.rectangle(img,(x,y),(x+w,y+h),(0,255,0),1)
 
-
+            extracted_words.append(extracted_word)
[…]

### 2017-12-03 CharacterRecognition/character_recognition.py
Puzzle everything together

@@ -312,6 +312,6 @@ def examples():
     for ex in reversed(examples):
         for i in range(4):
             print(ex,
-                  img_to_text(settings.EXAMPLE_PATH + ex + '_' + str(i) + ".png", n=62, session=session, _x=_x, _y=_y,
+                  img_to_text(settings.EXAMPLE_CHAR_PATH + ex + '_' + str(i) + ".png", n=62, session=session, _x=_x, _y=_y,
                               h=h))
 

### 2017-12-03 CharacterRecognition/preprocessing.py

@@ -105,7 +105,6 @@ def augmentImage(img, addNoise=True, addRotations=True, addTranslations=True, ad
         for sx in np.arange(min_scale_x, max_scale_x + step, step):
             for sy in np.arange(min_scale_y, max_scale_y + step, step):
                 if not sy == 1 and not sy == 1:
-                    # Todo: translate if not in bounds.
                     if inBounds(down * sy) and inBounds(right * sx):
                         images.append(scaleImage(img, sx, sy))
     return images
@@ -119,8 +118,10 @@ def read_image(file_name):
     :return: A normalized np array with correct dimensions
     """
     img = cv2.imread(file_name, cv2.IMREAD_GRAYSCALE)
-    # Data normalisation and inverting color values
-    img = np.subtract(1, np.divide(img, 255))
+    # Contrast normalisation and inverting color values
+    retVal, img = cv2.threshold(img, 127, 255, cv2.THRESH_BINARY_INV)
+    # Data normalisation
+    img = np.divide(img, 255)
     return cv2.resize(src=img, dsize=settings.SHAPE)
 
 

### 2017-12-03 Postprocessing/language_model.py

@@ -41,9 +41,8 @@ Punctuation needs spaces.
 
 
 def n_gram_model(sentence, words, n=3):
-    previous_words = ' '.join(sentence.split(' ')[-n:])
+    previous_words = ' '.join(sentence[-n:])
     query = previous_words + ' ' + ' / '.join(words)
-    print(query)
     '''
     Give a default value of 0 as probability.
     If Google dataset doesn't contain the given n-gram then the probability will remain 0. 
@@ -58,7 +57,7 @@ def n_gram_model(sentence, words, n=3):
             results[word] = phrase.score
     return results
 
-print(n_gram_model('She loves the', ['girl', 'boy', 'sheep'], n=3))
-print(n_gram_model('He loves the', ['girl', 'boy', 'sheep'], n=3))
-print(n_gram_model('She loves the', ['woman', 'man', 'sheep'], n=2))
+#print(n_gram_model(['She', 'loves', 'the'], ['girl', 'boy', 'sheep'], n=3))
+#print(n_gram_model('He loves the', ['girl', 'boy', 'sheep'], n=3))
+#print(n_gram_model('She loves the', ['woman', 'man', 'sheep'], n=2))
 

### 2017-12-03 Postprocessing/process.py

@@ -1,5 +1,10 @@
+from Tokenization.character_extraction_main import extract_characters
+from Tokenization.word_extraction import preprocess_image
+from CharacterRecognition.character_recognition import img_to_text
+from CharacterRecognition.character_recognition import img_to_prob
+from Postprocessing.language_model import n_gram_model
 from Postprocessing.vocabulary import most_likely_words
-
+import settings
 
 '''
 Postprocessing consists of 2 steps:
@@ -19,8 +24,8 @@ def predict_actual_word(cls_pred_list):
     words = most_likely_words(cls_pred_list)
     print(words)
 
-def sentence_img_to_text():
-    '''
+def sentence_img_to_text(image):
+    """
     Algorithm:
 
     1) Split the sentence in images of words.
@@ -30,11 +35,22 @@ def sentence_img_to_text():
 
     :param images: Image of a sentence
     :return: Text of the sentence
-    '''
-    pass
+    """
+    words = preprocess_image(image)
+    text = [] #Converted image into list of words
+    for word in words:
+        # Find a list of possible words using the vocabularium
+        likey_words = word_img_to_most_likely_words(word)
+        # Find the most likely word using the language model
+        most_likely_word = max(n_gram_model(text, likey_words), key=lambda x: x[1])
+        text.append(most_likely_word)
+    return ' '.join(text)
+
+
+
 
 def word_img_to_most_likely_words(image):
-    '''
+    """
     Algorithm:
 
     1) Split the sentence in images of characters.
@@ -42,21 +58,14 @@ def word_img_to_most_likely_words(image):
     3) Find the most likely words given the cls_pred_list. Now we have a list of words and their probabilities.
     (These probabilities are based on word distances with actual words in the English dictionary)
 
-
     :param image: Image of a word
     :return: A list of pairs, the pairs consist of likely words and their probabilities.
-    '''
-    pass
+    """
+    char_imgs = extract_characters(image)
+    cls_pred_list = []
+    for img in char_imgs:
+        cls_pred_list.append(img_to_prob(img))
+    return most_likely_words(cls_pred_list)
 
-def character_to_cls_pred(image):
-    '''
-    Algorithm:
 
-    1) Feed the image to the trained neural network
-
-    :param image: Image of a character
-    :return: A list of class prediction probabilities.
-    As there are 62 classes, this list will be of length 62.
[…]

### 2017-12-03 Postprocessing/vocabulary.py

@@ -89,12 +89,3 @@ def rand(not_indices, amount):
     else:
         return []
 
-
-# A   p   p   l   e
-# 10 52 52  48  41
-A_pred = rand_array([10])
-p_pred = rand_array([51])
-l_pred = rand_array([47])
-e_pred = rand_array([40])
-
-print(most_likely_words([A_pred, p_pred, p_pred, l_pred, e_pred]))

### 2017-12-03 Tokenization/character_extraction_main.py

@@ -11,8 +11,8 @@ import imutils
 from matplotlib import pyplot as plt
 
 #searched for a better skeletonizer and found this library
+# Scikit-image dependency! (https://www.lfd.uci.edu/~gohlke/pythonlibs/#scikit-image)
 from skimage.morphology import skeletonize
-
 dir = os.path.dirname(__file__)
 
 

### 2017-12-03 settings.py

@@ -17,5 +17,6 @@ PREPROCESSED_CHAR_DATA_TXT_PATH = TXT_PATH + 'preprocessed_chars.txt'
 
 CHAR_PATH = PROJECT_PATH + 'CharacterRecognition/'
 SAVE_PATH = CHAR_PATH + 'Model/model.ckpt'
-EXAMPLE_PATH = CHAR_PATH + 'Examples/'
+EXAMPLE_CHAR_PATH = CHAR_PATH + 'Examples/'
+EXAMPLE_TEXT_PATH = DATA_PATH + 'text/'
 EXPERIMENTS_CHAR_PATH = CHAR_PATH + 'Experiments/'

### 2017-12-04 CharacterRecognition/character_recognition.py
MORE EXPERIMENTSSS

@@ -7,6 +7,7 @@ from CharacterRecognition import graphics as gr
 import settings
 import cv2
 from CharacterRecognition import utils
+import CharacterRecognition.preprocessing as preprocess
 
 '''
 For this project we use the Chars74K dataset. It contains 62 classes, with about 3.4K handwritten characters.
@@ -36,26 +37,9 @@ def open_images():
     data = np.zeros((length, SIZE, SIZE, 1), dtype=np.float32)
     i = 0
     for name in file_names:
-        data[i] = read_image(name)
+        data[i] = preprocess.preprocess_image(preprocess.read_image(name))
         i += 1
     return labels, data, length
-
-
-def read_image(file_name, invert=False):
-    """
-    Read an image.
-    :param invert: Indicates if color values should be inverted.
-    :param file_name: The path to an image
-    :return: A normalized np array with correct dimensions
-    """
-    img = cv2.imread(file_name, cv2.IMREAD_GRAYSCALE)
-    # Data normalisation and inverting color values
-    if img.shape != settings.SHAPE:
-        img = cv2.resize(src=img, dsize=settings.SHAPE)
-    img = np.divide(img, 255)
-    if invert:
-        img = np.subtract(1, img)
-    return np.reshape(img, settings.IMG_SHAPE)
 
 
 def label2vector(label):
@@ -277,20 +261,14 @@ def train_save(epochs):
 def img_to_prob(img, session=None, _x=None, _y=None, h=None):
     """
     Converts an image to a character probabilities.
-    This function assumes there is a Model subdirectory with a trained network model.
-    :param file_name: Path to an image which contains a character.
+    This function assumes there is a Model_500it subdirectory with a trained network model.
+    :param img: Path to an image which contains a character.
     :return: A list containing the probabilities
     of the image being a certain class (representing a letter or number).
     """
-    # img = read_image(file_name, invert=True)
-    if (img.shape != settings.SHAPE):
-        img = cv2.resize(src=img, dsize=settings.SHAPE)
-        img = np.reshape(img, settings.IMG_SHAPE)
+    img = preprocess.preprocess_image(img)
     if session is None:
-        _x, _y, h = create_neural_net(train=False)
-        session = create_session()
-        session.run(tf.global_variables_initializer())
-        restore_session(session)
+        session, _x, _y, h = init_session()
[…]

### 2017-12-04 Postprocessing/process.py

@@ -6,6 +6,7 @@ import settings
 import cv2
 import tensorflow as tf
 import CharacterRecognition.character_recognition as cr
+
 '''
 Postprocessing consists of 2 steps:
 1) Find the most likely word given a list of probabilities for every character in the word. 
@@ -21,12 +22,13 @@ to handle the high error rate of character recognition (+/- 20%).
 
 
 def read_image(file_name):
-    return cv2.imread(file_name, 0)
+    return cv2.imread(file_name, cv2.IMREAD_GRAYSCALE)
 
 
 def predict_actual_word(cls_pred_list):
     words = most_likely_words(cls_pred_list)
     print(words)
+
 
 def sentence_img_to_text(image):
     """
@@ -40,10 +42,10 @@ def sentence_img_to_text(image):
     :param images: Image of a sentence
     :return: Text of the sentence
     """
-    alpha = 0.7 # Indicates importance of correct vocabulary
-    beta = 0.3 # Indicates importance of language model
+    alpha = 0.7  # Indicates importance of correct vocabulary
+    beta = 0.3  # Indicates importance of language model
     words = preprocess_image(image)
-    text = [] #Converted image into list of words
+    text = []  # Converted image into list of words
     # Network for recognising individual characters.
     # These variables need to be this high up in the python project.
     # In order to avoid initialising multiple sessions or neural networks.
@@ -53,10 +55,12 @@ def sentence_img_to_text(image):
         voc_words = word_img_to_most_likely_words(word, session, _x, _y, h)
         # Find the most likely words using the language model
         lang_words = n_gram_model(text, voc_words.keys())
-        most_likely_word = max([(word, alpha*voc_words[word] + beta*prob) for word,prob in lang_words.items()], key=lambda x: x[0])[0]
+        most_likely_word = \
+        max([(word, alpha * voc_words[word] + beta * prob) for word, prob in lang_words.items()], key=lambda x: x[0])[0]
         text.append(most_likely_word)
         print(most_likely_word)
     return ' '.join(text)
+
 
 def word_img_to_most_likely_words(image, session=None, _x=None, _y=None, h=None):
     """
@@ -71,11 +75,10 @@ def word_img_to_most_likely_words(image, session=None, _x=None, _y=None, h=None)
     :return: A list of pairs, the pairs consist of likely words and their probabilities.
     """
     char_imgs = extract_characters(image)
-    if session==None:
[…]

### 2017-12-04 settings.py

@@ -16,7 +16,7 @@ CHAR_DATA_TXT_PATH = TXT_PATH + 'chars.txt'
 PREPROCESSED_CHAR_DATA_TXT_PATH = TXT_PATH + 'preprocessed_chars.txt'
 
 CHAR_PATH = PROJECT_PATH + 'CharacterRecognition/'
-SAVE_PATH = CHAR_PATH + 'Model/model.ckpt'
+SAVE_PATH = CHAR_PATH + 'Model_500it/model.ckpt'
 EXAMPLE_CHAR_PATH = CHAR_PATH + 'Examples/'
 EXAMPLE_TEXT_PATH = DATA_PATH + 'texts/'
 EXPERIMENTS_CHAR_PATH = CHAR_PATH + 'Experiments/'

### 2017-12-04 Report/paper.tex

@@
 \usepackage{subcaption}
+\usepackage{qtree}
 
@@
 where $s_x$ is the horizontal scaling factor and $s_y$ is the vertical scaling factor. 
+\subparagraph{Shearing of data}
+In order to deal with different kinds of slants we add sheared versions of the image to the dataset. 
+Shearing displaces each point in a fixed direction, by an amount proportional to its signed distance from a line that is parallel to that direction \cite{Shear}.
+We used the following shear mapping to achieve this:
+\begin{equation}
+        \begin{bmatrix}
+                1 & s & 0 \\
+                0 & 1 & 0
+        \end{bmatrix}
+\end{equation}
 \subsubsection{Deep neural network}
 \label{sec:dnn}
+Humans have millions of connected neurons in the visual cortices which are capable of image processing.
+We, humans, can recognise most characters without any thought because of this complex neural network.
+For computers, character recognition is a more difficult task.
+We could try to write rule-based systems and define how every character should look like.
+This solution isn't flexible at all and is very hard to define.
+We use a neural network, inspired by how the brain processes images.
+These deep neural networks contain several layers of connected perceptrons, where each connection is defined by weight and each perceptron is defined by a bias.
+The perceptrons take an input, and calculate their output by calculating the activation function with the weights, input and bias.
+Images can be fed to the neural network, which then propagate trough several layers in the neural network and eventually produce as output the probabiliy that the input has a certain class, one of the possible characters.
 
+% I will add a discussion of the neural network once the neural network is finalized and not subject tot change. 
 \subsection{Tokenization}
@@
 \subsection{Natural language processing}
+To cope with some of the high error rates as discussed in section \ref{sec:expres}, there needs to be some kind of correction, in order to produce possibble written text even when a step of the larger algorithm fails.
+In order to do this we adapt popural natural language processing techniques as postprocessing steps.
+Natural language techniques can often be found in speech recognition, language generation and other systems, but this can be helpful here aswell to correct mistakes made in the previous steps of the algorithm.
+As a first step we match words against a dictionary to find out if they make sense, this is discussed in section \ref{sec:voc}.
+The second step checks if those given words make sense in the context, this is discussed in section \ref{sec:lm}.
 
 \subsubsection{Vocabularium}
+\label{sec:voc}
+The neural network does not only produce which character is recognised, but rather a list of probabilities indicating the chance of an image to be a certain character.
+We can use all of this information in this postprocessing step instead of throwing that information away and naively returning the most likely character as the actual character.
+We make a tree, as shown in \ref{fig:wordtree}, where every node is a tuple with a word (sequence of characters) and a probability (likeliness of that word based on the probabilities of individual characters).
+The probability is calculated as a product of probability for each character.
+\begin{figure}
+        %\Tree[.Y [.e [.s] [.5] [.l [.s] [.5]]]]
+        %\Tree[.Y]
+        \Tree [.{(Y, 0.9)} 
+        [.{(Ye , 0.72)} {(Yes, 0.432)} {(Ye5 , 0.288)} ] 
+        [.{(Yl , 0.18)} {(Yls, 0.108)} {(Yl5 , 0.072)} ]
+            ] 
+        \Tree [.{(T, 0.1)} 
+        [.{(Te , 0.08)} {(Tes, 0.048)} {(Te5 , 0.032)} ] 
+        [.{(Tl , 0.02)} {(Tls, 0.012)} {(Tl5 , 0.008)} ]
+            ] 
+\label{fig:wordtree} 
+\caption{A list of the trees for the word 'Yes' with branching factor 2.
+The characters have the following probabilities: $P(Y)=0.9, P(T)=0.1, P(e)=0.8, P(l)=0.2, P(s)=0.6, P(5)=0.4$.
+Other characters have probabilities which are neglible.}
+\end{figure} 
 \subsubsection{Language model}
-
+\label{sec:lm}
 \section{Experiments and results}

### 2017-12-05 Report/paper.tex
Visualisation code for experiments

@@
 \usepackage{qtree}
+\usepackage{listings}
+\usepackage{titlesec}
 
@@
 \label{sec:preproc}
-We make use of 2 types of preprocessing steps: normalization of the data (\ref{par:norm}) and augmentation of the data (\ref{par:aug}).
+We make use of 2 types of preprocessing steps: normalization of the data and augmentation of the data.
 Normalization of data adjusts the original data and data augmentation takes the data and uses augmentation techniques to create variations of the original data.
@@
 We can use all of this information in this postprocessing step instead of throwing that information away and naively returning the most likely character as the actual character.
-We make a tree, as shown in \ref{fig:wordtree}, where every node is a tuple with a word (sequence of characters) and a probability (likeliness of that word based on the probabilities of individual characters).
-The probability is calculated as a product of probability for each character.
 \begin{figure}
-        %\Tree[.Y [.e [.s] [.5] [.l [.s] [.5]]]]
-        %\Tree[.Y]
         \Tree [.{(Y, 0.9)} 
@@
             ] 
-\label{fig:wordtree} 
 \caption{A list of the trees for the word 'Yes' with branching factor 2.
@@
 Other characters have probabilities which are neglible.}
+\label{fig:wordtree} 
 \end{figure} 
+We make a tree, as shown in figure \ref{fig:wordtree}, where every node is a tuple with a word (sequence of characters) and a probability (likeliness of that word based on the probabilities of individual characters).
+The probability is calculated as a product of probability for each character.
+In order to keep the complexity low we only consider the 3 most probable characters.
+This will be the branching factor of every tree.
+There's a tree for every possible starting character.
+Now we have a list of words and their probabilities (every leaf).
+We can calculate the closest matches in the dictionary for each word.
+For this purpose we use the function \lstinline{get_close_matches} from the python library difflib.
+We use function to find close matches in the English dictionary and their corresponding score.
+We multiply this score with our previously calculated probability.
+This leaves us with a list of correct English words and a score for every word.
+If we only want to convert the image of one word into text, we can just return the word with the highest score.
 \subsubsection{Language model}
 \label{sec:lm}
+Besides checking the syntax of words, we can also look at the context.
+This is were language models are useful.
+We can check the previous words and calculate the likeness of a word to be the next word.
+N-gram models are used to calculate probabilities given n previous words.
+% TODO: More elaborate explanation of n-gram models. I need the internetz for this. 
+In order for these n-gram models to be useful.
+We need a dataset.
+Google has published n-gram corpora to the public and is free to use.
+Querying these datasets takes several minutes, and this needs to happen for every word.
+Instead we use a python library phrasefinder to make online queries to these corpora.
+This greatly increases speed but has the disadvantage of requiring internet connection.
+Now we can make queries such as "I like dogs / cats / sheep".
+Using the Google corpora probabilities are calculated with n-gram models for each of the 3 words.
+Now we have a probability for every word based on the context.
+Combining the results with the probabilities calculated in section \ref{sec:voc}, we can 
+calculate a new score.
+The word with the highest score is accepted as written word.
+By iterating over all of the words, we now have converted the image with written text into text.
+In the next section we will
+further discuss the results of these techniques and some experiments to check the performance and impact of optimizations. 
 \section{Experiments and results}

### 2017-12-05 CharacterRecognition/experiment_results.py

@@ -1,16 +1,75 @@
 import settings
 import os
+import numpy as np
+import matplotlib.pyplot as plt
+from pylab import savefig
+
+# from CharacterRecognition.experiments import ITERATIONS
 experiments = ['base', 'batchsize', 'decay', 'filtersize', 'keepprob', 'learningrate', 'preprocess']
+
+ITERATIONS = 4
 
 
 def get_experiment_results(experiment_name):
-    """
-    Finds the results of an experiment. 
-    :param experiment_name: The name of the experiment
-    :return: A list of tuples, each tuple contains a configuration (string) and the results (4 lists).
-    """
+    confs, accuracy_avgs, time_avgs = [], [], []
+    for conf, dir in find_experiment_location(experiment_name):
+        accuracy = []
+        time = []
+        for i in range(ITERATIONS):
+            file_name = os.path.join(dir, '{}_' + str(i) + '.txt')
+            accuracy.append(np.loadtxt(file_name.format('accuracy')))
+            time.append(np.loadtxt(file_name.format('time')))
+        confs.append(conf)
+        accuracy_avgs.append(np.mean(accuracy, axis=0))
+        time_avgs.append(np.mean(time, axis=0))
+    return confs, accuracy_avgs, time_avgs
 
-    for root, dirs, files in os.walk(settings.EXPERIMENTS_CHAR_PATH):
-        print(root, dirs, files)
 
-get_experiment_results("base")
+def get_configuration_results(experiment_name, conf):
+    accuracies, times = [], []
+    _, dir = find_experiment_conf_location(experiment_name, conf)
+    for i in range(ITERATIONS):
+        file_name = os.path.join(dir, '{}_' + str(i) + '.txt')
+        accuracies.append(np.loadtxt(file_name.format('accuracy')))
+        times.append(np.loadtxt(file_name.format('time')))
+    return accuracies, times
+
+
+def find_experiment_location(experiment_name):
+    return [(x[0].split('/')[-1], x[0]) for x in os.walk(settings.EXPERIMENTS_CHAR_PATH) if experiment_name in x[0]]
+
+
+def find_experiment_conf_location(experiment_name, conf):
+    return find_experiment_location('_'.join([experiment_name, str(conf)]))[0]
+
+
+def visualise(accuracies, times, name='', show=True, save=False, labels=None):
+    plots = []
+    for i in range(len(accuracies)):
+        if labels is None:
[…]

### 2017-12-05 Postprocessing/vocabulary.py

@@ -54,7 +54,7 @@ def most_likely_words(cls_pred_list):
     words = {}
     for characters, probability in most_possible_characters[-3:]:
         for word, score in correct_written_words(characters):
-            words[word] = probability
+            words[word] = probability * score
     return words
 
 

### 2017-12-05 settings.py

@@ -16,7 +16,8 @@ CHAR_DATA_TXT_PATH = TXT_PATH + 'chars.txt'
 PREPROCESSED_CHAR_DATA_TXT_PATH = TXT_PATH + 'preprocessed_chars.txt'
 
 CHAR_PATH = PROJECT_PATH + 'CharacterRecognition/'
-SAVE_PATH = CHAR_PATH + 'Model_500it/model.ckpt'
+SAVE_PATH = CHAR_PATH + 'Model/model.ckpt'
 EXAMPLE_CHAR_PATH = CHAR_PATH + 'Examples/'
 EXAMPLE_TEXT_PATH = DATA_PATH + 'texts/'
 EXPERIMENTS_CHAR_PATH = CHAR_PATH + 'Experiments/'
+EXPERIMENTS_CHAR_GRAPHS_PATH = CHAR_PATH + 'Graphs/'

### 2017-12-07 Report/paper.tex
moarrr verslag

@@
 \usepackage{hyperref}
-\usepackage{icml2016} 
+\usepackage[accepted]{icml2016} 
 \usepackage{amsmath}
@@
   \includegraphics[width=\linewidth]{images/good_char}
-  \caption{Picture of a 'z' which resembles a 3.}
+  \caption{Picture of a clean 'z'.}
 \end{subfigure}
@@
 This is were language models are useful.
-We can check the previous words and calculate the likeness of a word to be the next word.
-N-gram models are used to calculate probabilities given n previous words.
-% TODO: More elaborate explanation of n-gram models. I need the internetz for this. 
-In order for these n-gram models to be useful.
-We need a dataset.
-Google has published n-gram corpora to the public and is free to use.
-Querying these datasets takes several minutes, and this needs to happen for every word.
-Instead we use a python library phrasefinder to make online queries to these corpora.
+We can check the previous words and calculate the likeness of a word to be the next word in the context.
+N-gram models are used to calculate probabilities given n-1 previous words.
+The probability is computed as $P(w_i | w_{i-n},...,w_{i-1})$.
+We use the Markov assumption here, we can approximate the probability of a word so the word only depends on the previous n-1 words.
+This probability is calculated as follows:
+\begin{equation}
+        P(w_i | w_{i-n},...,w_{i-1}) = \frac{count(w_{i-n},...,w_{i-1},w_{i})}{count(w_{i-n},...,w_i)}
+\end{equation}
+where $w_i$ is the word of which we want to calculate the probability and $count(w_i,...w_{i+k})$ counts the amount of occurences of the sequence of words $w_i,...,w_{i+k}$.
+That amount of occurences is the same as counting the amount of n-grams with that sequence of words.
+An n-gram is a contiguous sequence of n items from a given sequence of text or speech.
+\cite{ngram}
+We use the n-grams from the Google dataset for this purpose.
+Querying these datasets locally takes several minutes, and this happens for every word.
+Instead we use a Python library phrasefinder to make online queries to these corpora.
 This greatly increases speed but has the disadvantage of requiring internet connection.
@@
 Using the Google corpora probabilities are calculated with n-gram models for each of the 3 words.
-Now we have a probability for every word based on the context.
+This gives us a probability for every word based on the context.
 Combining the results with the probabilities calculated in section \ref{sec:voc}, we can 
@@
 \section{Experiments and results}
+%TODO
+\subsection{Character Recognition}
 \label{sec:expres}
+We tested a lot of different configurations for the deep neural network in order to find a good functioning neural network given the low amounts of data.
+A problem which arrised when trying to test many diffeent configurations was the long time it takes to train a neural network.
+In order to keep these experiments from running several months, we added some small constraints.
+Next it's important to note that a lot of non-determinism is part of these neural networks.
+When running the neural network twice with the same configuration, it could have completely different results.
+In order to have meaningful results which run in a reasonable amount of time we ran every experiment 4 times but limited the amount of epochs to 200.
+This setup is important to note, as some results might be quite different if we run more tests or use more iterations.
+When discussing these results we take the best result of the 4 training results.
+We use this result as we are only interested in the potential of the neural network.
+If we'd let those experiments run for more iterations, they would reach the same accuracy because of the noise added to the neural network.
+The different configurations and results will be discussed next.
+\subsubsection{Experiments}
+\paragraph{Preprocessing techniques}
+Figure \ref{fig:preprocess} shows that preprocessing techniques have great influence on the performance of the neural network.
+We can see that the accuracy is much higher when using all of the preprocessing techniques.
+We achieve 72\% accuracy with all of the preprocessing techniques versus 35\% without any preprocessing techniques, which is a huge difference.
+We can also see the individual impact of preprocessing in the figure.
+Adding shearing (64\%) to handle different amounts of slant, adding noise (58\%) to cope to make the recognition more robust and adding scaling (59\%) to handle different handwriting sizes was the most effective to increase performance.
+Adding rotations (40\%) and translations (50\%) was less effective (but still effective).
+Adding these preprocessing techniques increased the amount of data greatly from 3410 examples to 37000 examples.
+This improves the performance but also increases the training time of our network significantly.
+One remarkable result is that the training time of preprocessing with all techniques enabled, takes less long than some other results.
+When we look at the average however, the training time is the highest for this case.
+\paragraph{Filter amount}
+% Not sure if I should discuss this, pretty boring results. 
+\paragraph{Filter size}
+In this experiment we check the influence of choosing different sizes of filters for the convolutional network.
+The filter size determines the size of our filters for convolving the input images.
+In figure \ref{fig:filtersize} we can see the time increases greatly, because more surrounding values have to be considered.
+The accuracy increases slightly when using bigger filters.
+Using different filter sizes allow other type of features to be extracted.
+% TODO: this is hard to explain..., maybe just leave this one out.
+% Convolutation layers allow extracting features from the image. 
+% Go over this again: good explanation: (See: http://www.deeplearningbook.org/)
+% Pixels far away from eachother are less correlated. This is why we subsample filter responses.
+\paragraph{Learning rate}
+Figure \ref{fig:learningrate} shows the impact of learning rate on the training of our neural network. 
+For this experiment we do have to note that we're using the Adam optimizer to train our neural network which computes individual learning rates for every parameter every training step.
+The learning rate which we pass is the initial learning rate.
+Choosing a high learning rate $10^-3$ allows the neural network to converge to a high accuracy (76 \%) very fast, but once reached the accuracy fails to grow any more.
+This in contrast to higher learning rates which reach a good accuracy more slowly but find a more optimal solution.
+Learning rate $10^-4$ reaches an accuracy of 68\% after 200 iterations but is still growing.
+Learning rate $10^-5$ only reaches an accuracy of 5\% after 200 iterations but should find a better solution eventually than the other two configurations.
+\paragraph{Batch size}
+In figure \ref{fig:batchsize} we can see the impact of choosing different batch sizes which are fed to the neural network.
+In order to increase to speed at which the neural network is trained, we feed batches of input images to the neural network instead of one image at a time.
+The backpropagation in order to calculate the weight adjustment only happens after every batch is fed.
+In the figure we can see bigger batch sizes increase the execution speed.
+However these differences are quites small.
+The configuration with batch size executes 200 iterations in one minute less than a configuration with batch size 128.
+These batch sizes also have an effect on our performance.
+When using small batch sizes backpropagation happens more frequently, which implies our weights will be more finetuned.
+The configuration with batch size 128 obtains an accuracy of 10\% higher than batch size 1024.
+\paragraph{Weight decay}
+In order to regulize the neural network we add weight decay to the network loss function in order to penalize big weights.
+The impact of different decays can be seen in \ref{fig:decay}.
+A high decay functions best (71\% accuracy) because we have a small amount of training examples.
+Without decay we only reach an accuracy of 65\%.
+%TODO: higher decay -> higher accuracy
+\paragraph{Optimizer}
+% Only useful optimizer is Adam. I'm not sure why, or if we shoul compare these at all. 
+\paragraph{Dropout}
+% Maybe redo this experiment?
+It's not clear at all at this time whether dropout has an actual effect on the accuracy.
+I believe it helps find a better global solution, but often the accuracy will be lower as we're only using half of the weights.
+Further research is required.
+\begin{figure}
+\centering
+    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/preprocess.png}
+    \caption{Impact of various preprocessing techniques.}
+    \label{fig:preprocess}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/filtersize.png}
+    \caption{Impact of various filter sizes.}
+    \label{fig:filtersize}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/learningrate.png}
+    \caption{Impact of different learning rates.}
+    \label{fig:learningrate}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/batchsize.png}
+    \caption{Impact of different batch sizes.}
+    \label{fig:batchsize}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/decay.png}
+    \caption{Impact of weight decay regularisation.}
+    \label{fig:decay}
+\end{figure}
+\subsubsection{Final results}
+% Add graph with several runs of the final network. Discuss which parameters we took. 
+% Maybe some more visualisation of mispredicted characters? Oh this isn't part of char recognition...
+% TODO TODO
 \section{Conclusion}

### 2017-12-07 CharacterRecognition/experiment_results.py

@@ -1,4 +1,5 @@
 import settings
+import math
 import os
 import numpy as np
 import matplotlib.pyplot as plt
@@ -10,19 +11,33 @@ experiments = ['base', 'batchsize', 'decay', 'filtersize', 'keepprob', 'learning
 ITERATIONS = 4
 
 
-def get_experiment_results(experiment_name):
-    confs, accuracy_avgs, time_avgs = [], [], []
+def get_experiment_results(experiment_name, max=True):
+    confs, accuracies, times = [], [], []
     for conf, dir in find_experiment_location(experiment_name):
-        accuracy = []
-        time = []
+        # Find the best result (result for which accuracy is highest in the end)
+        avg_accuracy = []
+        avg_time = []
+        max_accuracy = [0]
+        max_time = []
         for i in range(ITERATIONS):
             file_name = os.path.join(dir, '{}_' + str(i) + '.txt')
-            accuracy.append(np.loadtxt(file_name.format('accuracy')))
-            time.append(np.loadtxt(file_name.format('time')))
+            accuracy = np.loadtxt(file_name.format('accuracy'))
+            time = np.loadtxt(file_name.format('time'))
+            if max:
+                if accuracy[-1] > max_accuracy[-1]:
+                    max_accuracy = accuracy
+                    max_time = time
+            else:
+                avg_accuracy.append(accuracy)
+                avg_time.append(time)
         confs.append(conf)
-        accuracy_avgs.append(np.mean(accuracy, axis=0))
-        time_avgs.append(np.mean(time, axis=0))
-    return confs, accuracy_avgs, time_avgs
+        if max:
+            accuracies.append(max_accuracy)
+            times.append(max_time)
+        else:  # Calculate mean if not max
+            accuracies.append(np.mean(avg_accuracy, axis=0))
+            times.append(np.mean(avg_time, axis=0))
+    return confs, accuracies, times
 
 
 def get_configuration_results(experiment_name, conf):
@@ -46,6 +61,10 @@ def find_experiment_conf_location(experiment_name, conf):
 def visualise(accuracies, times, name='', show=True, save=False, labels=None):
     plots = []
     for i in range(len(accuracies)):
+        if labels[i]:
+            print("{}: {:.0f}%, {:.0f}".format(labels[i], 100 * accuracies[i][-1], times[i][-1]))
+        else:
+            print(accuracies[i][-1], times[i][-1])
[…]

### 2017-12-08 CharacterRecognition/character_recognition.py
method for creating training data

@@ -66,9 +66,10 @@ def vector2label(vector):
     return 1 + vector.index(1)
 
 
-def new_weights(shape):
+def new_weights(shape, weights):
     w = tf.get_variable('weights', shape, initializer=tf.truncated_normal_initializer(stddev=.1))
-    weights.append(w)
+    if weights:
+        weights.append(w)
     return w
 
 
@@ -76,10 +77,10 @@ def new_biases(length):
     return tf.get_variable('biases', shape=[length], initializer=tf.constant_initializer(0.1))
 
 
-def new_conv_layer(name, input, num_filters, filter_size, num_in_channels, use_pooling=True):
+def new_conv_layer(name, input, num_filters, filter_size, num_in_channels, global_weights, use_pooling=True):
     with tf.variable_scope("conv_layer_" + str(name)):
         shape = [filter_size, filter_size, num_in_channels, num_filters]
-        weights = new_weights(shape)
+        weights = new_weights(shape, global_weights)
         biases = new_biases(num_filters)
         layer = tf.nn.conv2d(input=input, filter=weights, strides=[1, 1, 1, 1], padding='SAME') + biases
         if use_pooling:
@@ -89,9 +90,9 @@ def new_conv_layer(name, input, num_filters, filter_size, num_in_channels, use_p
         return tf.nn.relu(layer)
 
 
-def new_fc_layer(name, input, num_in, num_out):
+def new_fc_layer(name, input, num_in, num_out, global_weights):
     with tf.variable_scope("fc_layer_" + str(name)):
-        weights = new_weights(shape=[num_in, num_out])
+        weights = new_weights(shape=[num_in, num_out], weights=global_weights)
         biases = new_biases(length=num_out)
         return tf.nn.relu(tf.matmul(input, weights) + biases)
 
@@ -125,7 +126,7 @@ def get_data():
     return (train_X, train_Y), (validation_X, validation_Y), (test_X, test_Y)
 
 
-def create_neural_net(train=True, base=1, filter_size=FILTER_SIZE, keep_prob=KEEP_PROB):
+def create_neural_net(global_weights=None, train=True, base=1, filter_size=FILTER_SIZE, keep_prob=KEEP_PROB):
     """
     Builds a neural network which can be trained with an optimizer to recognise characters.
     :param train: Indicates if the net is used for training.
[…]

### 2017-12-08 Tokenization/character_extraction_main.py

@@ -6,9 +6,7 @@
 import cv2
 import numpy as np
 import os
-import shutil
 import imutils
-from matplotlib import pyplot as plt
 
 # Found a good skeltonize-algorithm in the scikit-library
 # Scikit-image dependency! (https://www.lfd.uci.edu/~gohlke/pythonlibs/#scikit-image)

### 2017-12-08 Tokenization/segmentation_correction.py

@@ -1,0 +1,104 @@
+import settings
+import cv2
+import numpy as np
+import Tokenization.character_extraction_main as chrextr
+from pathlib import Path
+import sys
+
+np.set_printoptions(threshold=np.nan)
+
+WINDOW_SIZE = 5
+MATRIX_DX = 5
+START_LINE = 18
+
+
+def open_images(start=0, amount=1000):
+    file_entries = [x.split(' ') for x in
+                    open(settings.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[
+                    start + START_LINE:START_LINE + start + amount]]
+    print("Reading new dataset of {} images".format(len(file_entries)))
+    images = []
+    files = [x[0] for x in file_entries]
+    for file_name in files:
+        parts = file_name.split('-')
+        file_path = '/'.join([parts[0], parts[0] + '-' + parts[1], file_name + '.png'])
+        images.append((file_path, cv2.imread(settings.CHAR_SEGMENTATION_DATA_PATH + file_path)))
+    return images
+
+
+def find_start(file_name):
+    file_entries = [x.split(' ') for x in
+                    open(settings.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[START_LINE:]]
+    for (i, entry) in enumerate(file_entries):
+        if entry[0] + '.png' == file_name:
+            return i
+    raise FileNotFoundError
+
+
+def create_training_data(start=0, amount=1000):
+    show_range = 1
+    images = open_images(start, amount)
+    n = len(images)
+    with open(settings.CHAR_SEGMENTATION_TRAINING_OUT_PATH, "a") as out_file:
+        for (i, (img_path, img)) in enumerate(images):
+            print("Showing image number {} of {}".format(i, n))
+            split_data = []
+            splits = chrextr.extract_character_separations(img[:, :, 0])
+            for (x, y) in splits:
+                split = False
+                img_tmp = img.copy()
+                for y in range(-show_range, show_range + 1):
+                    img_tmp[:, x + y] = [0, 0, 255]
+                cv2.imshow("img", img_tmp)  # Present splitpoint to user
+                key = cv2.waitKey(0)
+                if key == 13 or key == 32:  # User decides x is a splitpoint
+                    split = True
+                    # Create pixel matrix around splitpoint
+                elif key == 27:  # escape
+                    return
+                else:  # user decides x is not a splitpoint
+                    pass
[…]

### 2017-12-08 settings.py

@@ -27,3 +27,7 @@ EXAMPLE_CHAR_PATH = CHAR_PATH + 'Examples/'
 EXAMPLE_TEXT_PATH = DATA_PATH + 'texts/'
 EXPERIMENTS_CHAR_PATH = CHAR_PATH + 'Experiments/'
 EXPERIMENTS_CHAR_GRAPHS_PATH = CHAR_PATH + 'Graphs/'
+
+CHAR_SEGMENTATION_DATA_TXT_PATH = TXT_PATH + 'words.txt'
+CHAR_SEGMENTATION_DATA_PATH = DATA_PATH + 'words/'
+CHAR_SEGMENTATION_TRAINING_OUT_PATH = DATA_PATH + 'segmentation/segmentation.txt'

### 2017-12-08 Tokenization/segmentation_correction.py
improve persistence

@@ -4,6 +4,7 @@ import numpy as np
 import Tokenization.character_extraction_main as chrextr
 from pathlib import Path
 import sys
+import pickle
 
 np.set_printoptions(threshold=np.nan)
 
@@ -16,7 +17,7 @@ def open_images(start=0, amount=1000):
     file_entries = [x.split(' ') for x in
                     open(settings.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[
                     start + START_LINE:START_LINE + start + amount]]
-    print("Reading new dataset of {} images".format(len(file_entries)))
+    print("Reading new dataset of {} images, starting at image {}".format(len(file_entries), start))
     images = []
     files = [x[0] for x in file_entries]
     for file_name in files:
@@ -39,7 +40,7 @@ def create_training_data(start=0, amount=1000):
     show_range = 1
     images = open_images(start, amount)
     n = len(images)
-    with open(settings.CHAR_SEGMENTATION_TRAINING_OUT_PATH, "a") as out_file:
+    with open(settings.CHAR_SEGMENTATION_TRAINING_OUT_PATH, "ab") as out_file:
         for (i, (img_path, img)) in enumerate(images):
             print("Showing image number {} of {}".format(i, n))
             split_data = []
@@ -55,22 +56,53 @@ def create_training_data(start=0, amount=1000):
                     split = True
                     # Create pixel matrix around splitpoint
                 elif key == 27:  # escape
+                    pickle.dump(i, out_file) # Write length to pickle
                     return
                 else:  # user decides x is not a splitpoint
                     pass
                 split_data.append((x, split))
-            out_file.write(img_path + ' ' + str(split_data) + '\n')
+            pickle.dump((start+i, img_path, split_data), out_file)
 
 
 def start_training(requested_start=0):
     path = Path(settings.CHAR_SEGMENTATION_TRAINING_OUT_PATH)
     if path.exists():
-        last_entry = path.open().readlines()[-1].split(' ')[0].split('/')[-1]
-        start = find_start(last_entry)
+        lines = path.open().readlines()
+        if len(lines) > 0:
+            last_entry = lines[-1].split(' ')[0].split('/')[-1]
+            start = find_start(last_entry)
+        else:
+            start = requested_start
     else:
         start = requested_start
     create_training_data(start + 1)
 
+
+def read_training_data():
+    with open(settings.CHAR_SEGMENTATION_TRAINING_OUT_PATH, "rb") as in_file:
[…]

### 2017-12-08 CharacterRecognition/character_recognition.py
Import optimisations

@@ -4,7 +4,7 @@ import tensorflow as tf
 from sklearn.utils import shuffle
 from sklearn.model_selection import train_test_split
 from CharacterRecognition import graphics as gr
-import settings
+import definitions
 import cv2
 from CharacterRecognition import utils
 import CharacterRecognition.preprocessing as preprocess
@@ -20,11 +20,8 @@ KEEP_PROB = 0.5
 BATCH_SIZE = 128
 FILTER_SIZE = 5
 NUM_CLASSES = 62
-SIZE = settings.SIZE
-NUM_CHANNELS = settings.NUM_CHANNELS
-
-# Global variable
-weights = []
+SIZE = definitions.SIZE
+NUM_CHANNELS = definitions.NUM_CHANNELS
 
 
 # Helper functions
@@ -33,7 +30,7 @@ def open_images():
     Opens the dataset and preprocesses the images.
     :return: The labels of images, numpy pixel arrays with the image data, amount of images
     """
-    file_names = open(settings.PREPROCESSED_CHAR_DATA_TXT_PATH, 'r').read().splitlines()
+    file_names = open(definitions.PREPROCESSED_CHAR_DATA_TXT_PATH, 'r').read().splitlines()
     print("Reading dataset of {} images".format(len(file_names)))
     file_names = shuffle(file_names)
     labels = [int(x.split('-')[1]) for x in file_names]
@@ -241,7 +238,7 @@ def save_session(session):
     :return: None 
     """
     saver = tf.train.Saver()
-    saver.save(session, settings.SAVE_PATH)
+    saver.save(session, definitions.SAVE_PATH)
     session.close()
 
 
@@ -250,7 +247,7 @@ def restore_session(session):
     Create a session initialized with the session as defined in the SAVE_PATH.
     :return: None
     """
-    tf.train.Saver().restore(session, settings.SAVE_PATH)
+    tf.train.Saver().restore(session, definitions.SAVE_PATH)
 
 
 def restore_train_save(epochs):
@@ -330,7 +327,7 @@ def examples():
     for ex in reversed(examples):
         for i in range(4):
             print(ex,
-                  img_to_text(cv2.imread(settings.EXAMPLE_CHAR_PATH + ex + '_' + str(i) + ".png", 0),
+                  img_to_text(cv2.imread(definitions.EXAMPLE_CHAR_PATH + ex + '_' + str(i) + ".png", 0),
                               sessionargs, n=1))
 
 train_net(200, restore=False)

### 2017-12-08 CharacterRecognition/experiment_nets.py

@@ -1,4 +1,4 @@
-import settings
+import definitions
 import tensorflow as tf
 from CharacterRecognition.character_recognition import new_conv_layer
 from CharacterRecognition.character_recognition import new_fc_layer
@@ -16,8 +16,8 @@ KEEP_PROB = 0.5
 BATCH_SIZE = 512
 FILTER_SIZE = 3
 NUM_CLASSES = 62
-SIZE = settings.SIZE
-NUM_CHANNELS = settings.NUM_CHANNELS
+SIZE = definitions.SIZE
+NUM_CHANNELS = definitions.NUM_CHANNELS
 
 f = 'fully connected layer'
 fx = 'fully connected layer with dropout'

### 2017-12-08 CharacterRecognition/experiment_results.py

@@ -1,5 +1,4 @@
-import settings
-import math
+import definitions
 import os
 import numpy as np
 import matplotlib.pyplot as plt
@@ -51,7 +50,7 @@ def get_configuration_results(experiment_name, conf):
 
 
 def find_experiment_location(experiment_name):
-    return [(x[0].split('/')[-1], x[0]) for x in os.walk(settings.EXPERIMENTS_CHAR_PATH) if experiment_name in x[0]]
+    return [(x[0].split('/')[-1], x[0]) for x in os.walk(definitions.EXPERIMENTS_CHAR_PATH) if experiment_name in x[0]]
 
 
 def find_experiment_conf_location(experiment_name, conf):
@@ -77,7 +76,7 @@ def visualise(accuracies, times, name='', show=True, save=False, labels=None):
     axes = plt.gca()
     axes.set_ylim([0, 0.9])
     if save and name != '':
-        savefig(os.path.join(settings.EXPERIMENTS_CHAR_GRAPHS_PATH, name + '.png'))
+        savefig(os.path.join(definitions.EXPERIMENTS_CHAR_GRAPHS_PATH, name + '.png'))
     if show:
         plt.show()
 

### 2017-12-08 CharacterRecognition/experiments.py

@@ -1,4 +1,4 @@
-import settings
+import definitions
 from numpy import savetxt
 from CharacterRecognition import character_recognition
 from CharacterRecognition import preprocessing
@@ -168,7 +168,7 @@ def save_output(name, accuracies, time, iteration=None):
     extension = '.txt'
     if iteration is not None:
         extension = '_' + str(iteration) + extension
-    out_path = settings.EXPERIMENTS_CHAR_PATH + name + '/'
+    out_path = definitions.EXPERIMENTS_CHAR_PATH + name + '/'
     try:
         os.makedirs(out_path)
     except OSError:

### 2017-12-08 CharacterRecognition/preprocessing.py

@@ -3,7 +3,7 @@ import time as t
 import cv2
 import numpy as np
 
-import settings
+import definitions
 
 '''
 For this project we use the Chars74K dataset. It contains 64 classes, with about 74K images.
@@ -11,7 +11,7 @@ For this project we use the Chars74K dataset. It contains 64 classes, with about
 
 # Global constants
 start = t.time()
-SIZE = settings.SIZE
+SIZE = definitions.SIZE
 EMPTY_VALUE = 0.0
 WRITE_VALUE = 1.0
 
@@ -62,22 +62,22 @@ def noisyImage(img, stddev):
 # Helper functions
 def rotateImage(image, angle):
     rotation_matrix = cv2.getRotationMatrix2D(center=(SIZE / 2, SIZE / 2), angle=angle, scale=1.0)
-    return cv2.warpAffine(image, rotation_matrix, settings.SHAPE)
+    return cv2.warpAffine(image, rotation_matrix, definitions.SHAPE)
 
 
 def translateImage(image, tx, ty):
     translation_matrix = np.float32([[1, 0, tx], [0, 1, ty]])
-    return cv2.warpAffine(image, translation_matrix, settings.SHAPE)
+    return cv2.warpAffine(image, translation_matrix, definitions.SHAPE)
 
 
 def scaleImage(image, sx, sy):
     scale_matrix = np.float32([[sx, 0, 0], [0, sy, 0]])
-    return cv2.warpAffine(image, scale_matrix, settings.SHAPE)
+    return cv2.warpAffine(image, scale_matrix, definitions.SHAPE)
 
 
 def shearImage(image, s):
     shear_matrix = np.float32([[1, s, 0], [0, 1, 0]])
-    return cv2.warpAffine(image, shear_matrix, settings.SHAPE)
+    return cv2.warpAffine(image, shear_matrix, definitions.SHAPE)
 
 def erodeImage(image):
     element = cv2.getStructuringElement(cv2.MORPH_RECT,(3,3))
@@ -144,26 +144,26 @@ def read_image(file_name):
 
 
 def preprocess_image(img, inverse=False):
-    if img.shape != settings.SHAPE:
-        img = cv2.resize(src=img, dsize=settings.SHAPE)
+    if img.shape != definitions.SHAPE:
+        img = cv2.resize(src=img, dsize=definitions.SHAPE)
     # Contrast normalisation and inverting color values
     # We add a simple threshold for distinguishing the foreground and background.
[…]

### 2017-12-08 CharacterRecognition/utils.py

@@ -1,4 +1,3 @@
-import sklearn
 
 def cls2str(x):
     if(x<11):

### 2017-12-08 Postprocessing/language_model.py

@@ -1,4 +1,4 @@
-from Postprocessing.lib import phrasefinder
+from Postprocessing import phrasefinder
 
 '''
 See: http://phrasefinder.io/documentation

### 2017-12-08 Postprocessing/process.py

@@ -1,12 +1,11 @@
-from Tokenization.character_extraction_main import extract_characters
-from Tokenization.word_extraction import preprocess_image
-from Tokenization.character_combinator import evaluate_character_combinations
+import cv2
+
+import CharacterRecognition.character_recognition as cr
 from Postprocessing.language_model import n_gram_model
 from Postprocessing.vocabulary import most_likely_words
-import settings
-import cv2
-import tensorflow as tf
-import CharacterRecognition.character_recognition as cr
+from Tokenization.character_combinator import evaluate_character_combinations
+from Tokenization.character_extraction_main import extract_characters
+from Tokenization.word_extraction import preprocess_image
 
 '''
 Postprocessing consists of 2 steps:

### 2017-12-08 Tokenization/character_combinator.py

@@ -1,4 +1,3 @@
-import CharacterRecognition.character_recognition as crj
 import Tokenization.character_normalizer as normalizer
 
 

### 2017-12-08 Tokenization/character_extraction_WIP.py

@@ -3,14 +3,13 @@
 # https://gist.github.com/jsheedy/3913ab49d344fac4d02bcc887ba4277d
 
 
-import cv2
-import numpy as np
 import os
-import shutil
-import imutils
-from matplotlib import pyplot as plt
 
-#searched for a better skeletonizer and found this library
+import cv2
+import imutils
+import numpy as np
+from matplotlib import pyplot as plt
+# searched for a better skeletonizer and found this library
 from skimage.morphology import skeletonize
 
 dir = os.path.dirname(__file__)

### 2017-12-08 Tokenization/main_preprocessing.py

@@ -1,7 +1,7 @@
 
-import word_extraction
-import character_extraction_main
-import character_normalizer
+import Tokenization.word_extraction as word_extraction
+import Tokenization.character_extraction_main as character_extraction_main
+import Tokenization.character_normalizer as character_normalizer
 import cv2
 import os
 

### 2017-12-08 Tokenization/test.py

@@ -1,13 +1,10 @@
 #http://opencvpython.blogspot.be/2012/05/skeletonization-using-opencv-python.html
 
-import cv2
-import numpy as np
 import os
-import shutil
+
+import cv2
 import imutils
-from matplotlib import pyplot as plt
-
-
+import numpy as np
 
 dir = os.path.dirname(__file__)
 

### 2017-12-08 Tokenization/word_extraction.py

@@ -3,10 +3,11 @@
 # http://opencv-python-tutroals.readthedocs.io/en/latest/py_tutorials/py_imgproc/py_contours/py_contour_features/py_contour_features.html
 # https://codereview.stackexchange.com/questions/31352/overlapping-rectangles
 
+import os
+
 import cv2
 import numpy as np
-import os
-import shutil
+
 
 # checks if rectangle 2 inside rectangle 1
 def rectangle_contains_rectangle(rectangle1, rectangle2):

### 2017-12-08 create_data.py

@@ -1,13 +1,10 @@
 import definitions
 import cv2
-import numpy as np
 import Tokenization.character_extraction_main as chrextr
 from pathlib import Path
 import sys
 import pickle
 
-WINDOW_SIZE = 5
-MATRIX_DX = 5
 START_LINE = 18
 
 
@@ -76,43 +73,6 @@ def start_training(requested_start=0):
     create_training_data(start + 1)
 
 
-def read_training_data():
-    with open(definitions.CHAR_SEGMENTATION_TRAINING_OUT_PATH, "rb") as in_file:
-        entries = []
-        EOF = False
-        while not EOF:  # Non ideal way of reading files... But easiest way to make program fool proof.
-            try:
-                entry = pickle.load(in_file)
-                entries.append(entry)
-            except EOFError:
-                EOF = True
-        for (i, file_name, split_data) in entries:
-            print(i, file_name, split_data)
-            # file_entries = [x.split(' ') for x in
-            # open(definitions.CHAR_SEGMENTATION_TRAINING_OUT_PATH, 'r').read().splitlines()]
-        '''
-        start = file_entries[0][0]
-        print(start)
-        print("Reading new dataset of {} images".format(len(file_entries),))
-        for entry in file_entries:
-            print(entry)
-            i = entry[0]
-            file_name = entry[1]
-            split_data = entry[2]
-            print(i, file_name, split_data)
-        '''
-
-
-def feature_extractor(img, x):
-    # This matrix has the shape: (2*MATRIX_DX, IMG_HEIGHT)
-    # Where IMG_HEIGHT is the height of the image
-    # Pixel matrix has reduced dimensions (color channels reduced to 1, grayscale)
-
-    # Normalize the pixel_matrix
-    pixel_matrix = img[:, x - MATRIX_DX:x + MATRIX_DX, 0]
-    return pixel_matrix
-
-
 """
 Source: https://research-repository.griffith.edu.au/bitstream/handle/10072/15242/11185.pdf%3Bsequence=1
 For each segmentation point in a particular word (given by its xcoordinate),
@@ -130,9 +90,6 @@ window are presented.
 if __name__ == "__main__":
     if len(sys.argv) == 2:
         arg = sys.argv[1]
-        if arg == '-t' or arg == '--test':
-            read_training_data()
-        else:
-            start_training(int(arg))
+        start_training(int(arg))
     else:
         start_training(0)

### 2017-12-08 CharacterRecognition/character_recognition.py
remove bug

@@ -4,7 +4,7 @@ import tensorflow as tf
 from sklearn.utils import shuffle
 from sklearn.model_selection import train_test_split
 from CharacterRecognition import graphics as gr
-import definitions
+import settings
 import cv2
 from CharacterRecognition import utils
 import CharacterRecognition.preprocessing as preprocess
@@ -20,8 +20,8 @@ KEEP_PROB = 0.5
 BATCH_SIZE = 128
 FILTER_SIZE = 5
 NUM_CLASSES = 62
-SIZE = definitions.SIZE
-NUM_CHANNELS = definitions.NUM_CHANNELS
+SIZE = settings.SIZE
+NUM_CHANNELS = settings.NUM_CHANNELS
 
 
 # Helper functions
@@ -30,7 +30,7 @@ def open_images():
     Opens the dataset and preprocesses the images.
     :return: The labels of images, numpy pixel arrays with the image data, amount of images
     """
-    file_names = open(definitions.PREPROCESSED_CHAR_DATA_TXT_PATH, 'r').read().splitlines()
+    file_names = open(settings.PREPROCESSED_CHAR_DATA_TXT_PATH, 'r').read().splitlines()
     print("Reading dataset of {} images".format(len(file_names)))
     file_names = shuffle(file_names)
     labels = [int(x.split('-')[1]) for x in file_names]
@@ -238,7 +238,7 @@ def save_session(session):
     :return: None 
     """
     saver = tf.train.Saver()
-    saver.save(session, definitions.SAVE_PATH)
+    saver.save(session, settings.SAVE_PATH)
     session.close()
 
 
@@ -247,7 +247,7 @@ def restore_session(session):
     Create a session initialized with the session as defined in the SAVE_PATH.
     :return: None
     """
-    tf.train.Saver().restore(session, definitions.SAVE_PATH)
+    tf.train.Saver().restore(session, settings.SAVE_PATH)
 
 
 def restore_train_save(epochs):
@@ -327,7 +327,7 @@ def examples():
     for ex in reversed(examples):
         for i in range(4):
             print(ex,
-                  img_to_text(cv2.imread(definitions.EXAMPLE_CHAR_PATH + ex + '_' + str(i) + ".png", 0),
+                  img_to_text(cv2.imread(settings.EXAMPLE_CHAR_PATH + ex + '_' + str(i) + ".png", 0),
                               sessionargs, n=1))
 
 train_net(200, restore=False)

### 2017-12-08 CharacterRecognition/experiment_nets.py

@@ -1,4 +1,4 @@
-import definitions
+import settings
 import tensorflow as tf
 from CharacterRecognition.character_recognition import new_conv_layer
 from CharacterRecognition.character_recognition import new_fc_layer
@@ -16,8 +16,8 @@ KEEP_PROB = 0.5
 BATCH_SIZE = 512
 FILTER_SIZE = 3
 NUM_CLASSES = 62
-SIZE = definitions.SIZE
-NUM_CHANNELS = definitions.NUM_CHANNELS
+SIZE = settings.SIZE
+NUM_CHANNELS = settings.NUM_CHANNELS
 
 f = 'fully connected layer'
 fx = 'fully connected layer with dropout'

### 2017-12-08 CharacterRecognition/experiment_results.py

@@ -1,4 +1,4 @@
-import definitions
+import settings
 import os
 import numpy as np
 import matplotlib.pyplot as plt
@@ -50,7 +50,7 @@ def get_configuration_results(experiment_name, conf):
 
 
 def find_experiment_location(experiment_name):
-    return [(x[0].split('/')[-1], x[0]) for x in os.walk(definitions.EXPERIMENTS_CHAR_PATH) if experiment_name in x[0]]
+    return [(x[0].split('/')[-1], x[0]) for x in os.walk(settings.EXPERIMENTS_CHAR_PATH) if experiment_name in x[0]]
 
 
 def find_experiment_conf_location(experiment_name, conf):
@@ -76,7 +76,7 @@ def visualise(accuracies, times, name='', show=True, save=False, labels=None):
     axes = plt.gca()
     axes.set_ylim([0, 0.9])
     if save and name != '':
-        savefig(os.path.join(definitions.EXPERIMENTS_CHAR_GRAPHS_PATH, name + '.png'))
+        savefig(os.path.join(settings.EXPERIMENTS_CHAR_GRAPHS_PATH, name + '.png'))
     if show:
         plt.show()
 

### 2017-12-08 CharacterRecognition/experiments.py

@@ -1,4 +1,4 @@
-import definitions
+import settings
 from numpy import savetxt
 from CharacterRecognition import character_recognition
 from CharacterRecognition import preprocessing
@@ -168,7 +168,7 @@ def save_output(name, accuracies, time, iteration=None):
     extension = '.txt'
     if iteration is not None:
         extension = '_' + str(iteration) + extension
-    out_path = definitions.EXPERIMENTS_CHAR_PATH + name + '/'
+    out_path = settings.EXPERIMENTS_CHAR_PATH + name + '/'
     try:
         os.makedirs(out_path)
     except OSError:

### 2017-12-08 CharacterRecognition/preprocessing.py

@@ -3,7 +3,7 @@ import time as t
 import cv2
 import numpy as np
 
-import definitions
+import settings
 
 '''
 For this project we use the Chars74K dataset. It contains 64 classes, with about 74K images.
@@ -11,7 +11,7 @@ For this project we use the Chars74K dataset. It contains 64 classes, with about
 
 # Global constants
 start = t.time()
-SIZE = definitions.SIZE
+SIZE = settings.SIZE
 EMPTY_VALUE = 0.0
 WRITE_VALUE = 1.0
 
@@ -62,22 +62,22 @@ def noisyImage(img, stddev):
 # Helper functions
 def rotateImage(image, angle):
     rotation_matrix = cv2.getRotationMatrix2D(center=(SIZE / 2, SIZE / 2), angle=angle, scale=1.0)
-    return cv2.warpAffine(image, rotation_matrix, definitions.SHAPE)
+    return cv2.warpAffine(image, rotation_matrix, settings.SHAPE)
 
 
 def translateImage(image, tx, ty):
     translation_matrix = np.float32([[1, 0, tx], [0, 1, ty]])
-    return cv2.warpAffine(image, translation_matrix, definitions.SHAPE)
+    return cv2.warpAffine(image, translation_matrix, settings.SHAPE)
 
 
 def scaleImage(image, sx, sy):
     scale_matrix = np.float32([[sx, 0, 0], [0, sy, 0]])
-    return cv2.warpAffine(image, scale_matrix, definitions.SHAPE)
+    return cv2.warpAffine(image, scale_matrix, settings.SHAPE)
 
 
 def shearImage(image, s):
     shear_matrix = np.float32([[1, s, 0], [0, 1, 0]])
-    return cv2.warpAffine(image, shear_matrix, definitions.SHAPE)
+    return cv2.warpAffine(image, shear_matrix, settings.SHAPE)
 
 def erodeImage(image):
     element = cv2.getStructuringElement(cv2.MORPH_RECT,(3,3))
@@ -144,26 +144,26 @@ def read_image(file_name):
 
 
 def preprocess_image(img, inverse=False):
-    if img.shape != definitions.SHAPE:
-        img = cv2.resize(src=img, dsize=definitions.SHAPE)
+    if img.shape != settings.SHAPE:
+        img = cv2.resize(src=img, dsize=settings.SHAPE)
     # Contrast normalisation and inverting color values
     # We add a simple threshold for distinguishing the foreground and background.
[…]

### 2017-12-08 Postprocessing/lib/demo.py

@@ -1,0 +1,5 @@
+#!/usr/bin/env python
+from __future__ import print_function
+import phrasefinder
+
+

### 2017-12-08 Postprocessing/lib/makedocs.sh

@@ -1,0 +1,3 @@
+#!/usr/bin/env bash
+
+pdoc --html --html-dir docs --overwrite src/phrasefinder.py

### 2017-12-08 Postprocessing/lib/phrasefinder.py

@@ -1,0 +1,153 @@
+# This file is part of PhraseFinder.  http://phrasefinder.io
+#
+# Copyright (C) 2016  Martin Trenkmann
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#    http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""
+Module phrasefinder provides routines for querying the PhraseFinder web service at
+http://phrasefinder.io.
+"""
+import sys
+if sys.version_info[0] < 3:
+    # Python 2.
+    import urllib as urllibx
+    from urllib import urlencode as urlencode
+else:
+    # Python 3.
+    import urllib.request as urllibx
+    from urllib.parse import urlencode as urlencode
+
+VERSION_MAJOR = 0
+VERSION_MINOR = 1
+VERSION_MICRO = 1
+VERSION = VERSION_MAJOR * 1000000 + VERSION_MINOR * 1000 + VERSION_MICRO
+"""Defines the version number as one integer."""
+
+class Corpus(object):
+    """Corpus contains numeric constants that represent corpora to be searched.
+
+    All corpora belong to version 2 of the Google Books Ngram Dataset
+    (http://storage.googleapis.com/books/ngrams/books/datasetsv2.html).
+    """
+    EnglishUS, EnglishGB, Spanish, French, German, Russian, Chinese = range(7)
+
+class Status(object):
+    """Status contains numeric constants that report whether a request was successful.
+
+    The value is derived from the HTTP status code sent along with a response. Note that the numeric
+    value does not correspond to the original HTTP code.
+    """
+    Ok, BadRequest, PaymentRequired, MethodNotAllowed, TooManyRequests, ServerError = range(6)
+
+class Token(object):
+    """Token represents a single token (word, punctuation mark, etc.) as part of a phrase."""
+    class Tag(object):
+        """Tag denotes the role of a token with respect to the query."""
+        Given, Inserted, Alternative, Completed = range(4)
+    def __init__(self):
+        self.text = ""
+        self.tag  = Token.Tag.Given
+
[…]

### 2017-12-08 Tokenization/segmentation_correction.py

@@ -1,4 +1,4 @@
-import definitions
+import settings
 import cv2
 import numpy as np
 import Tokenization.character_extraction_main as chrextr
@@ -13,7 +13,7 @@ START_LINE = 18
 
 def open_images(start=0, amount=1000):
     file_entries = [x.split(' ') for x in
-                    open(definitions.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[
+                    open(settings.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[
                     start + START_LINE:START_LINE + start + amount]]
     print("Reading new dataset of {} images, starting at image {}".format(len(file_entries), start))
     images = []
@@ -21,13 +21,13 @@ def open_images(start=0, amount=1000):
     for file_name in files:
         parts = file_name.split('-')
         file_path = '/'.join([parts[0], parts[0] + '-' + parts[1], file_name + '.png'])
-        images.append((file_path, cv2.imread(definitions.CHAR_SEGMENTATION_DATA_PATH + file_path)))
+        images.append((file_path, cv2.imread(settings.CHAR_SEGMENTATION_DATA_PATH + file_path)))
     return images
 
 
 def find_start(file_name):
     file_entries = [x.split(' ') for x in
-                    open(definitions.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[START_LINE:]]
+                    open(settings.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[START_LINE:]]
     for (i, entry) in enumerate(file_entries):
         if entry[0] + '.png' == file_name:
             return i
@@ -38,7 +38,7 @@ def create_training_data(start=0, amount=1000):
     show_range = 1
     images = open_images(start, amount)
     n = len(images)
-    with open(definitions.CHAR_SEGMENTATION_TRAINING_OUT_PATH, "ab") as out_file:
+    with open(settings.CHAR_SEGMENTATION_TRAINING_TXT_PATH, "ab") as out_file:
         for (i, (img_path, img)) in enumerate(images):
             print("Showing image number {} of {}".format(i, n))
             split_data = []
@@ -54,7 +54,6 @@ def create_training_data(start=0, amount=1000):
                     split = True
                     # Create pixel matrix around splitpoint
                 elif key == 27:  # escape
-                    pickle.dump(i, out_file)  # Write length to pickle
                     return
                 else:  # user decides x is not a splitpoint
[…]

### 2017-12-08 create_data.py

@@ -1,16 +1,19 @@
-import definitions
+import settings
 import cv2
+import numpy as np
 import Tokenization.character_extraction_main as chrextr
 from pathlib import Path
 import sys
 import pickle
 
+WINDOW_SIZE = 5
+MATRIX_DX = 5
 START_LINE = 18
 
 
 def open_images(start=0, amount=1000):
     file_entries = [x.split(' ') for x in
-                    open(definitions.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[
+                    open(settings.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[
                     start + START_LINE:START_LINE + start + amount]]
     print("Reading new dataset of {} images, starting at image {}".format(len(file_entries), start))
     images = []
@@ -18,13 +21,13 @@ def open_images(start=0, amount=1000):
     for file_name in files:
         parts = file_name.split('-')
         file_path = '/'.join([parts[0], parts[0] + '-' + parts[1], file_name + '.png'])
-        images.append((file_path, cv2.imread(definitions.CHAR_SEGMENTATION_DATA_PATH + file_path)))
+        images.append((file_path, cv2.imread(settings.CHAR_SEGMENTATION_DATA_PATH + file_path)))
     return images
 
 
 def find_start(file_name):
     file_entries = [x.split(' ') for x in
-                    open(definitions.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[START_LINE:]]
+                    open(settings.CHAR_SEGMENTATION_DATA_TXT_PATH, 'r').read().splitlines()[START_LINE:]]
     for (i, entry) in enumerate(file_entries):
         if entry[0] + '.png' == file_name:
             return i
@@ -35,7 +38,7 @@ def create_training_data(start=0, amount=1000):
     show_range = 1
     images = open_images(start, amount)
     n = len(images)
-    with open(definitions.CHAR_SEGMENTATION_TRAINING_OUT_PATH, "ab") as out_file:
+    with open(settings.CHAR_SEGMENTATION_TRAINING_TXT_PATH, "ab") as out_file:
         for (i, (img_path, img)) in enumerate(images):
             print("Showing image number {} of {}".format(i, n))
             split_data = []
@@ -51,7 +54,6 @@ def create_training_data(start=0, amount=1000):
                     split = True
                     # Create pixel matrix around splitpoint
                 elif key == 27:  # escape
[…]

### 2017-12-08 settings.py

@@ -30,4 +30,5 @@ EXPERIMENTS_CHAR_GRAPHS_PATH = CHAR_PATH + 'Graphs/'
 
 CHAR_SEGMENTATION_DATA_TXT_PATH = TXT_PATH + 'words.txt'
 CHAR_SEGMENTATION_DATA_PATH = DATA_PATH + 'words/'
-CHAR_SEGMENTATION_TRAINING_OUT_PATH = DATA_PATH + 'segmentation/segmentation.txt'
+CHAR_SEGMENTATION_TRAINING_PATH = DATA_PATH + 'segmentation/'
+CHAR_SEGMENTATION_TRAINING_TXT_PATH = CHAR_SEGMENTATION_TRAINING_PATH + 'segmentation.txt'

### 2017-12-08 Postprocessing/lib/demo.py
fix bug

@@ -1,5 +1,0 @@
-#!/usr/bin/env python
-from __future__ import print_function
-import phrasefinder
-
-

### 2017-12-08 Postprocessing/lib/makedocs.sh

@@ -1,3 +1,0 @@
-#!/usr/bin/env bash
-
-pdoc --html --html-dir docs --overwrite src/phrasefinder.py

### 2017-12-08 Tokenization/character_extraction_main.py
fixed exception

@@ -48,6 +48,9 @@ def extract_character_separations(word_image):
     but will yield the smallest percentage of undersegmentation.
     We start with taking the segmentation of the image without rotation as to ensure no unnecessary rotations are picked
     '''
+    print(rotated_splits)
+    if len(rotated_splits) == 0 or len(rotated_splits[0]) == 0:
+        return []
     most_splits = len( rotated_splits[ len(rotated_splits) // 2 ][0] )
     chosen_split_layout = rotated_splits[ len(rotated_splits) // 2 ]
     for rotated_split in rotated_splits:

### 2017-12-10 src/character_recognition.py
boop

@@ -32,7 +32,7 @@ def open_images():
     :return: The labels of images, numpy pixel arrays with the image data, amount of images
     """
     file_names = open(definitions.PREPROCESSED_CHARSET_INFO_PATH, 'r').read().splitlines()
-    print("Reading dataset of {} images".format(len(file_names)))
+    #print("Reading dataset of {} images".format(len(file_names)))
     file_names = shuffle(file_names)
     labels = [int(x.split('-')[1]) for x in file_names]
     length = len(file_names)
@@ -154,7 +154,7 @@ def create_training_operation(h, _y, learning_rate=LEARNING_RATE, decay=DECAY, g
     return training_operation
 
 
-def train_net(n, restore=False, min_save=1.0):
+def train_net(n, restore=False, min_save=1.0, iteration=1):
     """
     Trains the network for n epochs with a new session
     Note: if this function has never been run, set restore to false!
@@ -192,13 +192,13 @@ def train_net(n, restore=False, min_save=1.0):
         accuracies.append(validation_accuracy)
         t = time() - start
         times.append(t)
-        if i % 10 == 0:
-            print('EPOCH {} - {:.0f}: Validation Accuracy = {:.3f}'.format(i, t, validation_accuracy))
+        #if i % 10 == 0:
+            #print('EPOCH {} - {:.0f}: Validation Accuracy = {:.3f}'.format(i, t, validation_accuracy))
         if validation_accuracy > min_save:
             print("New maximum accuracy {} achieved.".format(validation_accuracy))
             save_session(session)
             min_save = validation_accuracy
-    save_output("all", time=times, accuracies=accuracies, iteration=1)
+    save_output("all", time=times, accuracies=accuracies, iteration=iteration)
     t = str(time() - start)
     print("The training took: " + str(t) + " seconds.")
     return session

### 2017-12-10 src/utils/character_combinator.py
mhm

@@ -9,7 +9,7 @@ def normalize_and_execute_character_recognition(character_images):
 # All values are in pixels
 SPLIT_POINT_MIN_DISTANCE = 5
 MINIMAL_PIXEL_COUNT = 100
-split_count_per_x_range = 1/8 (1 split point per 8 pixels) 
+split_count_per_x_range = 1/8 (1 split point per 8 pixels)
 def evaluate_character_combinations(character_images, session_args):
 
     # Prepare characters for evaluation
@@ -24,7 +24,14 @@ def evaluate_character_combinations(character_images, session_args):
 
     """
 
+    """
+    count pixels between splits, order by most probable unnessecary
+        => fitness
 
-    for iteration in range(1, 8):
-        identiefier = 0
-        nlist.append(normalizer.normalize_character([char_img]))
+    close to two other split splitpoints
+        => fitness minus'
+
+    Check if splits are closer thatn MIN_DISTANCE
+        remove which one?
+
+    """

### 2017-12-10 Report/paper.tex
remove bug

@@
 \begin{abstract} 
+This paper handles off-line handwriting recognition. We created a system which converts images with handwriting to text. First we explain how to recognise characters.
+This explanation involves the used dataset, preprocessing techniques and the deep neural network for classifying characters.
+Next we describe how to segment a text into words using a contouring algorithm with consequent postprocessing to remove unwanted noise.
+After this we present our approach to seperating the characters from words using vertical projections combined with postprocessing techniques including a neural network for splitpoint evaluation.
+An explanation of two postprocessing techniques follow.
+We try to correct errors word recognition using an English dictionary, and language models, n-gram models in particular.
+Finally we discuss our results and some problems we encountered.
 
@@
 \section{Introduction}
-
 \section{Handwriting Recognition}
@@
 We normalize the data by reducing the size from 1200x900x3 to 32x32x1 (width x height x number of color channels).
+A simple threshold is added to improve the contrast in our image.
+This allows the foreground to be more distinguishable from the background and allows the neural network to train and recognise more easily.
 
-%TODO BINARIZATION -> INCREASE CONTRAST BETWEEN BACKGROUND AND FOREGROUND
-%This might solve our problem that handwriting isn't recognised.
-
-%TODO INVERTING -> We still need to check influence of inverting color values. 
-%This surely helps for augmenting the data.
-
-% Size / Contrast
 \paragraph{Augmenting data}
@@
 \end{equation}
+\subparagraph{Erosion of data}
+The erosion operation of images adds thinner versions of characters to the dataset.
+Erosion uses a kernel B with some shape to convolute an image.
+The image is scanned using the kernel and, the maximum pixel value of overlapping part between the image and the anchor point of the kernel is replaced with this pixel value.
+This causes bright regions to get thinner, and dark regions to get bigger in the image.
+All of the previously discussed preprocessing techniques are also applied to the eroded image.
+The final dataset contains translated, orientated, sheared, scaled and noisy version of both the original image, and the eroded image.
 \subsubsection{Deep neural network}
@@
 The perceptrons take an input, and calculate their output by calculating the activation function with the weights, input and bias.
+These perceptrons make simple decisions using the activation function.
 Images can be fed to the neural network, which then propagate trough several layers in the neural network and eventually produce as output the probabiliy that the input has a certain class, one of the possible characters.
 
-% I will add a discussion of the neural network once the neural network is finalized and not subject tot change. 
+Now follows an overview of the final neural network.
+The input is a normalized image with one color channel.
+This is fed to a convolutational layer.
+Convolutional layers allow us extract features by scanning the image with filters.
+These features might be edges, corners, or any other type of shape in the image.
+The extracted features using such a filter are called a feature map.
+We can apply several of those filters using one convolutional layer to extract several types of features.
+This works because images have spatial information, close pixels have shared information.
+By applying a filter, pixels in the same region will use shared biases and weights, which allows the filter to extract particular features.
+Next we add a pooling layer, in particular a max-pooling layer.
+Pooling layers remove the positional information of the features in the image.
+This is useful because the exact location is less important than the location relative to the other features, and this allows us to reduce the amount of parameters required in the next layers.
+
+We use 3 convolutional layers combined with max pooling layers in order to extract low, medium, and high level features.
+Next we flatten the output of the last convolutional layer in order to extract a 1-dimensional array from the 3-dimensional feature maps.
+This 1-dimensional array can now be fed to a fully connected layer, which is a layer of connected perceptrons as previously discussed.
+There's 2 of these fully connected layers each connected with a dropout layer.
+The purpose of this layer is to remove activations at random.
+Dropout greatly reduces overfitting in the fully connected layers.
+We don't use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
+Next there's one more fully connected layer which has 62 output neurons, which correspond to the 62 possible output classes.
+We use rectified linear units in order to compute the activation function.
+This choise is based on experimental results as there aren't any theories yet describing how to choose the ideal activation function.
+The last layer in the neural network is a softmax layer with a cross-entropy cost function.
+The softmax layer compute a probability distribution (a set of 62 numbers which sum up to 1) using the 62 output neurons from the previous layer.
+An element in the probability distribution estimates the chance of the input image to classify as a certain character.
+The cross-entropy cost function uses the actual labels to calculate the error from the predicted labels (the previously discussed probabilities).
+The closer the cross-entropy function gets to 0, the better it is at classifying handwritten characters.
+
+Finally, in order to actually train the neural network, we need some kind of weight adjustment.
+This will effectively improve the decision making of perceptrons and thus, the classification of images.
+The Adam optimizer effectively fulfills this goal.
+In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow.
+First of all, stochastic gradient descent (SGD) is an iterative method which tries to find minima.
+This function takes a step size, which is known as the learning rate.
+SGD can be used to minimize the cross-entropy function iteratively by adjusting the weights.
+Momentum-based gradient descent makes the correction dependent based on an average of several previous corrections to move more quickly in the correct direction.
+AdaGrad allows the learning rate to adapt based on parameters.
+AdaDelta solves the problem of AdaGrad, concerning the learning rate to become so small that learning becomes impossible.
+Each of these methods effectively improve the previously mentionned method.
+The Adam optimizer is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
+This optimizer works best in our case and will be used.
+
 \subsection{Tokenization}
+\subsubsection{Correction of oversegmentation}
+As our algorithm for segmentation of handwritten splits the word into to many characters, we can verify the correctness of these splitting points.
+\cite{evalsplitpoints} discusses 3 approaches to do evaluate split point correctness: rule based approaches, cost function optimization based approaches and machine learning based approaches.
+In this paper we will only discuss the latter, machine learning based approaches, as we used this approach.
+In particular, our algorithm is based on \cite{evalsplitpointsnn}.
+The paper explains how to use neural networks the correctness of splitting points.
+The discussed algorithm is as follows.
+Manual classification of splitting points creates two classes, correct and incorrect splitting points.
+For every of these splitting points a pixel matrix is extracted surrounding the splitting point.
+This pixel matrix is normalised in size.
+Next, small windows of equal size are extracted from this pixel matrix and for each windows the density is calculated, as number of black pixels divided by amount of pixels.
+With every matrix a corresponding label is encoded for storage in the training file, this label is encoded as 0.9 if the splitting point is correct and 0.1 otherwise.
+Now these matrices and corresponding labels are used to train the neural network.
+
+We take a slightly different approach and consider a deep neural network to tackle this problem.
+This allows us to use convolutational layers in order to extract features, instead of manually creating features (features which show the density in this case).
+We use 2 convolutional layers, 2 fully connected layers and a dropout layer as neural network.
+This neural network takes the actual image of the splitting point as input (a pixel matrix of the pixels surrounding the splitting point) and returns a one-hot encoded vector indicating if the splitting is correct of incorrect.
+This allowed us to achieve better results than when trying to implement the algorithm exactly as described by the paper.
+Splitting points can now be filtered if the neural network decides a splitting point is incorrect.
 
@@
 That amount of occurences is the same as counting the amount of n-grams with that sequence of words.
-An n-gram is a contiguous sequence of n items from a given sequence of text or speech.
-\cite{ngram}
+An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+k}$
 We use the n-grams from the Google dataset for this purpose.
@@
 further discuss the results of these techniques and some experiments to check the performance and impact of optimizations. 
-\section{Experiments and results}
+\section{Experimental results}
 %TODO
@@
 When we look at the average however, the training time is the highest for this case.
-\paragraph{Filter amount}
-% Not sure if I should discuss this, pretty boring results. 
-\paragraph{Filter size}
-In this experiment we check the influence of choosing different sizes of filters for the convolutional network.
-The filter size determines the size of our filters for convolving the input images.
-In figure \ref{fig:filtersize} we can see the time increases greatly, because more surrounding values have to be considered.
-The accuracy increases slightly when using bigger filters.
-Using different filter sizes allow other type of features to be extracted.
-% TODO: this is hard to explain..., maybe just leave this one out.
-% Convolutation layers allow extracting features from the image. 
-% Go over this again: good explanation: (See: http://www.deeplearningbook.org/)
-% Pixels far away from eachother are less correlated. This is why we subsample filter responses.
 \paragraph{Learning rate}
@@
 Without decay we only reach an accuracy of 65\%.
-%TODO: higher decay -> higher accuracy
-\paragraph{Optimizer}
-% Only useful optimizer is Adam. I'm not sure why, or if we shoul compare these at all. 
 \paragraph{Dropout}
@@
 \centering
-    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/preprocess.png}
+    \includegraphics[width=\linewidth]{../Graphs/preprocess.png}
     \caption{Impact of various preprocessing techniques.}
@@
     \centering
-    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/filtersize.png}
-    \caption{Impact of various filter sizes.}
-    \label{fig:filtersize}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/learningrate.png}
+    \includegraphics[width=\linewidth]{../Graphs/learningrate.png}
     \caption{Impact of different learning rates.}
@@
     \centering
-    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/batchsize.png}
+    \includegraphics[width=\linewidth]{../Graphs/batchsize.png}
     \caption{Impact of different batch sizes.}
@@
     \centering
-    \includegraphics[width=\linewidth]{../CharacterRecognition/Graphs/decay.png}
+    \includegraphics[width=\linewidth]{../Graphs/decay.png}
     \caption{Impact of weight decay regularisation.}
@@
 
-\section{Future work}
+\section{Future research}
 

### 2017-12-10 Report/paper.tex
voorlopige versie paper

@@
 \usepackage{qtree}
-\usepackage{listings}
 \usepackage{titlesec}
@@
 \begin{abstract} 
-This paper handles off-line handwriting recognition. We created a system which converts images with handwriting to text. First we explain how to recognise characters.
-This explanation involves the used dataset, preprocessing techniques and the deep neural network for classifying characters.
+This paper handles off-line handwriting recognition. We created a system which converts images with handwriting to text. 
+First we explain how to recognise characters. This explanation involves the used dataset, preprocessing techniques and the deep neural network for classifying characters. 
 Next we describe how to segment a text into words using a contouring algorithm with consequent postprocessing to remove unwanted noise. 
 After this we present our approach to seperating the characters from words using vertical projections combined with postprocessing techniques including a neural network for splitpoint evaluation. 
-An explanation of two postprocessing techniques follow.
-We try to correct errors word recognition using an English dictionary, and language models, n-gram models in particular.
+We try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
 Finally we discuss our results and some problems we encountered.
@@
 The smallest distinguishable token in a text is a character. These handwritten characters are uniquely 
-written by each individual and there's even variation when written by the same individual.
-These variations include the inclination of characters, i.e., the slant, orientation and size.
-These variations can be removed by preprocessing the image.
-We however take another, more general approach.
-Instead of adding rules to minimize variation we made a deep neural net which can recognise any character with a certain probability.
+written by each individual and variations even occur when written by the same individual. These variations include the inclination of characters, i.e., the slant, orientation and size. 
+A way of dealing with these includes removing variations from the input image using preprocessing techniques. 
+We take another approach, instead of adding rules to minimize variations we made a deep neural net which can recognise any character with a certain probability.
 The architecture of this deep neural network is described in \ref{sec:dnn}.
-In order to make the deep neural network recognise characters in images with these variations, we augmented the dataset to include images which have these variations.
-We discuss this in \ref{sec:preproc}.
-The dataset of handwritten character is described in \ref{sec:data}.
+In order to make the deep neural network recognise characters in images with these variations, we augmented the dataset to include a bigger variety of images.
+We present these augmentation techniques in \ref{sec:preproc}.
+The dataset of handwritten characters is described in \ref{sec:data}.
 
@@
 \label{sec:data}
-A good data dataset with many examples, enough variation, and which is balanced between classes is very important to train the neural network.
-We use the Chars74K dataset for this purpose.
-It has 62 classes (0-9, A-Z, a-z) and 3410 handwritten images.
+A good data dataset with many examples, enough variation, and which is balanced between classes is very important to train the neural network effectively. 
+We use the Chars74K dataset for this purpose. It has 3410 images with handwritten characters, each belonging to one of 62 classes (0-9, A-Z, a-z). 
 This dataset contains both very easily recognisable images of characters but also characters which are even hard to recognise for humans, as shown in figure \ref{fig:char}. 
-3410 input images is quite low for a neural network to be effectively trained.
-To overcome this problem we use some data augmentation techniques discussed in \ref{par:aug}.
+3410 input images is quite low for a neural network to be effectively trained. To overcome this problem we employ some data augmentation techniques discussed in \ref{par:aug}.
 
 \begin{figure}
-\begin{subfigure}{.15\textwidth}
+\begin{subfigure}{.23\textwidth}
   \centering
@@
 \end{subfigure}
-\begin{subfigure}{.15\textwidth}
+\begin{subfigure}{.23\textwidth}
   \centering
@@
 \end{subfigure}
-\begin{subfigure}{0.15\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{images/good_char}
-  \caption{Picture of a clean 'z'.}
-\end{subfigure}
-\caption{Examples of written 'z' characters.}
+\caption{Examples of written 'z' characters, which are hard to recognise.}
 \label{fig:char}
@@
 \label{sec:preproc}
-We make use of 2 types of preprocessing steps: normalization of the data and augmentation of the data.
+We make use of two types of preprocessing steps: normalization and augmentation of the data.
 Normalization of data adjusts the original data and data augmentation takes the data and uses augmentation techniques to create variations of the original data.
@@
 In reality however, it's better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}.
-We normalize the data by reducing the size from 1200x900x3 to 32x32x1 (width x height x number of color channels).
-A simple threshold is added to improve the contrast in our image.
+We normalize the data by reducing the input dimensions from (1200,900,3) to (64,64,1). A simple threshold is added to improve the contrast in our image. 
 This allows the foreground to be more distinguishable from the background and allows the neural network to train and recognise more easily.
@@
 \label{par:aug}
-Data augmentation serves 2 purposes.
-It makes the dataset more robust to different kinds of handwriting as we add variations of the original data to the input data.
+Data augmentation serves two purposes. It makes the dataset more robust to different kinds of handwriting as we add variations of the original data to the input data. 
 The second purpose is to increase the size of our dataset.
-As our dataset only contains images of about 3400 images with handwritten characters, we need more data for the neural network to function to its fullest extend.
+As our dataset only contains 3410 images with handwritten characters, we need more data for the neural network to be able to function properly.
 Increasing the size of our dataset has great impact on the performance of the neural network as discussed in \ref{sec:expres}.
@@
 \subparagraph{Translations of data}
-The dataset is also extended with variations in padding of characters in images. For this purpose we multiply every coordinate of the original image with the follow transformation matrix: 
+The dataset is also extended with variations in padding of characters in images. For this purpose we multiply every coordinate of the original image with the following transformation matrix: 
 \begin{equation}
@@
 \subparagraph{Rotations of data}
-Rotations of the image are added to the dataset to make the network more robust to handwritings with different orientations.
-We rotate by -30 and 30 degrees.
-We use a builtin python function from the opencv2 library to create the transformation matrix.
-This matrix is equal to:
+Rotations of the image are added to the dataset to make the network more robust to handwritings with different orientations. We rotate by -30 and 30 degrees. This matrix is equal to:
 \begin{equation}
@@
 \end{equation}
-where $\alpha = cos(angle)$ and $\beta = sin(angle)$.
-Center is the coordinate in our image around which is rotated and angle describes the amount of degrees to rotate in the clockwise direction.
-Translated images outside of the 32x32 frame are discarded.
+where $\alpha = cos(angle)$ and $\beta = sin(angle)$. Center is the coordinate in our image around which is rotated and the angle describes the amount of degrees to rotate in the clockwise direction. 
 \subparagraph{Scaling of data}
 With the purpose of making the network more robust to different sizes of handwriting, we add different scalings of the data.
-We both scale in with and/or in height.
-The following transformation matrix scales the input image:
+We both scale in width and in height.
+The following transformation matrix allows us to scale the input image:
 \begin{equation}
@@
 \subparagraph{Erosion of data}
-The erosion operation of images adds thinner versions of characters to the dataset.
-Erosion uses a kernel B with some shape to convolute an image.
-The image is scanned using the kernel and, the maximum pixel value of overlapping part between the image and the anchor point of the kernel is replaced with this pixel value.
-This causes bright regions to get thinner, and dark regions to get bigger in the image.
+The erosion operation of images adds thinner versions of characters to the dataset. Erosion uses a kernel to convolute an image. 
+The image is scanned using the kernel.
+The maximum pixel value of the overlapping part between the image and the kernel is computed.
+This pixel value of the anchor point of the kernel is replaced with the maximum pixel value.
+This causes bright regions to get thinner, and dark regions to get bigger in the image. A visualisation of this is shown in \ref{fig:erosion}. 
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{images/erosion}
+  \caption{Visualisation of eroding an inverted image with a handwritten 'j'. The right picture underwent erosion \cite{erosion}.}
+\label{fig:erosion}
+\end{figure}
 All of the previously discussed preprocessing techniques are also applied to the eroded image.
-The final dataset contains translated, orientated, sheared, scaled and noisy version of both the original image, and the eroded image.
+The final dataset contains translated, orientated, sheared, scaled and noisy versions of both the original image, and the eroded image.
 \subsubsection{Deep neural network}
 \label{sec:dnn}
+
+\paragraph{Introduction}
 Humans have millions of connected neurons in the visual cortices which are capable of image processing.
-We, humans, can recognise most characters without any thought because of this complex neural network.
+We, humans, can recognise most characters without any thought because of this complex neural network inside our brain. 
 For computers, character recognition is a more difficult task.
@@
 These deep neural networks contain several layers of connected perceptrons, where each connection is defined by weight and each perceptron is defined by a bias.
-The perceptrons take an input, and calculate their output by calculating the activation function with the weights, input and bias.
-These perceptrons make simple decisions using the activation function.
-Images can be fed to the neural network, which then propagate trough several layers in the neural network and eventually produce as output the probabiliy that the input has a certain class, one of the possible characters.
+The perceptrons take an input and calculate an output using an activation function. This allows perceptrons to make decisions. 
+Normalized images can be fed to the neural network, which then propagate trough several layers and eventually produce as output the probabiliy that the input image has a certain class, one of the possible characters.
+Now follows a detailed overview of the actual deep neural network.
 
-Now follows an overview of the final neural network.
+\paragraph{Convolutional layers}
  The input is a normalized image with one color channel.
@@
 These features might be edges, corners, or any other type of shape in the image.
-The extracted features using such a filter are called a feature map.
+The extracted features are called a feature map.
  We can apply several of those filters using one convolutional layer to extract several types of features. 
@@
 This is useful because the exact location is less important than the location relative to the other features, and this allows us to reduce the amount of parameters required in the next layers.
-
 We use 3 convolutional layers combined with max pooling layers in order to extract low, medium, and high level features. 
+\paragraph{Fully-connected layers}
 Next we flatten the output of the last convolutional layer in order to extract a 1-dimensional array from the 3-dimensional feature maps.
 This 1-dimensional array can now be fed to a fully connected layer, which is a layer of connected perceptrons as previously discussed.
-There's 2 of these fully connected layers each connected with a dropout layer.
+There are 3 fully connected layers each connected with a dropout layer.
 The purpose of this layer is to remove activations at random.
@@
 We don't use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
-Next there's one more fully connected layer which has 62 output neurons, which correspond to the 62 possible output classes.
-We use rectified linear units in order to compute the activation function.
-This choise is based on experimental results as there aren't any theories yet describing how to choose the ideal activation function.
+The last fully connected layer has 62 output neurons, which correspond to the 62 possible output classes. We use rectified linear units in order to compute the activation function.
+\paragraph{Softmax layer}
  The last layer in the neural network is a softmax layer with a cross-entropy cost function.
@@
 An element in the probability distribution estimates the chance of the input image to classify as a certain character.
-The cross-entropy cost function uses the actual labels to calculate the error from the predicted labels (the previously discussed probabilities).
-The closer the cross-entropy function gets to 0, the better it is at classifying handwritten characters.
-
+ The cross-entropy cost function uses the actual labels to calculate the error between the actual labels and the predicted labels (the previously discussed probabilities).
+The closer the cross-entropy function is to zero, the better it is at classifying handwritten characters.
+ \paragraph{Weight adjustment}
 Finally, in order to actually train the neural network, we need some kind of weight adjustment.
 This will effectively improve the decision making of perceptrons and thus, the classification of images.
-The Adam optimizer effectively fulfills this goal.
-In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow.
+The Adam optimizer effectively fulfills this goal. In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow. 
 First of all, stochastic gradient descent (SGD) is an iterative method which tries to find minima.
@@
 Momentum-based gradient descent makes the correction dependent based on an average of several previous corrections to move more quickly in the correct direction. 
-AdaGrad allows the learning rate to adapt based on parameters.
-AdaDelta solves the problem of AdaGrad, concerning the learning rate to become so small that learning becomes impossible.
+AdaGrad allows the learning rate to adapt based on parameters. AdaDelta solves a problem of AdaGrad, concerning the learning rate to become so small that learning becomes impossible. 
 Each of these methods effectively improve the previously mentionned method. 
-The Adam optimizer is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
-This optimizer works best in our case and will be used.
+The Adam optimizer is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter. This optimizer works best in our case and will be used.
 
-\subsection{Tokenization}
+\subsection{Segmentation of text into words}
+\subsection{Segmentation of words into characters}
 \subsubsection{Correction of oversegmentation}
-As our algorithm for segmentation of handwritten splits the word into to many characters, we can verify the correctness of these splitting points.
-\cite{evalsplitpoints} discusses 3 approaches to do evaluate split point correctness: rule based approaches, cost function optimization based approaches and machine learning based approaches.
+As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these splitting points. 
+\cite{evalsplitpoints} discusses 3 approaches to evaluate split point correctness: rule based approaches, cost function optimization based approaches and machine learning based approaches. 
 In this paper we will only discuss the latter, machine learning based approaches, as we used this approach. 
 In particular, our algorithm is based on \cite{evalsplitpointsnn}.
-The paper explains how to use neural networks the correctness of splitting points.
+The paper explains how to use neural networks to classify the correctness of splitting points.
 The discussed algorithm is as follows.
-Manual classification of splitting points creates two classes, correct and incorrect splitting points.
-For every of these splitting points a pixel matrix is extracted surrounding the splitting point.
-This pixel matrix is normalised in size.
-Next, small windows of equal size are extracted from this pixel matrix and for each windows the density is calculated, as number of black pixels divided by amount of pixels.
+Manual classification of splitting points creates training data with two classes, correct and incorrect splitting points.
+For every of these splitting points a pixel matrix is extracted surrounding the splitting point. This pixel matrix is normalised in size. 
+Next, small windows of equal size are extracted from the pixel matrix and for each window the density is calculated.
+The is the number of black pixels divided by the total amount of pixels in the window.
 With every matrix a corresponding label is encoded for storage in the training file, this label is encoded as 0.9 if the splitting point is correct and 0.1 otherwise.
@@
 This allows us to use convolutational layers in order to extract features, instead of manually creating features (features which show the density in this case).
-We use 2 convolutional layers, 2 fully connected layers and a dropout layer as neural network.
-This neural network takes the actual image of the splitting point as input (a pixel matrix of the pixels surrounding the splitting point) and returns a one-hot encoded vector indicating if the splitting is correct of incorrect.
-This allowed us to achieve better results than when trying to implement the algorithm exactly as described by the paper.
+We use 1 convolutional layer and 3 fully connected layers in our neural network.
+This neural network takes the actual image of the splitting point as input (a pixel matrix of the pixels surrounding the splitting point) and returns a one-hot encoded vector indicating if the splitting is correct or incorrect.
+This allowed us to achieve better results than our implementation of the algorithm exactly as described by the paper achieved.
 Splitting points can now be filtered if the neural network decides a splitting point is incorrect.
@@
 \subsection{Natural language processing}
-To cope with some of the high error rates as discussed in section \ref{sec:expres}, there needs to be some kind of correction, in order to produce possibble written text even when a step of the larger algorithm fails.
-In order to do this we adapt popural natural language processing techniques as postprocessing steps.
-Natural language techniques can often be found in speech recognition, language generation and other systems, but this can be helpful here aswell to correct mistakes made in the previous steps of the algorithm.
+To cope with some of the high error rates as shown in section \ref{sec:expres}, there needs to be some kind of correction.
+This correction should allow the system to produce possibble written text even when a step of the larger system fails.
+In order to do this we adapt natural language processing techniques as postprocessing steps.
+Natural language techniques can often be found in speech recognition, language generation and other systems.
 As a first step we match words against a dictionary to find out if they make sense, this is discussed in section \ref{sec:voc}.
@@
 We make a tree, as shown in figure \ref{fig:wordtree}, where every node is a tuple with a word (sequence of characters) and a probability (likeliness of that word based on the probabilities of individual characters).
-The probability is calculated as a product of probability for each character.
+The probability is calculated as the product of probability for each character: $P(c_1, c_2,...,c_k) = \prod\limits_{i}{c_i}$ with $1 \leq i \leq k$.
 In order to keep the complexity low we only consider the 3 most probable characters.
@@
 There's a tree for every possible starting character.
-Now we have a list of words and their probabilities (every leaf).
-We can calculate the closest matches in the dictionary for each word.
-For this purpose we use the function \lstinline{get_close_matches} from the python library difflib.
-We use function to find close matches in the English dictionary and their corresponding score.
+Now we have a list of possible words and their probabilities (every leaf).
+The closest matches in the dictionary for each word are calculated using the function \textbf{get\_close\_matches} from the python library difflib.
+\textbf{get\_close\_matches} finds close matches in the English dictionary and their corresponding score.
 We multiply this score with our previously calculated probability.
@@
 If we only want to convert the image of one word into text, we can just return the word with the highest score.
+This finalises the logic for recognising handwritten words. Next we will discuss a postprocessing technique useful we the word is part of a bigger context.
 \subsubsection{Language model}
@@
 This is were language models are useful.
-We can check the previous words and calculate the likeness of a word to be the next word in the context.
+We can check the previous words and calculate the likeliness of a word to be the next word in the context.
 N-gram models are used to calculate probabilities given n-1 previous words.
 The probability is computed as $P(w_i | w_{i-n},...,w_{i-1})$.
-We use the Markov assumption here, we can approximate the probability of a word so the word only depends on the previous n-1 words.
-This probability is calculated as follows:
+We use the Markov assumption here: the probability of a word in the context can be approximated by the previous n-1 words. This probability is calculated as follows: 
 \begin{equation}
@@
 \end{equation}
-where $w_i$ is the word of which we want to calculate the probability and $count(w_i,...w_{i+k})$ counts the amount of occurences of the sequence of words $w_i,...,w_{i+k}$.
-That amount of occurences is the same as counting the amount of n-grams with that sequence of words.
-An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+k}$
+where $w_i$ is the word of which we want to calculate the probability and $count(w_i,...w_{i+k})$ indicates the amount of occurences of $w_i,...,w_{i+k}$.
+That amount is the same as counting the amount of n-grams with that sequence of words.
+An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+k}$.
 We use the n-grams from the Google dataset for this purpose.
@@
 Combining the results with the probabilities calculated in section \ref{sec:voc}, we can 
-calculate a new score.
-The word with the highest score is accepted as written word.
-By iterating over all of the words, we now have converted the image with written text into text.
-In the next section we will
-further discuss the results of these techniques and some experiments to check the performance and impact of optimizations. 
+calculate a new score. The word with the highest score is accepted as written word. 
+The concludes our full system of recognising handwritten texts.
+In the next section we will further discuss the results of these techniques and some experiments to check the performance and impact of optimizations.
 \section{Experimental results}
@@
 A problem which arrised when trying to test many diffeent configurations was the long time it takes to train a neural network.
-In order to keep these experiments from running several months, we added some small constraints.
+In order to limit the time these experiments run we added some small constraints.
 Next it's important to note that a lot of non-determinism is part of these neural networks.
 When running the neural network twice with the same configuration, it could have completely different results.
+Another note is that some experiments might not use the finalized setup of our neural network. These experiments do use a same setup to compare different configurations. 
 In order to have meaningful results which run in a reasonable amount of time we ran every experiment 4 times but limited the amount of epochs to 200.
@@
 When discussing these results we take the best result of the 4 training results.
-We use this result as we are only interested in the potential of the neural network.
-If we'd let those experiments run for more iterations, they would reach the same accuracy because of the noise added to the neural network.
-The different configurations and results will be discussed next.
+We use this result as we are only interested in the potential of the neural network, and not possible outliers which perform bad due to poor weight adjustments.  
 \subsubsection{Experiments}
@@
 This improves the performance but also increases the training time of our network significantly.
-One remarkable result is that the training time of preprocessing with all techniques enabled, takes less long than some other results.
+One remarkable result is that the training time of preprocessing with all techniques enabled, takes shorter time than some other results.
 When we look at the average however, the training time is the highest for this case.
@@
 In order to increase to speed at which the neural network is trained, we feed batches of input images to the neural network instead of one image at a time.
-The backpropagation in order to calculate the weight adjustment only happens after every batch is fed.
+Weight adjustment only occurs after feeding a complete batch to the network.
 In the figure we can see bigger batch sizes increase the execution speed.

### 2017-12-10 src/character_recognition.py
small bug in preprocessing fixed

@@ -117,7 +117,7 @@ def create_neural_net(global_weights=None, train=True, base=1, filter_size=FILTE
     :param base: experimental parameter, a higher base should produce better results. This should be a strict positive integer.
     :return: The input layer x, the output layer with the predicted values and a placeholder for the actual values.
     """
-    with tf.variable_scope(''):
+    with tf.variable_scope('CharacterRecognition'):
         _x = tf.placeholder(tf.float32, (None, SIZE, SIZE, NUM_CHANNELS))  # batch size - height - width - channels
         _y = tf.placeholder(tf.int64, (None, NUM_CLASSES))  # batch size - classes
         base2 = base * 1024
@@ -192,8 +192,8 @@ def train_net(n, restore=False, min_save=1.0, iteration=1):
         accuracies.append(validation_accuracy)
         t = time() - start
         times.append(t)
-        #if i % 10 == 0:
-            #print('EPOCH {} - {:.0f}: Validation Accuracy = {:.3f}'.format(i, t, validation_accuracy))
+        if i % 10 == 0:
+            print('EPOCH {} - {:.0f}s: Validation Accuracy = {:.3f}'.format(i, t, validation_accuracy))
         if validation_accuracy > min_save:
             print("New maximum accuracy {} achieved.".format(validation_accuracy))
             save_session(session)

### 2017-12-10 src/main.py

@@ -92,7 +92,7 @@ def recognise_text(file_name):
             max([(word, alpha * voc_words[word] + beta * prob) for word, prob in lang_words.items()],
                 key=lambda x: x[0])[0]
         text.append(most_likely_word)
-        print("Found word: " + most_likely_word)
+        print("Found word: " + most_likely_word[0])
     return ' '.join(text)
 
 
@@ -108,7 +108,7 @@ def main(argv):
             print(recognise_text(arg))
         elif option == '--train-rec' or option == '-tr':  # Train a character segmentation model for 'arg' epochs
             epochs = int(arg) if arg is not None else 500
-            cr.train_net(epochs, min_save=0.795)
+            cr.train_net(epochs, min_save=0.790)
         elif option == '--train-split' or option == '-ts':  # Train a character recognition model for 'arg' epochs
             epochs = int(arg) if arg is not None else 250
             sd.train_net(epochs, min_save=0.71)
@@ -130,21 +130,6 @@ def main(argv):
             -h, --help: Prints this output.
             '''
         )
-
-
-def train_several_models():
-    path = definitions.MODELS_PATH + 'CharacterRecognition/Model'
-    src_path = path + '/'
-    for i in range(3):
-        print("Iteration {}".format(i))
-        session = cr.train_net(500, min_save=0.795, iteration=i + 1)
-        session.close()
-        for dirpath, dirnames, filenames in os.walk(src_path):
-            for filename in filenames:
-                dst_path = path + str(i + 1) + '/'
-                shutil.copy(src_path + filename, dst_path)
-    cr.train_net(1500, min_save=0.795, iteration=4)
-
 
 if __name__ == "__main__":
     main(sys.argv[1:])

### 2017-12-10 src/character_combinator.py
normalization fix and last try for separation optimization

@@ -1,0 +1,35 @@
+import character_normalizer as normalizer
+import character_recognition as cr
+import character_utils
+
+
+def normalize_and_combine_characters(split_characters_images, sessionargs_char_recognition):
+    split_characters_list = list()
+
+    max_iterations = min(8, len(split_characters_images))
+    identifier = 1;
+
+    for character_image in split_characters_images:
+        base_percentage, max_percentage, identiefier_list = 0, 0, list()
+        split_characters_list.append( (character_image, base_percentage, identiefier_list, max_percentage) )
+
+    for iteration_length in range(1, max_iterations):
+        print("##########################")
+        print(iteration_length)
+        for startindex in range(len(split_characters_list) - 1 - iteration_length):
+            print("@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@")
+            print(startindex)
+
+            combined_character_splices = split_characters_images[startindex:startindex + iteration_length]
+            normalized_combined_character_splices = normalizer.normalize_character(combined_character_splices)
+
+            for char in normalized_combined_character_splices:
+                print ("CHARACTER")
+                probabilities = cr.img_to_prob(char, sessionargs_char_recognition)
+                #print(cr.most_probable_chars(probabilities, 5))
+                #print(sorted([(character_utils.cls2str(i),x) for (i,x) in enumerate(probabilities)], key=lambda x:-x[1])[:5])
+
+            identifier += 1;
+
+
+    return list()

### 2017-12-10 src/character_extraction.py

@@ -11,6 +11,7 @@ import numpy as np
 # Found a good skeltonize-algorithm in the scikit-library
 # Scikit-image dependency! (https://www.lfd.uci.edu/~gohlke/pythonlibs/#scikit-image)
 from skimage.morphology import skeletonize
+import character_split_evaluator as cse
 
 import splitpoint_decision as toc
 
@@ -72,11 +73,15 @@ def extract_characters(word_image, sessionargs=None):
 
     chosen_split_layout = extract_character_separations(word_image, sessionargs=sessionargs)
 
+
     blur = cv2.GaussianBlur(word_image,(1,1),0)
     ret3,threshold = cv2.threshold(blur,0,255,cv2.THRESH_BINARY+cv2.THRESH_OTSU)
     # chosen split layout variables
     finalsplits = chosen_split_layout
     height, width = threshold.shape[:2]
+
+
+    cse.evaluate_character_combinations(threshold, finalsplits)
 
     # Splitting the characters on the chosen split locations (finalsplits)
     splitcharacters = list()

### 2017-12-10 src/character_normalizer.py

@@ -1,0 +1,84 @@
+import cv2
+import numpy as np
+
+# This is necessary because test data is also centered in this rectangle
+CHARACTER_DIMS = 44
+CONTAINER_DIMS = 64
+side_padding = (CONTAINER_DIMS - CHARACTER_DIMS) // 2
+
+def delete_white_border(character_image):
+    # Invert image to make whites zero
+    inverted = cv2.bitwise_not(character_image)
+    height, width = inverted.shape[:2]
+
+    col_summation = cv2.reduce(inverted, 0, cv2.REDUCE_SUM, dtype=cv2.CV_32S);
+    row_summation = cv2.reduce(inverted, 1, cv2.REDUCE_SUM, dtype=cv2.CV_32S);
+
+    # Getting border sizes
+    left_bound, right_bound, up_bound, down_bound = 0, len(col_summation[0])-1, 0, len(row_summation)-1
+
+    # Left
+    while ( left_bound < len(col_summation[0])-1 and col_summation[0][left_bound] == 0 ):
+        left_bound += 1
+    # RIGHT
+    while ( right_bound > 0  and col_summation[0][right_bound] == 0 ):
+        right_bound -= 1
+    # Up
+    while ( up_bound < len(row_summation)-1  and row_summation[up_bound] == 0):
+        up_bound += 1
+    # Down
+    while (down_bound > 0  and row_summation[down_bound] == 0):
+            down_bound -= 1
+
+    if (left_bound > right_bound or up_bound > down_bound):
+        return character_image
+
+    borderless_image = character_image[up_bound:down_bound, left_bound:right_bound]
+    return(borderless_image)
+
+def normalize_character(charlist):
+    """
+    Precondition:: All character slices in the input have the same height! This is correct when the character slices come from the same word.
+    :return: a normalized image of the chosen character splits all pasted together
+    """
+
+
+    if len(charlist) == 0:
+        return None
+
+    # Horizontal concatenation of the character splices
+    character_img = charlist[0]
+    index = 0;
+    while (index + 1) < len(charlist):
+        character_img = cv2.hconcat(character_img, charlist[index + 1])
+
+
+    # Treshold the image
+    blur = cv2.GaussianBlur(character_img,(3,3),0)
+    ret3,thresh = cv2.threshold(blur,120,255,cv2.THRESH_BINARY)
+
+    # Delete white border around images
+    borderless_image = delete_white_border(thresh)
+
+    # Add padding
[…]

### 2017-12-10 src/character_split_evaluator.py

@@ -1,16 +1,10 @@
-import character_normalizer as normalizer
-#import CharacterRecognition.CharacterRecognition as cr
-
-
-def normalize_and_execute_character_recognition(character_images):
-
-    print("lel")
-
-# All values are in pixels
 SPLIT_POINT_MIN_DISTANCE = 5
 MINIMAL_PIXEL_COUNT = 100
-split_count_per_x_range = 1/8 (1 split point per 8 pixels)
-def evaluate_character_combinations(character_images, session_args):
+split_count_per_x_range = 1/8 #(1 split point per 8 pixels)
+def evaluate_character_combinations(word_image, splitpoints):
+
+    for xval, val in splitpoints:
+        print(xval)
 
     # Prepare characters for evaluation
 

### 2017-12-10 src/main.py

@@ -1,7 +1,7 @@
 import cv2
 import character_recognition as chr
 import splitpoint_decision as toc
-from character_extraction import extract_characters
+import  character_extraction as ce
 from language_model import n_gram_model
 from vocabulary import most_likely_words
 from word_extraction import preprocess_image
@@ -9,6 +9,9 @@ import splitpoint_decision as sd
 import character_recognition as cr
 from character_preprocessing import augment_data
 import word_normalizer as wn
+
+import character_combinator as ccb
+
 from pathlib import Path
 import shutil
 import definitions
@@ -56,9 +59,10 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     :param image: Image of a word
     :return: A list of pairs, the pairs consist of likely words and their probabilities.
     """
-    print("########################################################")
     normalized_word_image = wn.normalize_word(img)
-    char_imgs = extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction)
+    char_imgs = ce.extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction)
+
+    ccb.normalize_and_combine_characters(char_imgs, sessionargs_char_recognition)
 
     # Call character_combinator
 
@@ -148,4 +152,3 @@ def train_several_models():
 
 if __name__ == "__main__":
     main(sys.argv[1:])
-

### 2017-12-10 src/character_combinator.py
fixed image concatenation normalizer, finishing combinator

@@ -1,7 +1,7 @@
 import character_normalizer as normalizer
 import character_recognition as cr
 import character_utils
-
+import cv2
 
 def normalize_and_combine_characters(split_characters_images, sessionargs_char_recognition):
     split_characters_list = list()
@@ -13,23 +13,27 @@ def normalize_and_combine_characters(split_characters_images, sessionargs_char_r
         base_percentage, max_percentage, identiefier_list = 0, 0, list()
         split_characters_list.append( (character_image, base_percentage, identiefier_list, max_percentage) )
 
-    for iteration_length in range(1, max_iterations):
+    print (len (split_characters_list))
+    for iteration_length in range(1, max_iterations+1):
         print("##########################")
         print(iteration_length)
-        for startindex in range(len(split_characters_list) - 1 - iteration_length):
+        for startindex in range(0, len(split_characters_list) - iteration_length + 1):
             print("@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@")
             print(startindex)
 
             combined_character_splices = split_characters_images[startindex:startindex + iteration_length]
+
             normalized_combined_character_splices = normalizer.normalize_character(combined_character_splices)
+            print("normalized")
 
-            for char in normalized_combined_character_splices:
-                print ("CHARACTER")
-                probabilities = cr.img_to_prob(char, sessionargs_char_recognition)
-                #print(cr.most_probable_chars(probabilities, 5))
-                #print(sorted([(character_utils.cls2str(i),x) for (i,x) in enumerate(probabilities)], key=lambda x:-x[1])[:5])
 
-            identifier += 1;
 
+
+            print ("CHARACTER")
+            probabilities = cr.img_to_prob(normalized_combined_character_splices, sessionargs_char_recognition)
+            print(cr.most_probable_chars(probabilities, 5))
+
+            #cv2.imshow("c2", normalized_combined_character_splices)
+            #cv2.waitKey(0)
 
     return list()

### 2017-12-10 src/character_normalizer.py

@@ -50,8 +50,8 @@ def normalize_character(charlist):
     character_img = charlist[0]
     index = 0;
     while (index + 1) < len(charlist):
-        character_img = cv2.hconcat(character_img, charlist[index + 1])
-
+        character_img = np.concatenate((character_img, charlist[index + 1]), axis=1)
+        index += 1
 
     # Treshold the image
     blur = cv2.GaussianBlur(character_img,(3,3),0)
@@ -79,6 +79,5 @@ def normalize_character(charlist):
 
     container = cv2.bitwise_not(np.zeros((CONTAINER_DIMS,CONTAINER_DIMS), np.uint8))
     container[side_padding:side_padding+CHARACTER_DIMS, side_padding:side_padding+CHARACTER_DIMS] = final_character_image
-
 
     return (container)

### 2017-12-10 src/character_combinator.py
fixed commit

@@ -3,6 +3,9 @@ import character_recognition as cr
 import character_utils
 import cv2
 
+
+UPPER_ERROR_MARGIN = 0.85
+MAX_WIDTH_HEIGHT_RATIO = 3
 def normalize_and_combine_characters(split_characters_images, sessionargs_char_recognition):
     split_characters_list = list()
 
@@ -10,30 +13,23 @@ def normalize_and_combine_characters(split_characters_images, sessionargs_char_r
     identifier = 1;
 
     for character_image in split_characters_images:
-        base_percentage, max_percentage, identiefier_list = 0, 0, list()
-        split_characters_list.append( (character_image, base_percentage, identiefier_list, max_percentage) )
+        split_characters_list.append( (identifier, 0) )
+        identifier += 1
 
-    print (len (split_characters_list))
     for iteration_length in range(1, max_iterations+1):
-        print("##########################")
-        print(iteration_length)
         for startindex in range(0, len(split_characters_list) - iteration_length + 1):
-            print("@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@")
-            print(startindex)
 
             combined_character_splices = split_characters_images[startindex:startindex + iteration_length]
 
-            normalized_combined_character_splices = normalizer.normalize_character(combined_character_splices)
-            print("normalized")
+            combined_character = normalizer.combine_characters(combined_character_splices)
+            height, width = combined_character.shape[:2]
 
+            normalized_combined_character_splices = normalizer.normalize_character(combined_character)
 
-
-
-            print ("CHARACTER")
             probabilities = cr.img_to_prob(normalized_combined_character_splices, sessionargs_char_recognition)
             print(cr.most_probable_chars(probabilities, 5))
 
-            #cv2.imshow("c2", normalized_combined_character_splices)
-            #cv2.waitKey(0)
+            cv2.imshow("c2", normalized_combined_character_splices)
+            cv2.waitKey(0)
 
     return list()

### 2017-12-10 src/character_normalizer.py

@@ -36,23 +36,26 @@ def delete_white_border(character_image):
     borderless_image = character_image[up_bound:down_bound, left_bound:right_bound]
     return(borderless_image)
 
-def normalize_character(charlist):
+def combine_characters(characterlist):
+    if len(characterlist) == 0:
+        return None
+
+    # Horizontal concatenation of the character splices
+    character_img = characterlist[0]
+    index = 0;
+    while (index + 1) < len(characterlist):
+        character_img = np.concatenate((character_img, characterlist[index + 1]), axis=1)
+        index += 1
+    return character_img
+
+def normalize_character(character):
     """
     Precondition:: All character slices in the input have the same height! This is correct when the character slices come from the same word.
     :return: a normalized image of the chosen character splits all pasted together
     """
 
-
-    if len(charlist) == 0:
-        return None
-
     # Horizontal concatenation of the character splices
-    character_img = charlist[0]
-    index = 0;
-    while (index + 1) < len(charlist):
-        character_img = np.concatenate((character_img, charlist[index + 1]), axis=1)
-        index += 1
-
+    character_img = character
     # Treshold the image
     blur = cv2.GaussianBlur(character_img,(3,3),0)
     ret3,thresh = cv2.threshold(blur,120,255,cv2.THRESH_BINARY)

### 2017-12-10 src/main.py

@@ -11,6 +11,8 @@ import word_normalizer as wn
 
 import character_combinator as ccb
 
+import character_normalizer as cn
+
 from pathlib import Path
 import shutil
 import definitions
@@ -61,7 +63,8 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     normalized_word_image = wn.normalize_word(img)
     char_imgs = ce.extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction)
 
-    ccb.normalize_and_combine_characters(char_imgs, sessionargs_char_recognition)
+    # normalized_characters = [cm.normalize_character(character) for character in char_imgs]
+    # ccb.normalize_and_combine_characters(char_imgs, sessionargs_char_recognition)
 
     # Call character_combinator
 

### 2017-12-10 src/demo.py
some changes

@@ -8,12 +8,12 @@ from character_preprocessing import augmentImage
 from character_preprocessing import preprocess_image
 from character_preprocessing import erodeImage
 from character_utils import index2str
-import  numpy as np
-
+import numpy as np
 
 aug_demo = False
 char_rec_demo = True
 word_splitting_demo = False
+
 
 def start_demo():
     char_file = definitions.PROJECT_PATH + 'Data/charset/Img/Sample056/img056-003.png'
@@ -22,7 +22,6 @@ def start_demo():
     char_image = main.read_image(char_file)
     tocsessionargs = toc.init_session()
     chrsessionargs = cr.init_session()
-
 
     if aug_demo:
         input("Data augmentation demo")
@@ -33,7 +32,6 @@ def start_demo():
     if word_splitting_demo:
         input("Word splitting demo")
         demo_word_splitting(word_image, tocsessionargs)
-
 
 
 def resize_word(img):
@@ -63,6 +61,7 @@ def demo_augmentation(img):
     for augimg in images:
         cv2.imshow("Augmented image", cv2.resize(np.multiply(255, np.subtract(1, augimg)), dsize=(300, 300)))
         cv2.waitKey(0)
+
 
 if __name__ == "__main__":
     start_demo()

### 2017-12-10 src/main.py

@@ -1,3 +1,5 @@
+
+
 import cv2
 import splitpoint_decision as toc
 import  character_extraction as ce
@@ -8,9 +10,7 @@ import splitpoint_decision as sd
 import character_recognition as cr
 from character_preprocessing import augment_data
 import word_normalizer as wn
-
 import character_combinator as ccb
-
 import character_normalizer as cn
 
 from pathlib import Path

### 2017-12-10 src/demo.py
improve demo

@@ -7,17 +7,20 @@ import character_recognition as cr
 from character_preprocessing import augmentImage
 from character_preprocessing import preprocess_image
 from character_preprocessing import erodeImage
-from character_utils import index2str
 import numpy as np
+import word_normalizer as wn
+import vocabulary
 
 aug_demo = False
-char_rec_demo = True
-word_splitting_demo = False
+char_rec_demo = False
+word_splitting_demo = True
+word_rec_demo = True
 
 
 def start_demo():
     char_file = definitions.PROJECT_PATH + 'Data/charset/Img/Sample056/img056-003.png'
-    word_file = definitions.PROJECT_PATH + 'Data/words/d01/d01-104/d01-104-00-02.png'
+    # word_file = definitions.PROJECT_PATH + 'Data/words/b04/b04-060/b04-060-00-07.png' # Government
+    word_file = 'words/beer.png'
     word_image = main.read_image(word_file)
     char_image = main.read_image(char_file)
     tocsessionargs = toc.init_session()
@@ -32,6 +35,9 @@ def start_demo():
     if word_splitting_demo:
         input("Word splitting demo")
         demo_word_splitting(word_image, tocsessionargs)
+    if word_rec_demo:
+        input("Word recognition demo")
+        demo_word_recognition(word_image, chrsessionargs, tocsessionargs)
 
 
 def resize_word(img):
@@ -46,6 +52,7 @@ def demo_word_splitting(word_image, sessionargs):
     splitpoints = [split for (is_split, split) in zip(splitpoint_decisions, splitpoints) if is_split]
     cv2.imshow('Splitpoints after neural network', resize_word(toc.show_splitpoints(word_image, splitpoints)))
     cv2.waitKey(0)
+    cv2.destroyAllWindows()
 
 
 def demo_augmentation(img):
@@ -61,6 +68,15 @@ def demo_augmentation(img):
     for augimg in images:
         cv2.imshow("Augmented image", cv2.resize(np.multiply(255, np.subtract(1, augimg)), dsize=(300, 300)))
         cv2.waitKey(0)
+
+
+def demo_word_recognition(img, chrsessionargs, tocsessionargs):
+    normalized_word_image = wn.normalize_word(img)
+    char_imgs = chrext.extract_characters(normalized_word_image, sessionargs=tocsessionargs, postprocess=False)
+    char_probabilities = cr.imgs_to_text(char_imgs, chrsessionargs, n=3)
[…]

### 2017-12-10 src/main.py

@@ -47,7 +47,7 @@ def recognise_word(file_name):
                key=lambda x: x[1])
 
 
-def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_oversegmentation_correction):
+def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_oversegmentation_correction, postprocess=True, verbose=False):
     """
     Algorithm:
 
@@ -61,7 +61,7 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     :return: A list of pairs, the pairs consist of likely words and their probabilities.
     """
     normalized_word_image = wn.normalize_word(img)
-    char_imgs = ce.extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction)
+    char_imgs = ce.extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction, postprocess=postprocess)
 
     # normalized_characters = [cm.normalize_character(character) for character in char_imgs]
     # ccb.normalize_and_combine_characters(char_imgs, sessionargs_char_recognition)
@@ -69,7 +69,7 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     # Call character_combinator
 
     # evaluated_chars = evaluate_character_combinations(char_imgs, sessionargs_char_recognition)
-    char_probabilities = cr.imgs_to_text(char_imgs, sessionargs_char_recognition, n=3)
+    char_probabilities = cr.imgs_to_text(char_imgs, sessionargs_char_recognition, n=3, verbose=verbose)
     return most_likely_words(char_probabilities)
 
 

### 2017-12-10 src/main.py
try to fix bug where edges are not taken as contours, pushing demo data

@@ -9,8 +9,6 @@ import character_recognition as cr
 from character_preprocessing import augment_data
 import word_normalizer as wn
 
-import character_combinator as ccb
-
 import character_normalizer as cn
 
 from pathlib import Path
@@ -66,9 +64,6 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     # normalized_characters = [cm.normalize_character(character) for character in char_imgs]
     # ccb.normalize_and_combine_characters(char_imgs, sessionargs_char_recognition)
 
-    # Call character_combinator
-
-    # evaluated_chars = evaluate_character_combinations(char_imgs, sessionargs_char_recognition)
     cls_pred_list = cr.imgs_to_prob_list(char_imgs, sessionargs_char_recognition)
     return most_likely_words(cls_pred_list)
 

### 2017-12-10 src/demo.py
renames and improve demo

@@ -10,12 +10,13 @@ from character_preprocessing import erodeImage
 import numpy as np
 import word_normalizer as wn
 import vocabulary
+import splitpoint_decision as sd
 
-aug_demo = False
-char_rec_demo = False
+aug_demo = True
+char_rec_demo = True
 word_splitting_demo = True
 word_rec_demo = True
-
+data_creation_demo = True
 
 def start_demo():
     char_file = definitions.PROJECT_PATH + 'Data/charset/Img/Sample056/img056-003.png'
@@ -27,21 +28,29 @@ def start_demo():
     chrsessionargs = cr.init_session()
 
     if aug_demo:
-        input("Data augmentation demo")
+        # input("Data augmentation demo")
         demo_augmentation(char_image)
     if char_rec_demo:
-        input("Character recognition demo")
-        print(cr.img_to_text(char_image, chrsessionargs, n=3))
+        # input("Character recognition demo")
+        demo_char_recognition(char_image, chrsessionargs)
+    if data_creation_demo:
+        # input("Data creation demo")
+        demo_data_creation()
     if word_splitting_demo:
-        input("Word splitting demo")
+        # input("Word splitting demo")
         demo_word_splitting(word_image, tocsessionargs)
     if word_rec_demo:
-        input("Word recognition demo")
+        # input("Word recognition demo")
         demo_word_recognition(word_image, chrsessionargs, tocsessionargs)
 
 
 def resize_word(img):
     return cv2.resize(img, dsize=(800, 300))
+
+
+def demo_data_creation():
+    sd.start_data_creation()
+    cv2.destroyAllWindows()
 
 
 def demo_word_splitting(word_image, sessionargs):
@@ -68,15 +77,26 @@ def demo_augmentation(img):
     for augimg in images:
         cv2.imshow("Augmented image", cv2.resize(np.multiply(255, np.subtract(1, augimg)), dsize=(300, 300)))
         cv2.waitKey(0)
+    cv2.destroyAllWindows()
 
+
+def demo_char_recognition(img, sessionargs):
+    cv2.imshow("Character recognition", img)
+    cv2.waitKey(0)
+    print(cr.img_to_text(img, sessionargs, n=3))
+    cv2.destroyAllWindows()
 
 def demo_word_recognition(img, chrsessionargs, tocsessionargs):
+    cv2.imshow("Word recognition", img)
+    cv2.waitKey(0)
     normalized_word_image = wn.normalize_word(img)
[…]

### 2017-12-10 src/splitpoint_decision.py

@@ -39,7 +39,7 @@ def open_images(start=0, amount=-1):
     return images
 
 
-def create_training_data(start=0, amount=1000):
+def create_training_data(start=0, amount=10):
     images = open_images(start, amount)
     n = len(images)
     with open(definitions.WORD_SPLITTING_PATH, "ab") as out_file:
@@ -49,14 +49,13 @@ def create_training_data(start=0, amount=1000):
             split_data = manual_split_point_detection(img, splits)
             if split_data is None:
                 return
-            pickle.dump((start + i, img_path, split_data), out_file)
+            pickle.dump((start + i, img_path, split_data), out_file, protocol=2)
 
 
 def show_splitpoints(img, splits):
     if len(img.shape) == 2:
         img = np.expand_dims(img, axis=2)
         img = np.repeat(img, 3, axis=2)
-    split_data = []
     show_range = 1
     for (x, _) in splits:
         for y in range(-show_range, show_range + 1):
@@ -206,6 +205,8 @@ def create_neural_net(global_weights=None, train=True):
         _y = tf.placeholder(tf.float32, (None, OUT_SIZE))
         h1 = net.new_conv_layer(name=1, input=_x, num_in_channels=NUM_CHANNELS, num_filters=3, filter_size=5,
                                 global_weights=global_weights, use_pooling=False)
+        #h2 = net.new_conv_layer(name=2, input=h1, num_in_channels=3, num_filters=6, filter_size=5,
+        #                        global_weights=global_weights, use_pooling=False)
         h3 = tf.contrib.layers.flatten(h1)
         h4 = net.new_fc_layer(name=3, input=h3, num_in=h3.shape[1], num_out=64, global_weights=global_weights)
         if False:
@@ -322,4 +323,8 @@ def decide_splitpoints(img, potential_split_points, sessionargs):
         return []
 
 
-
+def convert_pickle_data():
+    entries = read_training_data()
+    with open(definitions.WORD_SPLITTING_PATH, "wb") as out_file:
+        for entry in entries:
+            pickle.dump(entry, out_file, protocol=2)

### 2017-12-10 src/main.py
executing normalization on word, removed bug where words that cut the edge of the image would not be found

@@ -10,11 +10,6 @@ import splitpoint_decision as sd
 import character_recognition as cr
 from character_preprocessing import augment_data
 import word_normalizer as wn
-<<<<<<< HEAD
-
-=======
-import character_combinator as ccb
->>>>>>> 3475134a6ac20274c559b39a8cd625de0b495bf6
 import character_normalizer as cn
 
 from pathlib import Path
@@ -73,7 +68,8 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     # Call character_combinator
 
     # evaluated_chars = evaluate_character_combinations(char_imgs, sessionargs_char_recognition)
-    char_probabilities = cr.imgs_to_text(char_imgs, sessionargs_char_recognition, n=3, verbose=verbose)
+    normalized_character_images = [cn.normalize_character(character_im) for character_im in char_imgs]
+    char_probabilities = cr.imgs_to_text(normalized_character_images, sessionargs_char_recognition, n=3, verbose=verbose)
     return most_likely_words(char_probabilities)
 
 

### 2017-12-10 src/demo.py
final version

@@ -12,14 +12,15 @@ import word_normalizer as wn
 import vocabulary
 import splitpoint_decision as sd
 
-aug_demo = True
-char_rec_demo = True
+aug_demo = False
+char_rec_demo = False
 word_splitting_demo = True
 word_rec_demo = True
-data_creation_demo = True
+data_creation_demo = False
 
 def start_demo():
     char_file = definitions.PROJECT_PATH + 'Data/charset/Img/Sample056/img056-003.png'
+    demo_chars = definitions.PROJECT_PATH + 'demo_data/char/{}.png'
     # word_file = definitions.PROJECT_PATH + 'Data/words/b04/b04-060/b04-060-00-07.png' # Government
     word_file = 'words/beer.png'
     word_image = main.read_image(word_file)
@@ -32,7 +33,7 @@ def start_demo():
         demo_augmentation(char_image)
     if char_rec_demo:
         # input("Character recognition demo")
-        demo_char_recognition(char_image, chrsessionargs)
+        demo_char_recognition(demo_chars, chrsessionargs)
     if data_creation_demo:
         # input("Data creation demo")
         demo_data_creation()
@@ -80,11 +81,14 @@ def demo_augmentation(img):
     cv2.destroyAllWindows()
 
 
-def demo_char_recognition(img, sessionargs):
-    cv2.imshow("Character recognition", img)
-    cv2.waitKey(0)
-    print(cr.img_to_text(img, sessionargs, n=3))
-    cv2.destroyAllWindows()
+def demo_char_recognition(demo_chars, sessionargs):
+    for ordinal in range(ord('a'), ord('z')+1):
+        char = chr(ordinal)
+        img = main.read_image(demo_chars.format(char))
+        cv2.imshow("Character recognition of the letter {}".format(char), img)
+        print(cr.img_to_text(img, sessionargs, n=3))
+        cv2.waitKey(0)
+        cv2.destroyAllWindows()
 
 def demo_word_recognition(img, chrsessionargs, tocsessionargs):
     cv2.imshow("Word recognition", img)

### 2017-12-10 src/main.py

@@ -10,13 +10,6 @@ import splitpoint_decision as sd
 import character_recognition as cr
 from character_preprocessing import augment_data
 import word_normalizer as wn
-<<<<<<< HEAD
-
-=======
-import character_combinator as ccb
->>>>>>> 3475134a6ac20274c559b39a8cd625de0b495bf6
-import character_normalizer as cn
-
 from pathlib import Path
 import shutil
 import definitions

### 2017-12-10 src/splitpoint_decision.py

@@ -39,7 +39,7 @@ def open_images(start=0, amount=-1):
     return images
 
 
-def create_training_data(start=0, amount=10):
+def create_training_data(start=0, amount=1000):
     images = open_images(start, amount)
     n = len(images)
     with open(definitions.WORD_SPLITTING_PATH, "ab") as out_file:

### 2017-12-10 src/vocabulary.py

@@ -108,3 +108,9 @@ def rand(not_indices, amount):
             return rands
     else:
         return []
+
+def test():
+    print(correct_written_words("artificol"))
+    print(correct_written_words("inteligance"))
+
+test()

### 2017-12-10 src/character_extraction.py
small bugfixes

@@ -46,11 +46,11 @@ def extract_character_separations(word_image, postprocess=True, sessionargs=None
     blur = cv2.GaussianBlur(word_image, (1, 1), 0)
     ret3, rotated_threshold = cv2.threshold(blur, 0, 255, cv2.THRESH_BINARY + cv2.THRESH_OTSU)
     splitpoints = find_splits_img(word_image)
+
     if postprocess and not sessionargs is None:  # Use a neural network to check if predicted splitting points are actual splitting points.
         splitpoint_decisions = toc.decide_splitpoints(word_image, splitpoints, sessionargs)
         splitpoints = [split for (is_split, split) in zip(splitpoint_decisions, splitpoints) if is_split]
     rotated_splits.append(splitpoints)
-
     '''
     We have the rotated image and the splits for each of the chosen angles
     Now we identify the image with the most splits,
@@ -141,6 +141,9 @@ def find_splits_img(image):
     ret3, thresh = cv2.threshold(blur, 0, 255, cv2.THRESH_BINARY + cv2.THRESH_OTSU)
 
     skel = skeletonize_thresholded_image(thresh)
+
+    cv2.imshow("skel", skel)
+    cv2.waitKey(0)
 
     # We reduce each collumn to its summation, and normalize it to 1 for each black pixel in the original threshold image
     col_summation = cv2.reduce(skel, 0, cv2.REDUCE_SUM, dtype=cv2.CV_32S) // 255;

### 2017-12-10 src/main.py

@@ -59,7 +59,16 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     :param image: Image of a word
     :return: A list of pairs, the pairs consist of likely words and their probabilities.
     """
+
+    cv2.imshow("n", img)
+    cv2.waitKey(0)
+
     normalized_word_image = wn.normalize_word(img)
+
+
+    cv2.imshow("n", normalized_word_image)
+    cv2.waitKey(0)
+
     char_imgs = ce.extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction, postprocess=postprocess)
 
     # normalized_characters = [cm.normalize_character(character) for character in char_imgs]

### 2017-12-10 src/word_extraction.py

@@ -450,16 +450,17 @@ def preprocess_image(img, file_index = 0):
             # Saving the found word images to a file
             write_word_image(extracted_word, file_number, line_index, word_index)
 
+    new_img = img.copy()
     greyscale = 73
     color_increment = 180 // len(lines)
     for line_index, line in enumerate(lines):
         for word_index, rectangle in enumerate(line):
             (x, y, w, h) = rectangle
-            cv2.rectangle(img,(x,y),(x+w,y+h),(greyscale,255,0),5)
+            cv2.rectangle(new_img,(x,y),(x+w,y+h),(greyscale,255,0),5)
         greyscale += color_increment
 
     # Write threshold image for demonstration pirposes
-    write_threshold_image(img, file_index)
+    write_threshold_image(new_img, file_index)
     write_threshold_image(thresh, file_index+10)
 
     return (extracted_words)

### 2017-12-10 src/word_normalizer.py

@@ -1,13 +1,21 @@
 import cv2
+import numpy as np
 
 NORMALIZED_HEIGHT = 100 #pixels
+PADDING = 5
 def normalize_word(word_image):
 
     height, width = word_image.shape[:2]
+    
 
     difference_percentage = NORMALIZED_HEIGHT / height
     normalized_width = width * difference_percentage
 
     normalized_word = cv2.resize(word_image, (int(normalized_width), NORMALIZED_HEIGHT))
 
-    return normalized_word
+    norm_height, norm_width = normalized_word.shape[:2]
+
+    container = cv2.bitwise_not(np.zeros((norm_height + (2*PADDING), norm_width + (2*PADDING)), np.uint8))
+    container[PADDING:norm_height+PADDING ,PADDING:norm_width+PADDING] = normalized_word
+
+    return container

### 2017-12-10 src/character_extraction.py
removed image showings

@@ -142,9 +142,6 @@ def find_splits_img(image):
 
     skel = skeletonize_thresholded_image(thresh)
 
-    cv2.imshow("skel", skel)
-    cv2.waitKey(0)
-
     # We reduce each collumn to its summation, and normalize it to 1 for each black pixel in the original threshold image
     col_summation = cv2.reduce(skel, 0, cv2.REDUCE_SUM, dtype=cv2.CV_32S) // 255;
     col_summation_list = col_summation[0].tolist()

### 2017-12-10 src/main.py

@@ -57,15 +57,7 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     :param image: Image of a word
     :return: A list of pairs, the pairs consist of likely words and their probabilities.
     """
-
-    cv2.imshow("n", img)
-    cv2.waitKey(0)
-
     normalized_word_image = wn.normalize_word(img)
-
-
-    cv2.imshow("n", normalized_word_image)
-    cv2.waitKey(0)
 
     char_imgs = ce.extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction, postprocess=postprocess)
 

### 2017-12-11 src/word_normalizer.py
undone padding, not needed, was another bug

@@ -6,16 +6,16 @@ PADDING = 5
 def normalize_word(word_image):
 
     height, width = word_image.shape[:2]
-    
+
 
     difference_percentage = NORMALIZED_HEIGHT / height
     normalized_width = width * difference_percentage
 
     normalized_word = cv2.resize(word_image, (int(normalized_width), NORMALIZED_HEIGHT))
 
-    norm_height, norm_width = normalized_word.shape[:2]
+    #norm_height, norm_width = normalized_word.shape[:2]
 
-    container = cv2.bitwise_not(np.zeros((norm_height + (2*PADDING), norm_width + (2*PADDING)), np.uint8))
-    container[PADDING:norm_height+PADDING ,PADDING:norm_width+PADDING] = normalized_word
+    #container = cv2.bitwise_not(np.zeros((norm_height + (2*PADDING), norm_width + (2*PADDING)), np.uint8))
+    #container[PADDING:norm_height+PADDING ,PADDING:norm_width+PADDING] = normalized_word
 
-    return container
+    return normalized_word

### 2017-12-13 Report/paper.tex
redudant copy of model, and new notes to paper
renames and improve demo
demo, updates report, updates preprocessing

@@
 ]
-
+% Notes:
+% - Add https://tel.archives-ouvertes.fr/tel-01249405/document reference, (see abstract) 
+% in order to compare results. 
+% - Disciss possible flaws. (Use images from IAM in order to do so)
+% - Update outdated parts, like neural net of splitpoint evaluation
+%  Reference datasets, Chars74K, IAM, Google n-grams
+%  redo the preprocessing experiment and explain why some techniques increase the accuracy more than others. 
 \begin{abstract} 
-This paper handles off-line handwriting recognition. We created a system which converts images with handwriting to text. 
-First we explain how to recognise characters. This explanation involves the used dataset, preprocessing techniques and the deep neural network for classifying characters. 
-Next we describe how to segment a text into words using a contouring algorithm with consequent postprocessing to remove unwanted noise. 
-After this we present our approach to seperating the characters from words using vertical projections combined with postprocessing techniques including a neural network for splitpoint evaluation. 
-We try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
-Finally we discuss our results and some problems we encountered.
-
+The aim of this paper is to present a system for off-line handwriting recognition. 
+We discuss how to segment texts with contouring. Next we present our method of splitting words into characters. 
+Then we describe how to recognise characters with a deep neural network approach. 
+Finally we present some postprocessing techniques to correct mispredicted text. 
 \end{abstract}
@@
 \section{Introduction}
-\section{Handwriting Recognition}
+Despite the expontially increasing amount of digital information in the contemporary world, a lot of new documents are still handwritten. 
+Handwritten documents are very inefficient as they need to be read manually in order to extract information.
+These documents need manual labour and physical space in order to be stored in an organized manner. 
+Finally, it takes a lot of time, effort and resources to move these documents to another location.
 
-\subsection{Character Recognition}
+This paper presents an off-line handwriting recognition system.
+This system digitizes handwritten texts, which allows modern data processing, storage a transmission techniques to be applied to the information stored in the handwritten documents.
+We only cover off-line handwriting recognition in this paper, where we have the full handwritten text in advance.
+This problem is similar to Optical Character recognition (OCR).
+OCR digitizes printed texts, this is generally easier to recognise as it contains less variation than handwritten texts.
+On-line recognition on the other hand converts text as it is written (on a touchscreen for example).
+Other information is available for this problem, e.g.
+the exact movements of the pen.
+\cite{olcr} present a technology for online recognition of Japanese handwritten text with an accuracy of 94.6\%.
+
+First we explain how to recognise characters in Section \ref{sec:charrec}. This explanation involves the used dataset, preprocessing techniques and the deep neural network for classifying characters. 
+In Section \ref{sec:segtext} we describe how to segment a text into words using a contouring algorithm with consequent postprocessing to remove unwanted noise. 
+After this we present our approach to seperating the characters from words in \ref{seg:segword}.
+We use vertical projections combined with postprocessing techniques including a neural network for splitpoint evaluation.
+Postprocessing steps are presented in Section \ref{seg:postproc}, we try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
+We conclude this paper with some experimental results, the problems we encountered combined with possible improvements.
+
+\section{Character Recognition}
+\label{sec:charrec}
 The smallest distinguishable token in a text is a character. These handwritten characters are uniquely 
@@
 We take another approach, instead of adding rules to minimize variations we made a deep neural net which can recognise any character with a certain probability.
-The architecture of this deep neural network is described in \ref{sec:dnn}.
+The architecture of this deep neural network is described in Section \ref{sec:dnn}.
 In order to make the deep neural network recognise characters in images with these variations, we augmented the dataset to include a bigger variety of images.
-We present these augmentation techniques in \ref{sec:preproc}.
-The dataset of handwritten characters is described in \ref{sec:data}.
+We present these augmentation techniques in Section \ref{sec:preproc}.
+The dataset of handwritten characters is described in Section \ref{sec:data}.
 
-\subsubsection{Data}
+\subsection{Data}
 \label{sec:data}
-A good data dataset with many examples, enough variation, and which is balanced between classes is very important to train the neural network effectively. 
+Before exposing the neural network to new characters, it needs to be familiar with many training samples to be able to classify any new character correctly. 
+This is a why a good dataset is very important. A dataset should have many examples, much variation, and an even balance between classes. A dataset with these properties 
+allow the neural network to train effectively and recognise any new character given to the network.
+
 We use the Chars74K dataset for this purpose. It has 3410 images with handwritten characters, each belonging to one of 62 classes (0-9, A-Z, a-z). 
-This dataset contains both very easily recognisable images of characters but also characters which are even hard to recognise for humans, as shown in figure \ref{fig:char}. 
-3410 input images is quite low for a neural network to be effectively trained. To overcome this problem we employ some data augmentation techniques discussed in \ref{par:aug}.
+This dataset contains both very easily recognisable images of characters but also characters which are even hard to recognise for humans, as shown in Figure \ref{fig:char}. 
+3410 input images is quite low for a neural network to be effectively trained. To overcome this problem we employ some data augmentation techniques discussed in Section \ref{par:aug}.
 
@@
 
-\subsubsection{Preliminary steps}
+\subsection{Preliminary steps}
 \label{sec:preproc}
@@
 
-\paragraph{Normalizing data} 
+\subsubsection{Normalizing data} 
 \label{par:norm}
 Theoretically feeding normalized data to the neural network returns the same output as before. 
-In reality however, it's better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}.
+In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}.
 We normalize the data by reducing the input dimensions from (1200,900,3) to (64,64,1). A simple threshold is added to improve the contrast in our image. 
@@
 
-\paragraph{Augmenting data}
+\subsubsection{Data augmentation}
 \label{par:aug}
@@
 As our dataset only contains 3410 images with handwritten characters, we need more data for the neural network to be able to function properly.
-Increasing the size of our dataset has great impact on the performance of the neural network as discussed in \ref{sec:expres}.
+Increasing the size of our dataset has great impact on the performance of the neural network as discussed in Section \ref{sec:expres}.
 Now we will describe the data augmentation techniques we used.
@@
 
-\subparagraph{Adding noise to data}
+\paragraph{Adding noise to data}
 We add a variant of each image which includes noise troughout the image with the purpose of reducing overfitting of the neural network \cite{DataNoise}.
-This noise is taken from the Gaussian distribution and summed with the original image.
-\subparagraph{Translations of data}
+This noise is taken from the Gaussian distribution with a mean of 0 and standard deviation of 0.004.
+The noise is summed with the original image of which the pixels are normalized between $[-1, 1]$.
+\paragraph{Translations of data}
 The dataset is also extended with variations in padding of characters in images. For this purpose we multiply every coordinate of the original image with the following transformation matrix: 
@@
 where $t_x$ is the horizontal shift and $t_y$ is the vertical shift.
-\subparagraph{Rotations of data}
-Rotations of the image are added to the dataset to make the network more robust to handwritings with different orientations. We rotate by -30 and 30 degrees. This matrix is equal to:
+The values for chosen for $t_x$ are -8, 0 and 8, the values chosen for $t_y$ are -4, 0 and 4.
+All combinations are used except $t_x=0$ and $t_y=0$.
+The translation only occurs if it does not displace the character outside of the image.
+\paragraph{Rotations of data}
+Rotations of the image are added to the dataset to make the network more robust to handwritings with different orientations. This matrix is equal to:
 \begin{equation}
@@
 \end{equation}
-where $\alpha = cos(angle)$ and $\beta = sin(angle)$. Center is the coordinate in our image around which is rotated and the angle describes the amount of degrees to rotate in the clockwise direction. 
-\subparagraph{Scaling of data}
+where $\alpha = cos(angle)$ and $\beta = sin(angle)$.
+$center$ is the coordinate in our image around which is rotated and the angle describes the amount of degrees to rotate in the clockwise direction.
+We use -30 and 30 as angle.
+\paragraph{Scaling of data}
 With the purpose of making the network more robust to different sizes of handwriting, we add different scalings of the data.
-We both scale in width and in height.
+We scale in both the horizontal and vertical direction.
 The following transformation matrix allows us to scale the input image:
@@
 where $s_x$ is the horizontal scaling factor and $s_y$ is the vertical scaling factor. 
-\subparagraph{Shearing of data}
+We use combinations of values in $[0.75, 1, 1.25, 1.5]$ for $s_x$ and $s_y$ depending on the original size of the character.
+\paragraph{Shearing of data}
 In order to deal with different kinds of slants we add sheared versions of the image to the dataset. 
@@
 \end{equation}
-\subparagraph{Erosion of data}
+\paragraph{Erosion of data}
 The erosion operation of images adds thinner versions of characters to the dataset. Erosion uses a kernel to convolute an image. 
@@
 This pixel value of the anchor point of the kernel is replaced with the maximum pixel value.
-This causes bright regions to get thinner, and dark regions to get bigger in the image. A visualisation of this is shown in \ref{fig:erosion}. 
+This causes bright regions to get thinner and dark regions to get bigger in the image.  
+All of the previously discussed preprocessing techniques are also applied to the eroded image.
+The final dataset contains translated, orientated, sheared, scaled and noisy versions of both the original image and the eroded image.
+A visualitation of the augmentation of an augmented picture can be seen in Figure \ref{fig:augmented}.
+
 \begin{figure}
+\begin{subfigure}{0.11\textwidth}
   \centering
-  \includegraphics[width=\linewidth]{images/erosion}
-  \caption{Visualisation of eroding an inverted image with a handwritten 'j'. The right picture underwent erosion \cite{erosion}.}
-\label{fig:erosion}
+  \includegraphics[width=\linewidth]{images/original}
+  \caption{Original image}
+\end{subfigure}
+\begin{subfigure}{0.11\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{images/noisy}
+  \caption{Noisy image}
+\end{subfigure}
+\begin{subfigure}{0.11\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{images/translated}
+  \caption{Translated image}
+\end{subfigure}
+\begin{subfigure}{0.11\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{images/rotated}
+  \caption{Rotated image}
+\end{subfigure}
+\begin{subfigure}{0.11\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{images/scaled}
+  \caption{Scaled image}
+\end{subfigure}
+\begin{subfigure}{0.11\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{images/sheared}
+  \caption{Sheared image}
+\end{subfigure}
+\begin{subfigure}{0.11\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{images/eroded}
+  \caption{Eroded image}
+\end{subfigure}
+\caption{All of the augmentations applied to one image.}
+\label{fig:augmented}
 \end{figure}
-All of the previously discussed preprocessing techniques are also applied to the eroded image.
-The final dataset contains translated, orientated, sheared, scaled and noisy versions of both the original image, and the eroded image.
-\subsubsection{Deep neural network}
+\subsection{Deep neural network}
 \label{sec:dnn}
-
-\paragraph{Introduction}
+% TODO: make a drawing, see http://www.ais.uni-bonn.de/deep_learning/images/Convolutional_NN.jpg
 Humans have millions of connected neurons in the visual cortices which are capable of image processing.
-We, humans, can recognise most characters without any thought because of this complex neural network inside our brain. 
+We can recognise most characters without any thought because of this complex neural network inside our brain. 
 For computers, character recognition is a more difficult task.
-We could try to write rule-based systems and define how every character should look like.
-This solution isn't flexible at all and is very hard to define.
+We could try to write rule-based systems and define how every character should look.
+This solution is not flexible at all and is very hard to define.
 We use a neural network, inspired by how the brain processes images.
@@
 The perceptrons take an input and calculate an output using an activation function. This allows perceptrons to make decisions.  
-Normalized images can be fed to the neural network, which then propagate trough several layers and eventually produce as output the probabiliy that the input image has a certain class, one of the possible characters.
-Now follows a detailed overview of the actual deep neural network.
+Normalized images can be fed to the neural network, which then propagate through several layers and eventually produce as output the probability that the input image has a certain class, one of the possible characters.
+Now follows a detailed overview of our deep neural network. A detailed explanation on how deep neural networks function is explained by \cite{nnbook}.
 
-\paragraph{Convolutional layers}
+\subsubsection{Convolutional layers}
  The input is a normalized image with one color channel.
 This is fed to a convolutational layer.
-Convolutional layers allow us extract features by scanning the image with filters.
-These features might be edges, corners, or any other type of shape in the image.
+Convolutional layers allow us to extract features by scanning the image with filters.
+These features might be edges, corners or any other type of shape in the image.
 The extracted features are called a feature map.
  We can apply several of those filters using one convolutional layer to extract several types of features. 
- This works because images have spatial information, close pixels have shared information.
+ This works because images have a spatial structure: close pixels have shared information.
 By applying a filter, pixels in the same region will use shared biases and weights, which allows the filter to extract particular features.
@@
 Pooling layers remove the positional information of the features in the image.
-This is useful because the exact location is less important than the location relative to the other features, and this allows us to reduce the amount of parameters required in the next layers.
-We use 3 convolutional layers combined with max pooling layers in order to extract low, medium, and high level features. 
-\paragraph{Fully-connected layers}
-Next we flatten the output of the last convolutional layer in order to extract a 1-dimensional array from the 3-dimensional feature maps.
-This 1-dimensional array can now be fed to a fully connected layer, which is a layer of connected perceptrons as previously discussed.
-There are 3 fully connected layers each connected with a dropout layer.
-The purpose of this layer is to remove activations at random.
+This is useful because the exact location is less important than the location relative to the other features and this allows us to reduce the amount of parameters required in the next layers.
+We use three convolutional layers combined with max pooling layers in order to extract low, medium and high level features. 
+\subsubsection{Fully-connected layers}
+Next we flatten the output of the last convolutional layer in order to extract a one-dimensional array from the three-dimensional feature maps.
+This one-dimensional array can now be fed to a fully-connected layer, which is a layer of connected perceptrons as previously discussed.
+There are three fully-connected layers each connected with a dropout layer.
+Dropout randomly drops units and their connections from the network during training.
 Dropout greatly reduces overfitting in the fully connected layers.
-We don't use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
-The last fully connected layer has 62 output neurons, which correspond to the 62 possible output classes. We use rectified linear units in order to compute the activation function.
-\paragraph{Softmax layer}
+\cite{dropout} We do not use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
+The last fully connected layer has 62 output neurons, which correspond to the 62 possible output classes.
+We use rectified linear units in order to compute the activation function.
+A rectified linear unit or rectified linear neuron is a unit which uses a rectifier as linear unit.
+It's output is given by
+\begin{equation}
+        max(0, w*x+b)
+\end{equation}
+where $x$ is the input, $w$ is the weight vector and $b$ is the bias. 
+\subsubsection{Softmax layer}
  The last layer in the neural network is a softmax layer with a cross-entropy cost function.
 The softmax layer compute a probability distribution (a set of 62 numbers which sum up to 1) using the 62 output neurons from the previous layer.
-An element in the probability distribution estimates the chance of the input image to classify as a certain character.
+An element in the probability distribution estimates the probability of the input image to classify as a certain character.
  The cross-entropy cost function uses the actual labels to calculate the error between the actual labels and the predicted labels (the previously discussed probabilities).
 The closer the cross-entropy function is to zero, the better it is at classifying handwritten characters.
- \paragraph{Weight adjustment}
+ \subsubsection{Weight adjustment}
 Finally, in order to actually train the neural network, we need some kind of weight adjustment.
@@
 Each of these methods effectively improve the previously mentionned method. 
-The Adam optimizer is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter. This optimizer works best in our case and will be used.
+The Adam optimizer, introduced by \cite{adam}, is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
+This optimizer achieves the highest accuracy in our case and will be used.
 
-\subsection{Segmentation of text into words}
-\subsection{Segmentation of words into characters}
-\subsubsection{Correction of oversegmentation}
+\section{Segmentation of text into words}
+\ref{sec:segtext}
+\section{Segmentation of words into characters}
+\ref{sec:segword}
+\subsection{Correction of oversegmentation}
 As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these splitting points. 
-\cite{evalsplitpoints} discusses 3 approaches to evaluate split point correctness: rule based approaches, cost function optimization based approaches and machine learning based approaches. 
-In this paper we will only discuss the latter, machine learning based approaches, as we used this approach. 
+\cite{evalsplitpoints} discusses three approaches to evaluate split point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
+In this paper we will only discuss the latter, machine learning based procedures, as we used this approach. 
 In particular, our algorithm is based on \cite{evalsplitpointsnn}.
@@
 Next, small windows of equal size are extracted from the pixel matrix and for each window the density is calculated.
-The is the number of black pixels divided by the total amount of pixels in the window.
+This is the number of black pixels divided by the total amount of pixels in the window.
 With every matrix a corresponding label is encoded for storage in the training file, this label is encoded as 0.9 if the splitting point is correct and 0.1 otherwise.
@@
 This allows us to use convolutational layers in order to extract features, instead of manually creating features (features which show the density in this case).
-We use 1 convolutional layer and 3 fully connected layers in our neural network.
+We use one convolutional layer and three fully connected layers in our neural network.
 This neural network takes the actual image of the splitting point as input (a pixel matrix of the pixels surrounding the splitting point) and returns a one-hot encoded vector indicating if the splitting is correct or incorrect.
-This allowed us to achieve better results than our implementation of the algorithm exactly as described by the paper achieved.
+This method allowed us to achieve a higher accuracy than an implementation of the method described by \cite{evalsplitpointsnn}.
 Splitting points can now be filtered if the neural network decides a splitting point is incorrect.
@@
 \subsection{Natural language processing}
-To cope with some of the high error rates as shown in section \ref{sec:expres}, there needs to be some kind of correction.
-This correction should allow the system to produce possibble written text even when a step of the larger system fails.
+To cope with some of the high error rates as shown in Section \ref{sec:expres}, there needs to be some kind of correction.
+This correction should allow the system to reproduce possibble written text even when a step of the larger system fails.
 In order to do this we adapt natural language processing techniques as postprocessing steps.
 Natural language techniques can often be found in speech recognition, language generation and other systems.
-As a first step we match words against a dictionary to find out if they make sense, this is discussed in section \ref{sec:voc}.
-The second step checks if those given words make sense in the context, this is discussed in section \ref{sec:lm}.
+As a first step we match words against a dictionary to find out if they are syntactically correct, this is discussed in Section \ref{sec:voc}.
+The second step checks if those given words make sense in the context, this is discussed in Section \ref{sec:lm}.
 
@@
 \label{sec:voc}
-The neural network does not only produce which character is recognised, but rather a list of probabilities indicating the chance of an image to be a certain character.
+The neural network does not only produce which character is recognised, but rather a list of probabilities indicating the likelihood of an image to be a certain character.
 We can use all of this information in this postprocessing step instead of throwing that information away and naively returning the most likely character as the actual character.
@@
 \end{figure} 
-We make a tree, as shown in figure \ref{fig:wordtree}, where every node is a tuple with a word (sequence of characters) and a probability (likeliness of that word based on the probabilities of individual characters).
+We make a tree, as shown in Figure \ref{fig:wordtree}, where every node is a tuple with a word (sequence of characters) and a probability (likeliness of that word based on the probabilities of individual characters).
 The probability is calculated as the product of probability for each character: $P(c_1, c_2,...,c_k) = \prod\limits_{i}{c_i}$ with $1 \leq i \leq k$.
-In order to keep the complexity low we only consider the 3 most probable characters.
+In order to keep the complexity low we only consider the three most probable characters.
 This will be the branching factor of every tree.
-There's a tree for every possible starting character.
+A tree exists for every possible starting character.
 Now we have a list of possible words and their probabilities (every leaf).
@@
 If we only want to convert the image of one word into text, we can just return the word with the highest score.
-This finalises the logic for recognising handwritten words. Next we will discuss a postprocessing technique useful we the word is part of a bigger context.
+This finalises the logic for recognising handwritten words. Next we will discuss a postprocessing technique which is useful when the word is part of a bigger context.
 \subsubsection{Language model}
@@
 We can check the previous words and calculate the likeliness of a word to be the next word in the context.
-N-gram models are used to calculate probabilities given n-1 previous words.
+N-gram models are used to calculate probabilities given $n-1$ previous words.
 The probability is computed as $P(w_i | w_{i-n},...,w_{i-1})$.
-We use the Markov assumption here: the probability of a word in the context can be approximated by the previous n-1 words. This probability is calculated as follows: 
+We use the Markov assumption here: the probability of a word in the context can be approximated by the previous $n-1$ words. \cite{markov} This probability is calculated as follows: 
 \begin{equation}
@@
 \end{equation}
-where $w_i$ is the word of which we want to calculate the probability and $count(w_i,...w_{i+k})$ indicates the amount of occurences of $w_i,...,w_{i+k}$.
+where $w_i$ is the word of which we want to calculate the probability and $count(w_i,...w_{i+n})$ indicates the amount of occurences of $w_i,...,w_{i+n}$.
 That amount is the same as counting the amount of n-grams with that sequence of words.
-An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+k}$.
+An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+n}$.
+\cite{ngrams} % TODO google citation%
 We use the n-grams from the Google dataset for this purpose.
-Querying these datasets locally takes several minutes, and this happens for every word.
+\cite{google} Querying these datasets locally takes several minutes and this happens for every word.
 Instead we use a Python library phrasefinder to make online queries to these corpora.
@@
 Now we can make queries such as "I like dogs / cats / sheep".
-Using the Google corpora probabilities are calculated with n-gram models for each of the 3 words.
+Using the Google corpora probabilities are calculated with n-gram models for each of the three words.
 This gives us a probability for every word based on the context.
-Combining the results with the probabilities calculated in section \ref{sec:voc}, we can 
+Combining the results with the probabilities calculated in Section \ref{sec:voc}, we can 
 calculate a new score. The word with the highest score is accepted as written word. 
@@
 We tested a lot of different configurations for the deep neural network in order to find a good functioning neural network given the low amounts of data.
-A problem which arrised when trying to test many diffeent configurations was the long time it takes to train a neural network.
-In order to limit the time these experiments run we added some small constraints.
-Next it's important to note that a lot of non-determinism is part of these neural networks.
+A problem which arrised when trying to test many different configurations was the long time it takes to train a neural network.
+In order to limit the time these experiments run we added some small constraints. We only run every experiment 200 epochs, running these experiments longer might result in different conclusions. 
+Next it is important to note that a lot of stochastic behaviour is part of these neural networks because of the dropout layers.
 When running the neural network twice with the same configuration, it could have completely different results.
-Another note is that some experiments might not use the finalized setup of our neural network. These experiments do use a same setup to compare different configurations. 
-In order to have meaningful results which run in a reasonable amount of time we ran every experiment 4 times but limited the amount of epochs to 200.
-This setup is important to note, as some results might be quite different if we run more tests or use more iterations.
-When discussing these results we take the best result of the 4 training results.
-We use this result as we are only interested in the potential of the neural network, and not possible outliers which perform bad due to poor weight adjustments.  
+This is why we run every configuration four times.
+Only the result with the highest accuracy is considered.
+This way we can avoid outliers.
+Another note is that some experiments might not use the finalized setup of our neural network. These experiments do use the same setup to compare different configurations. 
 \subsubsection{Experiments}
@@
 Figure \ref{fig:learningrate} shows the impact of learning rate on the training of our neural network. 
-For this experiment we do have to note that we're using the Adam optimizer to train our neural network which computes individual learning rates for every parameter every training step.
+For this experiment we do have to note that we are using the Adam optimizer to train our neural network which computes individual learning rates for every parameter every training step.
 The learning rate which we pass is the initial learning rate.
-Choosing a high learning rate $10^-3$ allows the neural network to converge to a high accuracy (76 \%) very fast, but once reached the accuracy fails to grow any more.
+Choosing a high learning rate $10^{-3}$ allows the neural network to converge to a high accuracy (76 \%) very fast, but once reached the accuracy fails to grow any more.
 This in contrast to higher learning rates which reach a good accuracy more slowly but find a more optimal solution.
-Learning rate $10^-4$ reaches an accuracy of 68\% after 200 iterations but is still growing.
-Learning rate $10^-5$ only reaches an accuracy of 5\% after 200 iterations but should find a better solution eventually than the other two configurations.
+Learning rate $10^{-4}$ reaches an accuracy of 68\% after 200 iterations but is still growing.
+Learning rate $10^{-5}$ only reaches an accuracy of 5\% after 200 iterations but should find a better solution eventually than the other two configurations.
 \paragraph{Batch size}
-In figure \ref{fig:batchsize} we can see the impact of choosing different batch sizes which are fed to the neural network.
-In order to increase to speed at which the neural network is trained, we feed batches of input images to the neural network instead of one image at a time.
+n Figure \ref{fig:batchsize} we can see the impact of choosing different batch sizes which are fed to the neural network.
+In order to increase the speed at which the neural network is trained, we feed batches of input images to the neural network instead of one image at a time.
 Weight adjustment only occurs after feeding a complete batch to the network.
 In the figure we can see bigger batch sizes increase the execution speed.
-However these differences are quites small.
+However these differences are quite small.
 The configuration with batch size executes 200 iterations in one minute less than a configuration with batch size 128.
@@
 In order to regulize the neural network we add weight decay to the network loss function in order to penalize big weights.
-The impact of different decays can be seen in \ref{fig:decay}.
+The impact of different decays can be seen in Figure \ref{fig:decay}.
 A high decay functions best (71\% accuracy) because we have a small amount of training examples.

### 2017-12-13 src/character_preprocessing.py

@@ -2,7 +2,7 @@ import time as t
 
 import cv2
 import numpy as np
-from random import getrandbits
+import random
 import definitions
 
 '''
@@ -87,34 +87,55 @@ def erodeImage(image):
 
 # Expects a normalized image as input.
 # Returns an array of augmented images.
-def augmentImage(image, add_noise, add_rotations, add_translations, add_scales, add_shearing):
+def augmentImage(image, add_noise, add_rotations, add_translations, add_scales, add_shearing, add_one=False):
+    """
+    Augment an image with the chosen techniques.
+    :param image: Image to augment
+    :param add_noise:
+    :param add_rotations:
+    :param add_translations:
+    :param add_scales:
+    :param add_shearing:
+    :param add_one: True -> Indicates one of the given techniques, False -> add atleast one of the given technique.
+    :return:
+    """
     # Array with augmented images
     images = [image]
 
     up, down, left, right = imageBorders(image)
     ## Addition of Gaussian noise
     if add_noise:
-        images.append(noisyImage(image, stddev=0.004))
+        images.append(noisyImage(image, stddev=0.05))
 
     ## Rotations of image
     if add_rotations:
+        rotatedImages = []
         for angle in np.arange(-30, 60, 30):  # evenly spaced values within a given interval
-            if (angle != 0):
-                if bool(getrandbits(1)):  # Enough variation in rotation, but limit amount of examples
-                    images.append(rotateImage(image, angle))
+            if angle != 0:
+                    rotatedImages.append(rotateImage(image, angle))
+        if add_one:
+            images.append(random.choice(rotatedImages))
+        else:
+            images.extend(rotatedImages)
 
     ## Translation of image
     if add_translations:
-        for tx in range(-8, 16, 8):
-            for ty in range(-4, 4, 4):
-                if bool(getrandbits(1)):  # Enough variation in translation, but limit amount of examples
+        translatedImages = []
+        for tx in range(-16, 32, 16):
+            for ty in range(-8, 8, 8):
+                # if bool(getrandbits(1)):  # Enough variation in translation, but limit amount of examples
                     canTranslate = inBounds(up + ty) and inBounds(down + ty) and inBounds(left + tx) and inBounds(
                         right + tx)
                     if not tx == 0 and not ty == 0 and canTranslate:
[…]

### 2017-12-13 src/character_recognition.py

@@ -155,7 +155,7 @@ def create_training_operation(h, _y, learning_rate=LEARNING_RATE, decay=DECAY, g
     return training_operation
 
 
-def train_net(n, restore=False, min_save=1.0, iteration=1):
+def train_net(n, restore=False, min_save=1.0, iteration=1, name='all'):
     """
     Trains the network for n epochs with a new session
     Note: if this function has never been run, set restore to false!
@@ -199,7 +199,7 @@ def train_net(n, restore=False, min_save=1.0, iteration=1):
             print("New maximum accuracy {} achieved.".format(validation_accuracy))
             save_session(session)
             min_save = validation_accuracy
-    save_output("all", time=times, accuracies=accuracies, iteration=iteration)
+    save_output(name, time=times, accuracies=accuracies, iteration=iteration)
     t = str(time() - start)
     print("The training took: " + str(t) + " seconds.")
     return session

### 2017-12-13 src/demo.py

@@ -12,11 +12,11 @@ import word_normalizer as wn
 import vocabulary
 import splitpoint_decision as sd
 
-aug_demo = False
-char_rec_demo = False
+aug_demo = True
+char_rec_demo = True
+data_creation_demo = True
 word_splitting_demo = True
 word_rec_demo = True
-data_creation_demo = False
 
 def start_demo():
     char_file = definitions.PROJECT_PATH + 'Data/charset/Img/Sample056/img056-003.png'

### 2017-12-13 src/experiments/experiment_results.py

@@ -99,13 +99,14 @@ def visualise_experiment_configuration(experiment_name, conf, save=False):
 #visualise_experiment_configuration("preprocess", conf="all", save=False)
 
 def visualise_all():
-    accuracy_path = definitions.EXPERIMENTS_CHAR_PATH + 'all/accuracy_907.txt'
-    time_path = definitions.EXPERIMENTS_CHAR_PATH + 'all/time_907.txt'
+    accuracy_path = definitions.EXPERIMENTS_CHAR_PATH + 'all/accuracy_917.txt'
+    time_path = definitions.EXPERIMENTS_CHAR_PATH + 'all/time_917.txt'
     accuracy = np.loadtxt(accuracy_path)
     time = np.loadtxt(time_path)
-    end = 500
-    visualise([accuracy[:end]], [time[:end]], name='final_{}'.format(end), save=True)
-    print(np.max([accuracy[:end]]))
+    start = 0
+    end = 3000
+    visualise([accuracy[start:end]], [time[start:end]], name='final_{}'.format(end), save=True)
+    max_i = np.argmax(accuracy[start:end])
+    print(max_i, accuracy[max_i])
 
 
-visualise_all()

### 2017-12-13 src/main.py

@@ -11,6 +11,7 @@ import character_recognition as cr
 from character_preprocessing import augment_data
 import character_normalizer as cn
 from pathlib import Path
+import word_normalizer as wn
 import shutil
 import definitions
 import os
@@ -57,9 +58,9 @@ def recognise_possible_words(img, sessionargs_char_recognition, sessionargs_over
     :param image: Image of a word
     :return: A list of pairs, the pairs consist of likely words and their probabilities.
     """
-    normalized_word_image = wn.normalize_word(img)
+    # normalized_word_image = wn.normalize_word(img)
 
-    char_imgs = ce.extract_characters(normalized_word_image, sessionargs=sessionargs_oversegmentation_correction, postprocess=postprocess)
+    char_imgs = ce.extract_characters(img, sessionargs=sessionargs_oversegmentation_correction, postprocess=postprocess)
 
     # normalized_characters = [cm.normalize_character(character) for character in char_imgs]
     # ccb.normalize_and_combine_characters(char_imgs, sessionargs_char_recognition)
@@ -97,7 +98,7 @@ def recognise_text(file_name):
             max([(word, alpha * voc_words[word] + beta * prob) for word, prob in lang_words.items()],
                 key=lambda x: x[0])[0]
         text.append(most_likely_word)
-        print("Found word: " + most_likely_word[0])
+        print("Found word: " + most_likely_word)
     return ' '.join(text)
 
 

### 2017-12-13 src/splitpoint_decision.py

@@ -39,7 +39,7 @@ def open_images(start=0, amount=-1):
     return images
 
 
-def create_training_data(start=0, amount=1000):
+def create_training_data(start=0, amount=10):
     images = open_images(start, amount)
     n = len(images)
     with open(definitions.WORD_SPLITTING_PATH, "ab") as out_file:

### 2017-12-13 src/utils/experiments.py

@@ -4,7 +4,7 @@ import time as t
 import tensorflow as tf
 from numpy import savetxt
 
-import experiments.experiment_nets as en
+from src.experiments import experiment_nets as en
 import definitions, character_recognition, character_preprocessing
 
 """
@@ -94,6 +94,12 @@ def compare_epochs():
         print("Running experiments with " + str(x) + " iterations")
         run_experiment("epochs_" + str(x), batch_size=x)
 
+def run_experiment_with_orig_net(name, iterations=200):
+    for i in range(iterations):
+        with tf.variable_scope("Experiment_" + name + "_" + str(i)):
+            print("Experiment: " + str(i))
+            character_recognition.train_net(name=name, iteration=i, n=iterations)
+
 
 def compare_preprocessing_params():
     """
@@ -102,26 +108,29 @@ def compare_preprocessing_params():
     """
 
     character_preprocessing.augment_data(add_noise=False, add_rotations=False, add_scales=False, add_translations=False,
-                                         add_shearing=False)
-    run_experiment("preprocess_" + "none")
+                                         add_shearing=False, add_one=True, confirm=False, add_erode=False)
+    run_experiment_with_orig_net("preprocess_" + "none")
     character_preprocessing.augment_data(add_noise=True, add_rotations=False, add_scales=False, add_translations=False,
-                                         add_shearing=False)
-    run_experiment("preprocess_" + "noise")
+                                         add_shearing=False, add_one=True, confirm=False, add_erode=False)
+    run_experiment_with_orig_net("preprocess_" + "noise")
     character_preprocessing.augment_data(add_noise=False, add_rotations=True, add_scales=False, add_translations=False,
-                                         add_shearing=False)
-    run_experiment("preprocess_" + "rotate")
+                                         add_shearing=False, add_one=True, confirm=False, add_erode=False)
+    run_experiment_with_orig_net("preprocess_" + "rotate")
     character_preprocessing.augment_data(add_noise=False, add_rotations=False, add_scales=True, add_translations=False,
-                                         add_shearing=False)
-    run_experiment("preprocess_" + "scale")
+                                         add_shearing=False, add_one=True, confirm=False, add_erode=False)
+    run_experiment_with_orig_net("preprocess_" + "scale")
     character_preprocessing.augment_data(add_noise=False, add_rotations=False, add_scales=False, add_translations=True,
-                                         add_shearing=False)
[…]

### 2017-12-13 src/word_normalizer.py

@@ -5,6 +5,8 @@ NORMALIZED_HEIGHT = 100 #pixels
 PADDING = 5
 def normalize_word(word_image):
 
+    return (word_image)
+
     height, width = word_image.shape[:2]
 
 
@@ -19,3 +21,4 @@ def normalize_word(word_image):
     #container[PADDING:norm_height+PADDING ,PADDING:norm_width+PADDING] = normalized_word
 
     return normalized_word
+

### 2017-12-17 src/character_preprocessing.py
experiments

@@ -153,7 +153,7 @@ def augmentImage(image, add_noise, add_rotations, add_translations, add_scales, 
                 images.extend(scaledImages)
     if add_shearing:
         # Shear images to make the dataset more robust to different slant when writing.
-        shearedImages = [shearImage(image, 0.2), shearImage(image), -0.2]
+        shearedImages = [shearImage(image, 0.2), shearImage(image, -0.2)]
         if(add_one):
             images.append(shearedImages[random.randrange(0, len(shearedImages))])
         else:
@@ -183,7 +183,7 @@ def preprocess_image(img, inverse=False):
     return np.reshape(img, definitions.IMG_SHAPE)
 
 
-def augment_data(add_noise=True, add_rotations=True, add_translations=True, add_scales=True, add_shearing=True,  add_erode = True, add_one=False, confirm=True):
+def augment_data(add_noise=False, add_rotations=True, add_translations=True, add_scales=True, add_shearing=True,  add_erode = True, add_one=False, confirm=True):
     if confirm:
         inp = input("Are you sure you want to preprocess the data? Insert y(es) to continue.\n")
         if inp != 'y' and inp != 'yes':
@@ -222,6 +222,59 @@ def augment_data(add_noise=True, add_rotations=True, add_translations=True, add_
     end = t.time()
     print('\nThe execution time is {}'.format(end - start))
 
+def augment_data_fake():
+    in_file_names = open(definitions.CHARSET_INFO_PATH, 'r').read().splitlines()
+    out_file_names = open(definitions.PREPROCESSED_CHARSET_INFO_PATH, 'w')
+
+    i = 0
+    for file_name in in_file_names:
+        # Input
+        label = int(file_name.split('/')[1][-2:])
+        file_name = definitions.CHARSET_PATH + file_name
+        img = preprocess_image(read_image(file_name), inverse=True)
+        # Data augmentation
+        images = [img, img]
+        # Output
+        for aug_img in images:
+            aug_img = np.multiply(255, np.subtract(1,
+                                                   aug_img))  # Save augmented images as images without inverted color values.
+            new_file_name = definitions.PREPROCESSED_CHARSET_PATH + 'image-{}-{}.png'.format(label, i)
+            cv2.imwrite(new_file_name, aug_img)
+            out_file_names.write(new_file_name + '\n')
+            i += 1
+            # Keeping track of time...
+            if i % 250 == 0:
[…]

### 2017-12-17 src/character_recognition.py

@@ -121,22 +121,22 @@ def create_neural_net(global_weights=None, train=True, base=1, filter_size=FILTE
     with tf.variable_scope('CharacterRecognition'):
         _x = tf.placeholder(tf.float32, (None, SIZE, SIZE, NUM_CHANNELS))  # batch size - height - width - channels
         _y = tf.placeholder(tf.int64, (None, NUM_CLASSES))  # batch size - classes
-        base2 = base * 1024
-        h1 = new_conv_layer(name=1, input=_x, filter_size=filter_size, num_filters=8, num_in_channels=NUM_CHANNELS,
+        h1 = new_conv_layer(name=1, input=_x, filter_size=7, num_filters=8, num_in_channels=NUM_CHANNELS,
                             use_pooling=True, global_weights=global_weights)
-        h2 = new_conv_layer(name=2, input=h1, filter_size=filter_size, num_filters=16, num_in_channels=8,
+        h2 = new_conv_layer(name=2, input=h1, filter_size=5, num_filters=16, num_in_channels=8,
                             use_pooling=True, global_weights=global_weights)
-        h3 = new_conv_layer(name=3, input=h2, filter_size=filter_size, num_filters=24, num_in_channels=16,
+
+        h3 = new_conv_layer(name=3, input=h2, filter_size=3, num_filters=24, num_in_channels=16,
                             use_pooling=True, global_weights=global_weights)
         h4 = tf.contrib.layers.flatten(h3)
         if train:
-            h5 = new_fc_layer(name=4, input=h4, num_in=h4.shape[1], num_out=base2, global_weights=global_weights)
+            h5 = new_fc_layer(name=4, input=h4, num_in=h4.shape[1], num_out=1024, global_weights=global_weights)
             h6 = tf.nn.dropout(h5, keep_prob=keep_prob)
-            h7 = new_fc_layer(name=7, input=h6, num_in=h6.shape[1], num_out=base2 / 2, global_weights=global_weights)
+            h7 = new_fc_layer(name=7, input=h6, num_in=h6.shape[1], num_out=512, global_weights=global_weights)
             h8 = tf.nn.dropout(h7, keep_prob=keep_prob)
         else:
-            h7 = new_fc_layer(name=4, input=h4, num_in=h4.shape[1], num_out=base2, global_weights=global_weights)
-            h8 = new_fc_layer(name=7, input=h7, num_in=h7.shape[1], num_out=base2 / 2, global_weights=global_weights)
[…]

### 2017-12-17 src/experiments/experiment_results.py

@@ -110,3 +110,4 @@ def visualise_all():
     print(max_i, accuracy[max_i])
 
 
+visualise_experiment_configuration("batchsize", "1024")

### 2017-12-17 Report/paper.tex

@@
 The aim of this paper is to present a system for off-line handwriting recognition. 
-We discuss how to segment texts with contouring. Next we present our method of splitting words into characters. 
+We discuss how to segment texts with contouring. Next we present our method of segmentating words into characters. 
 Then we describe how to recognise characters with a deep neural network approach. 
@@
 After this we present our approach to seperating the characters from words in \ref{seg:segword}.
-We use vertical projections combined with postprocessing techniques including a neural network for splitpoint evaluation.
+We use vertical projections combined with postprocessing techniques including a neural network for segmentation point evaluation.
 Postprocessing steps are presented in Section \ref{seg:postproc}, we try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
@@
 \ref{sec:segtext}
+
+To be able to recognise the characters in a given text, we will first need to extract the words from the text before we can start evaluating the characters.
+This word extraction has many possible approaches, but in this paper we will focus solely on the word contouring approach.
+This method uses a heuristic algorithm to extract the contours of the words from a given text, and through postprocessing we can rearrange these extracted words into lines.
+The algorithm proposed is only to processes text images, and does not take into account images and other non-textual areas found in the text.
+Dependant on the situation, additional techniques for page segmentation, such as the Run-Length Smearing Algorithm, can be utilized.
+% TODO:: Run length Smearing Algorithm reference --> ....
+
+\subsection{binarization}
+The first preprocessing step is binarizing the text image. 
+If the image is not yet greyscaled, this is done at the start of the preprocessing phase.
+The greyscaled image is converted from greyscaled into a binary image.
+This process will allows us to easily separate the text from the background.
+
+To achieve this there are a lot of thresholding algorithms to choose from, but the one we will use throughout the paper is a combination of executing Gaussian blur in combination with OTSU's thresholding algorithm on the binary text image.
+The algorithms used are as implemented in the python opencv library. %TODO:: where do i put this????
+
+The Gaussian blur will reduce the noise in the image, and smoothen the edges of the text.
+The smoothening of the background noise will result in less noise picked up by the chosen thresholding algorithm, so it is less likely to confuse background noise with the words from the text image.
+
+The tresholding algorithm used in this paper is OTSU's thresholding algorithm.
+This algorithm will deliver adequate result for noisy text documents.
+The choise of tresholding algorithm depends largely on intended purposes of the algorithm.
+If very noisy documents will need to be processed, additional noise removal techniques and other preprocessing techniques may be required to reduce the excess noise picked up by the tresholding algorithm.
+
+\subsection{contouring}
+Our heuristic text segmentation algorithm is based on finding the contours in the text image, so we can group these contours in postprocessing stages to recover the original words from the text image.
+The contouring algorithm used is the one implemented in the opencv library for python.
+This algorithm generates a lot of unneeded contours, which we will filter through postprocessing.
+Also since words in handwritten text are not always connected into one piece, we will need to connect these contours in the postprocessing stages as well.
+
+\subsection{postprocessing}
+In the postprocessing step, we eliminate the redundant extracted contours, and reconnect the contours of words that consist of multiple parts.
+
+\subsubsection{contour removal}
+Since the contouring algorithm produces contours that are not contours of the words in the text, these contours need to be filtered out and removed.
+To accomplish this we make use of the bounding box we can draw around a contour.
+This bounding box is the smallest rectangle containing said contour, and it will allow us to easily compare contour positions in the text image.
+
+The first filter is removing the contours of which the bounding box is contained by the bounding box of another contour.
+This is allowed because the contents of this inner bounding box will be regarded when processing the contents of the outer bounding box, and so we can remove the contours belonging to the inner bounding boxes.
+
+The second filter is removing the contours that are too small to contain parts of a word.
+Since the contouring algorithm extracts the contours for any size, and because our character recognition network does not process punctuation, 
+we need to eliminate these from our results.
+
+We achieve this by taking the average height of all the found contours in the text.
+From these contours we calculate the average height, after which we remove the contours with a height lesser than the average height multiplied by a chosen parameter $minimalHeightMultiplier$.
+The optimal results were achieved when we initialize with $minimalHeightMultiplier = 2/3$, so that we retain only contours higher than the average height $ * 2/3$.
+
+\subsubsection{contour combination}
+With the excess contours removed, we can begin to reconstruct the original words from the text by combining multiple contours into one where needed.
+The algorithm used for this is a straightforward overlapping algorithm, that combines word segments based on the overlap between the bounding box of their contours.
+The parameters for this algorithm are the $alowedWordGap$ and the $allowedVerticalOverlap$ paramters.
+These parameters are initialized as $interWordSegmentDistance = $ the average width of the found words $ / 16 $ and $allowedVerticalOverlap = 1/2$ to give the best results in the our testset, however further experimentation with these parameters and the algorithm is required to find a more generalized optimal solution.
+
+\subsection{line selection}
+Now that the words are extracted from the text image, the next goal is to try and reconstruct the original text by reordering these words in their original lines.
+For the line selection process, an iterative algorithm, Algorithm \ref{alg:linereconstr}, is used.
+\begin{algorithm}[tb]
+   \caption{Line reconstruction algorithm}
+   \label{alg:linereconstr}
+\begin{algorithmic}
+   \STATE {\bfseries Input:} $extractedWords$, $textlines$
+   \REPEAT
+        \STATE $highestWord = findHighestWord $
+        \STATE $newline = [highestWord]$
+        \STATE Initialize $noChange = true$.
+        \STATE $avgY= newline$.findAverageYpos
+        \REPEAT
+            \FOR{$word$ {\bfseries in} $extractedWords$}
+            \STATE $boundRect = word.contour.boundingRect$
+            \IF{ $avgY > word.y$ {\bfseries and} $avgY < word.y + word.h$ } 
+            \STATE $line$.add $word$
+            \STATE $noChange = false$
+            \ENDIF
+        \ENDFOR
+        \UNTIL{$noChange = true$}
+        \STATE $textlines$.add $newline$
+   \UNTIL{$words$.empty $= true$}
+\end{algorithmic}
+\end{algorithm}
+
+This algorithm reorders the found words in lines, and return these lines as ordered in the original text image.
+
+This algorithm fares well processing text images, where the text consists of parrallel text lines.
+Angeled parallel text lines can be compensated for by rotating and padding the thresholded text image with different angles, and comparing the results.
+
+Combinations of parrallel and slanted text lines however are not handled well by the algorithm, and will require further adaptations or more advanced algorithms to map the words to text lines correctly.
+
 \section{Segmentation of words into characters}
 \ref{sec:segword}
-\subsection{Correction of oversegmentation}
-As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these splitting points. 
-\cite{evalsplitpoints} discusses three approaches to evaluate split point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
-In this paper we will only discuss the latter, machine learning based procedures, as we used this approach. 
-In particular, our algorithm is based on \cite{evalsplitpointsnn}.
-The paper explains how to use neural networks to classify the correctness of splitting points.
+The next step in the process is to extract the character images from the given word images.
+Again, there is the choise between several different approaches.
+In this paper we focus solely on a word oversegmentation approach to character extraction.
+The purpose is to segment the words into the individual characters.
+This will be done through initial oversegmentation of the word image, followed by postprocessing steps to remove the excess separation points to recover the original character images.
+
+\subsection{Normalization} % TODO:: necessary? never taken advantage of? maybe for split points count etc???
+The first action is normalizing the word image.
+This normalization step enables us to reason about words written in different pixel scalings the same way.
+The process of normalizing the word image consists of resizing the image to a fixed height of $100px$.
+
+\subsection{Binarization}
+Next we will again need to execute a binarization algorithm on the word image.
+Before thresholding the image, Gaussian blur is executed to reduce the noise of the word image, and to smoothen the edges of the characters.
+Afterwards we use OTSU's thresholding algorithm again to binarize the blurred word image.
+This thresholding step is essential for out skeletonization algorithm in the next step, as it requires a binarized image as input.
+
+\subsection{Skeletonization}
+The skeletonization of the thresholded word image will allow us to ignore the thickness of the lines in further processing. 
+This normalization is needed to reason about potential segmentation points between characters in the word image.
+After comparing different skeletonization algorithms, the algorithm of choise is the skeletonize function found in the scipy python library.
+This is an adaptation of the Zhang-Suen thinning algorithm, one of the most commonly used and adapted thinning algorithms.
+%TODO:: GIVE REFERENCE TO ZHANG SUEN AND SCIPY MANUAL FOR SKELETONIZATION
+
+\subsection{Vertical projection histogram}
+Now that we have the skeletonized word image, we can proceed with determining the potential segmentation points.
+We do this by taking the vertical projection histogram of this skeletonized word image.
+Before calculating the projection histogram, we first have to normalize the image to contain values $1$ for black pixels and $0$ for white pixels in the image.
+When the projection vertical projection is executed, we have the summation of every collumn of the word image.
+Because the word image is skeletonized, we can regard the collumns with a total pixel count of zero or one pixel as potential segmentation points, because connections between handwritten words consist of a single stroke, or whiitespace if the word consists of multiple parts.
+
+Now that we have our potential segmentation points, we need to decide which ones are correct segmentation points, and which ones need to be removed.
+This is done through correction of our oversegmentation.
+
+\subsection{Oversegmentation correction}
+In this section we provide a method to fix the oversegmentation of the word caused by our previous steps and to decide which segmentation points to retain.
+To achieve this we first process the potential segmentation points, to remove redundant and incorrect segmentation points.
+THe final decision to which segmentation points we retain is made by a neural network.
+
+\subsubsection{rule based}
+The main oversegmentation correction is handled by our neural network.
+However some simple solutions can be added to remove unwanted potential segmentation points throught deterministic processing.
+
+The first filter we utilize here, is one that removes the segmentation points at the start and the end of the word image.
+These segmentation points are an artifact of the used segmentation approach, and can be removed without information loss, as the alphabet does not contain letters that convert to a single pixel line after skeletonization.
+
+A second filter groups consequent potential segmentation points.
+The occurence of these consequent points is another artifact from our segmentation approach.
+From these grouped points, the collumns beloning to these x-coordinates are checked to be empty or containing a single pixel in the vertical projection histogram.
+if the group contains collumns that are empty on the histogram, only the middle segmentation point from this group is retained as potential segmentation point.
+If none of the potential segmentation points has an empty collumn on the histogram, the retained segmentation point is the middle points from the group.
+
+A third filter is that groups of consequent potential segmentation points, separated by collumns containing less pixels than a certain threshold value $maxSumThresh$, are grouped together.
+The value chosen for this threshold is 3.
+This filter remedies the segmentation of slated lines between the characters.
+
+This value for this threshold is pretty conservative, and could be further investigated for a more optimal rule based approach.
+More rule based filter could be added for more optimal processing of the potential segmentation points.
+However the focus the segmentation correction is the neural network used to remove incorrect segmenatation points.
+
+%As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these segmentation points points. 
+%\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
+%In this paper we will only discuss the latter, machine learning based procedures, as we used this approach.
+
+\subsubsection{Neural network}
+The neural network we use to judge the potential segmentation points, is based on \cite{evalsplitpointsnn}.
+The paper explains how to use neural networks to classify the correctness of segmentation points.
 The discussed algorithm is as follows.
-Manual classification of splitting points creates training data with two classes, correct and incorrect splitting points.
-For every of these splitting points a pixel matrix is extracted surrounding the splitting point. This pixel matrix is normalised in size. 
+Manual classification of segmentation points creates training data with two classes, correct and incorrect segmentation points.
+For every of these segmentation points a pixel matrix is extracted surrounding the segmentation point. This pixel matrix is normalised in size. 
 Next, small windows of equal size are extracted from the pixel matrix and for each window the density is calculated.
 This is the number of black pixels divided by the total amount of pixels in the window.
-With every matrix a corresponding label is encoded for storage in the training file, this label is encoded as 0.9 if the splitting point is correct and 0.1 otherwise.
+With every matrix a corresponding label is encoded for storage in the training file, this label is encoded as 0.9 if the segmentation point is correct and 0.1 otherwise.
 Now these matrices and corresponding labels are used to train the neural network.
@@
 We use one convolutional layer and three fully connected layers in our neural network.
-This neural network takes the actual image of the splitting point as input (a pixel matrix of the pixels surrounding the splitting point) and returns a one-hot encoded vector indicating if the splitting is correct or incorrect.
+This neural network takes the actual image of the segmentation point as input (a pixel matrix of the pixels surrounding the segmentation point) and returns a one-hot encoded vector indicating if the segmentation is correct or incorrect.
 This method allowed us to achieve a higher accuracy than an implementation of the method described by \cite{evalsplitpointsnn}.
-Splitting points can now be filtered if the neural network decides a splitting point is incorrect.
+Segmentation points can now be filtered if the neural network decides a segmentation point is incorrect.
+
+\subsection{Normalization}
+The character normalization step consists of two parts.
+Firstly the character is padded untill the height equals the width.
+A precondition for the padding step is that the image is already binarized, so that we can use white borders as padding, and this will not distort the image should binarization take place in a later stage.
+The image is then resized to $44px x 44px$
+Secondly padding of $10px$ is added on all sides of the character image to create whitespace around the letter.
+We end up with a $64px x 64px$ character image.
+The extra $10px$ padding is to make sure the letters do not stick to the edges, since the Chars74K dataset we work with utilizes borders around the provided characters.
+Should these borders be removed in the training process however, this should be reflected in this step by ignoring this new padding.
 

### 2017-12-17 Report/paper.tex
update title

--- 2017-12-13 Report/paper.tex
@@
 \begin{document} 
-\twocolumn[\icmltitle{Handwriting Recognition}
+\twocolumn[\icmltitle{An Approach to Off-line Handwriting Recognition through Segmentation Techniques and Deep Neural Networks}
 \icmlauthor{Jarre Knockaert}{Jarre.Knockaert@UGent.be}
@@
 % - Disciss possible flaws. (Use images from IAM in order to do so)
-% - Update outdated parts, like neural net of splitpoint evaluation
 %  Reference datasets, Chars74K, IAM, Google n-grams
@@
 
-\paragraph{Adding noise to data}
-We add a variant of each image which includes noise troughout the image with the purpose of reducing overfitting of the neural network \cite{DataNoise}.
-This noise is taken from the Gaussian distribution with a mean of 0 and standard deviation of 0.004.
-The noise is summed with the original image of which the pixels are normalized between $[-1, 1]$.
 \paragraph{Translations of data}
@@
 All of the previously discussed preprocessing techniques are also applied to the eroded image.
-The final dataset contains translated, orientated, sheared, scaled and noisy versions of both the original image and the eroded image.
+The final dataset contains translated, orientated, sheared and scaled versions of both the original image and the eroded image.
 A visualitation of the augmentation of an augmented picture can be seen in Figure \ref{fig:augmented}.
@@
   \caption{Original image}
-\end{subfigure}
-\begin{subfigure}{0.11\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{images/noisy}
-  \caption{Noisy image}
 \end{subfigure}
@@
 We take a slightly different approach and consider a deep neural network to tackle this problem.
-This allows us to use convolutational layers in order to extract features, instead of manually creating features (features which show the density in this case).
+This allows us to use the power of convolutational layers to extract features, instead of manually creating features (features which show the density in this case).
 We use one convolutional layer and three fully connected layers in our neural network.
@@
 This way we can avoid outliers.
-Another note is that some experiments might not use the finalized setup of our neural network. These experiments do use the same setup to compare different configurations. 
+Finally, a non augmented dataset is used to increase the speed at which experiments execute.
+These experiments do use the same setup to compare different configurations unless when mentionned otherwise.
 \subsubsection{Experiments}
 \paragraph{Preprocessing techniques} 
-Figure \ref{fig:preprocess} shows that preprocessing techniques have great influence on the performance of the neural network.
-We can see that the accuracy is much higher when using all of the preprocessing techniques.
-We achieve 72\% accuracy with all of the preprocessing techniques versus 35\% without any preprocessing techniques, which is a huge difference. 
-We can also see the individual impact of preprocessing in the figure.
-Adding shearing (64\%) to handle different amounts of slant, adding noise (58\%) to cope to make the recognition more robust and adding scaling (59\%) to handle different handwriting sizes was the most effective to increase performance.
-Adding rotations (40\%) and translations (50\%) was less effective (but still effective).
-Adding these preprocessing techniques increased the amount of data greatly from 3410 examples to 37000 examples.
-This improves the performance but also increases the training time of our network significantly.
-One remarkable result is that the training time of preprocessing with all techniques enabled, takes shorter time than some other results.
-When we look at the average however, the training time is the highest for this case.
+Figure \ref{fig:preprocess} shows that preprocessing techniques have great influence on the performance of the neural network. However these results might seem counterintuitive. 
+Adding preprocessing techniques lowers the accuracy of the neural network.
+In order to explain this, it has to be noted that every experiment uses the original data in addition with the augmented data using a preprocessing technique.
+The base case without preprocessing techniques, just uses a copy of every image as augmented dataset.
+This way every experiment uses the same amount of images.
+The accuracy decreases when using preprocessing techniques because the variety of our dataset increases but these techniques allow a more robust neural network which can recognise characters which includes images with more variation.
+
+Now we will explain the variance between several preprocessing techniques.
+All preprocessing techniques have a lower accuracy than using no preprocessing techniques.
+The amount of decrease in accuracy can be explained by how much the original image and augmented image differ.
+The configuration with erosion achieves the second highest accuracy, as these images resemble the original images most, the shape, position and rotation of the image remains, the density of regions in the image has only changed.
+A dataset with scaled images instead has much lower accuracy.
+As the image is completely different, the pixels are in other positions and the original shape has changed, the shape has a smaller or bigger surface depending on the kind of scaling.
+
+In the final dataset we use augment every image once or more with every technique which causes the dataset to significantly increase in size.
+The accuracy of the configuration with augmented images is only higher because of the increase in data which allows more weight adjustment.
+Figure \ref{fig:allprep} shows the comparison of using no preprocessing techniques and all preprocessing techniques once or more.
+
 \paragraph{Learning rate}

### 2017-12-17 Report/paper.tex

--- 2017-12-17 Report/paper.tex
@@
 % Notes:
-% - Add https://tel.archives-ouvertes.fr/tel-01249405/document reference, (see abstract) 
-% in order to compare results. 
-% - Disciss possible flaws. (Use images from IAM in order to do so)
-% - Update outdated parts, like neural net of splitpoint evaluation
-%  Reference datasets, Chars74K, IAM, Google n-grams
 %  redo the preprocessing experiment and explain why some techniques increase the accuracy more than others. 
 \begin{abstract} 
-The aim of this paper is to present a system for off-line handwriting recognition. 
+The aim of this paper is to present a complete system for off-line handwriting recognition. 
 We discuss how to segment texts with contouring. Next we present our method of segmentating words into characters. 
@@
 We conclude this paper with some experimental results, the problems we encountered combined with possible improvements.
+A complete flow of the recognition of an image is shown in Figure \ref{fig:flow}.
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=\linewidth]{images/flow_hor}
+  \caption{All of the steps in our system to convert an image with handwritten text to digitalized text.}
+  \label{fig:flow}
+\end{figure*}
 
@@
 
-\subsection{Data}
+\subsection{Dataset}
 \label{sec:data}
@@
   \includegraphics[width=\linewidth]{images/bad_char1}
-  \caption{Picture of a 'z' which resembles a 7.}
 \end{subfigure}
@@
   \includegraphics[width=\linewidth]{images/bad_char2}
-  \caption{Picture of a 'z' which resembles a 3.}
 \end{subfigure}
-\caption{Examples of written 'z' characters, which are hard to recognise.}
+\caption{Examples of written 'z' characters, which are hard to recognise or ambiguous.}
 \label{fig:char}
@@
 \begin{figure}
-\begin{subfigure}{0.11\textwidth}
+\begin{subfigure}{0.15\textwidth}
   \centering
   \includegraphics[width=\linewidth]{images/original}
-  \caption{Original image}
+  \caption{Original}
 \end{subfigure}
-\begin{subfigure}{0.11\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{images/noisy}
-  \caption{Noisy image}
-\end{subfigure}
-\begin{subfigure}{0.11\textwidth}
+\begin{subfigure}{0.15\textwidth}
   \centering
   \includegraphics[width=\linewidth]{images/translated}
-  \caption{Translated image}
+  \caption{Translated}
 \end{subfigure}
-\begin{subfigure}{0.11\textwidth}
+\begin{subfigure}{0.15\textwidth}
   \centering
   \includegraphics[width=\linewidth]{images/rotated}
-  \caption{Rotated image}
+  \caption{Rotated}
 \end{subfigure}
-\begin{subfigure}{0.11\textwidth}
+\begin{subfigure}{0.15\textwidth}
   \centering
   \includegraphics[width=\linewidth]{images/scaled}
-  \caption{Scaled image}
+  \caption{Scaled}
 \end{subfigure}
-\begin{subfigure}{0.11\textwidth}
+\begin{subfigure}{0.15\textwidth}
   \centering
   \includegraphics[width=\linewidth]{images/sheared}
-  \caption{Sheared image}
+  \caption{Sheared}
 \end{subfigure}
-\begin{subfigure}{0.11\textwidth}
+\begin{subfigure}{0.15\textwidth}
   \centering
   \includegraphics[width=\linewidth]{images/eroded}
-  \caption{Eroded image}
+  \caption{Eroded}
 \end{subfigure}
-\caption{All of the augmentations applied to one image.}
+\caption{An image transformed by the discussed augmentation techniques.}
 \label{fig:augmented}
@@
 Normalized images can be fed to the neural network, which then propagate through several layers and eventually produce as output the probability that the input image has a certain class, one of the possible characters.
-Now follows a detailed overview of our deep neural network. A detailed explanation on how deep neural networks function is explained by \cite{nnbook}.
+Now follows a detailed overview of our deep neural network.
+This overview is visualized in Figure \ref{fig:dnn}.
+A detailed explanation on how deep neural networks function is explained by \cite{nnbook}.
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{images/dnn}
+  \caption{Visualization of the deep neural network for character recognition.
+The first row shows the shape a layer.
+The second row shows the visualization of that layer, the third row shows information about the layer.
+'C' indicates a convolutional layer, 'MP' a max-pooling layer and 'FC' a fully-connected layer.}
+  \label{fig:dnn}
+\end{figure*}
 
@@
 
+The Adam optimizer minimizes the error calculated by the cross-entropy cost function adjusted with a weight decay factor.
+Weight decay regulizes the neural network by penalizing big weights and promoting smaller weights.
+Neural networks with this constraint tend to overfit less because small weights are less sensible to small changes of the input.
+\cite{presham}
 \section{Segmentation of text into words}
@@
 We take a slightly different approach and consider a deep neural network to tackle this problem.
-This allows us to use convolutational layers in order to extract features, instead of manually creating features (features which show the density in this case).
+This allows us to use the power of convolutational layers to extract features, instead of manually creating features (features which show the density in this case).
 We use one convolutional layer and three fully connected layers in our neural network.
-This neural network takes the actual image of the segmentation point as input (a pixel matrix of the pixels surrounding the segmentation point) and returns a one-hot encoded vector indicating if the segmentation is correct or incorrect.
+This neural network takes the actual image of the segmentation point as input (a pixel matrix of the pixels surrounding the splitting point) and returns a one-hot encoded vector indicating if the splitting is correct or incorrect.
 This method allowed us to achieve a higher accuracy than an implementation of the method described by \cite{evalsplitpointsnn}.
@@
 An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+n}$.
-\cite{ngrams} % TODO google citation%
+\cite{ngrams}
 We use the n-grams from the Google dataset for this purpose.
@@
 \subsection{Character Recognition}
-\label{sec:expres}
-We tested a lot of different configurations for the deep neural network in order to find a good functioning neural network given the low amounts of data.
-A problem which arrised when trying to test many different configurations was the long time it takes to train a neural network.
+The deep neural network can recognise characters with an accuracy of 91 \%.
+This neural network was trained for 1000 iterations which took 8650 seconds.
+We used a learning rate of $10^{-4}$, 256 as batch size, dropout layers with a keep probability of 0.5 and $10^{-3}$ as weight decay.
+Experimental findings led us to these values.
+
+Comparisons are hard to find as few papers discuss off-line handwritten character recognition.
+One paper we found, \cite{charrec}, achieves an accuracy of 98.5 \%, they classify 38 classes (10 digits, 26 letters, 2 special symbols) and manually extract features for the neural network instead of using convolutional layers.
+
+\subsection{Influence of data augmentation on character recognition}
+
+We're using the following setup for this experiment.
+Every experiment uses the original data in addition with the augmented data using one particular preprocessing technique.
+The base case without preprocessing techniques, just uses a copy of every image as augmented dataset.
+
 In order to limit the time these experiments run we added some small constraints. We only run every experiment 200 epochs, running these experiments longer might result in different conclusions. 
 Next it is important to note that a lot of stochastic behaviour is part of these neural networks because of the dropout layers.
-When running the neural network twice with the same configuration, it could have completely different results.
-This is why we run every configuration four times.
-Only the result with the highest accuracy is considered.
-This way we can avoid outliers.
-Another note is that some experiments might not use the finalized setup of our neural network. These experiments do use the same setup to compare different configurations. 
-\subsubsection{Experiments}
-\paragraph{Preprocessing techniques} 
-Figure \ref{fig:preprocess} shows that preprocessing techniques have great influence on the performance of the neural network.
-We can see that the accuracy is much higher when using all of the preprocessing techniques.
-We achieve 72\% accuracy with all of the preprocessing techniques versus 35\% without any preprocessing techniques, which is a huge difference. 
-We can also see the individual impact of preprocessing in the figure.
-Adding shearing (64\%) to handle different amounts of slant, adding noise (58\%) to cope to make the recognition more robust and adding scaling (59\%) to handle different handwriting sizes was the most effective to increase performance.
-Adding rotations (40\%) and translations (50\%) was less effective (but still effective).
-Adding these preprocessing techniques increased the amount of data greatly from 3410 examples to 37000 examples.
-This improves the performance but also increases the training time of our network significantly.
-One remarkable result is that the training time of preprocessing with all techniques enabled, takes shorter time than some other results.
-When we look at the average however, the training time is the highest for this case.
-\paragraph{Learning rate}
-Figure \ref{fig:learningrate} shows the impact of learning rate on the training of our neural network. 
-For this experiment we do have to note that we are using the Adam optimizer to train our neural network which computes individual learning rates for every parameter every training step.
-The learning rate which we pass is the initial learning rate.
-Choosing a high learning rate $10^{-3}$ allows the neural network to converge to a high accuracy (76 \%) very fast, but once reached the accuracy fails to grow any more.
-This in contrast to higher learning rates which reach a good accuracy more slowly but find a more optimal solution.
-Learning rate $10^{-4}$ reaches an accuracy of 68\% after 200 iterations but is still growing.
-Learning rate $10^{-5}$ only reaches an accuracy of 5\% after 200 iterations but should find a better solution eventually than the other two configurations.
-\paragraph{Batch size}
-n Figure \ref{fig:batchsize} we can see the impact of choosing different batch sizes which are fed to the neural network.
-In order to increase the speed at which the neural network is trained, we feed batches of input images to the neural network instead of one image at a time.
-Weight adjustment only occurs after feeding a complete batch to the network.
-In the figure we can see bigger batch sizes increase the execution speed.
-However these differences are quite small.
-The configuration with batch size executes 200 iterations in one minute less than a configuration with batch size 128.
-These batch sizes also have an effect on our performance.
-When using small batch sizes backpropagation happens more frequently, which implies our weights will be more finetuned.
-The configuration with batch size 128 obtains an accuracy of 10\% higher than batch size 1024.
-\paragraph{Weight decay}
-In order to regulize the neural network we add weight decay to the network loss function in order to penalize big weights.
-The impact of different decays can be seen in Figure \ref{fig:decay}.
-A high decay functions best (71\% accuracy) because we have a small amount of training examples.
-Without decay we only reach an accuracy of 65\%.
-\paragraph{Dropout}
-% Maybe redo this experiment?
-It's not clear at all at this time whether dropout has an actual effect on the accuracy.
-I believe it helps find a better global solution, but often the accuracy will be lower as we're only using half of the weights.
-Further research is required.
+We run every configuration four times.
+Only the result with the highest accuracy is considered to avoid outliers.
+The accuracy decreases when using preprocessing techniques because the variety of our dataset increases but these techniques allow a more robust neural network which can recognise characters which includes images with more variation.
+
+Figure \ref{fig:preprocess} shows that augmentation techniques have great influence on the performance of the neural network. However these results might seem counterintuitive. 
+All preprocessing techniques have a lower accuracy than the base case which does not use preprocessing techniques.
+The amount of decrease in accuracy can be explained by how much the original image and augmented image differ.
+The configuration with erosion achieves the second highest accuracy, as these images resemble the original images most, the shape, position and rotation of the image remains, the density of regions in the image has only changed.
+A dataset with scaled images instead has much lower accuracy.
+As the image is completely different, the pixels are in other positions and the original shape has changed, the shape has a smaller or bigger surface depending on the kind of scaling.
+
+In the final dataset we use augment every image once or more with every technique which causes the dataset to significantly increase in size.
+The accuracy of the configuration with augmented images is only higher because of the increase in data which allows more weight adjustment.
+Figure \ref{fig:allprep} shows the comparison of using no preprocessing techniques and all preprocessing techniques once or more.
+
 \begin{figure}
@@
 \end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{../Graphs/learningrate.png}
-    \caption{Impact of different learning rates.}
-    \label{fig:learningrate}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{../Graphs/batchsize.png}
-    \caption{Impact of different batch sizes.}
-    \label{fig:batchsize}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{../Graphs/decay.png}
-    \caption{Impact of weight decay regularisation.}
-    \label{fig:decay}
-\end{figure}
-\subsubsection{Final results}
-% Add graph with several runs of the final network. Discuss which parameters we took. 
-% Maybe some more visualisation of mispredicted characters? Oh this isn't part of char recognition...
-% TODO TODO
+
 \section{Conclusion}
@@
 \section{Future research}
-
+State-of-the-art systems currently use segmentation-free approaches, this involves handwriting recognition without explicitly splitting words into characters.
+These methods are mostly based on hidden Markov models and neural networks.
+This concept is explained and applied by \cite{presham}.
+\section{Acknowledgements}
+We thank Sarah Vluymans for her comments on this paper.
 \bibliography{paper}

### 2017-12-17 Report/paper.tex
working on paper

--- 2017-12-17 Report/paper.tex
@@
 This is an adaptation of the Zhang-Suen thinning algorithm, one of the most commonly used and adapted thinning algorithms.
-%TODO:: GIVE REFERENCE TO ZHANG SUEN AND SCIPY MANUAL FOR SKELETONIZATION
+\cite{zsthinning}
 
@@
 We do this by taking the vertical projection histogram of this skeletonized word image.
-Before calculating the projection histogram, we first have to normalize the image to contain values $1$ for black pixels and $0$ for white pixels in the image.
-When the projection vertical projection is executed, we have the summation of every collumn of the word image.
-Because the word image is skeletonized, we can regard the collumns with a total pixel count of zero or one pixel as potential segmentation points, because connections between handwritten words consist of a single stroke, or whiitespace if the word consists of multiple parts.
-
-Now that we have our potential segmentation points, we need to decide which ones are correct segmentation points, and which ones need to be removed.
-This is done through correction of our oversegmentation.
+With the vertical projection executed, we now have the summation of every collumn of the word image.
+This is done in a way that black pixels are represented by a $1$ value, and white pixels are represented by a $0$ value.
+Now we can regard the collumns with a total pixel count of zero or one pixel as potential segmentation points, as we can interpret columns containing only a single pixel to represent a connection between different characters, and collumns without pixels to represent spacing between characters.
+With these points as our potential segmentation points, we need to separate the correct segmentation points from the incorrect ones.
+For this we will use oversegmentation correction techniques.
 
 \subsection{Oversegmentation correction}
-In this section we provide a method to fix the oversegmentation of the word caused by our previous steps and to decide which segmentation points to retain.
-To achieve this we first process the potential segmentation points, to remove redundant and incorrect segmentation points.
-THe final decision to which segmentation points we retain is made by a neural network.
+In this section we provide a method to fix the oversegmentation of the word image caused by the previous steps.
+To achieve this we process the potential segmentation points with a few rule based filters, to remove redundant and incorrect segmentation points.
+The final decision to which segmentation points we retain is made by a neural network.
 

### 2017-12-17 Report/paper.tex
fix most references

--- 2017-12-17 Report/paper.tex
@@
 In Section \ref{sec:segtext} we describe how to segment a text into words using a contouring algorithm with consequent postprocessing to remove unwanted noise. 
-After this we present our approach to seperating the characters from words in \ref{seg:segword}.
+After this we present our approach to seperating the characters from words in \ref{sec:segword}.
 We use vertical projections combined with postprocessing techniques including a neural network for segmentation point evaluation.
-Postprocessing steps are presented in Section \ref{seg:postproc}, we try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
+Postprocessing steps are presented in Section \ref{sec:postproc}, we try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
 We conclude this paper with some experimental results, the problems we encountered combined with possible improvements.
@@
 As our dataset only contains 3410 images with handwritten characters, we need more data for the neural network to be able to function properly.
-Increasing the size of our dataset has great impact on the performance of the neural network as discussed in Section \ref{sec:expres}.
-Now we will describe the data augmentation techniques we used.
-For these techniques we based us on \cite{DataAug}.
-A more thorough explanation can be found there.
+Next we will describe the data augmentation techniques we used. For these techniques we based us on \cite{DataAug}. A more thorough explanation can be found there.
 
@@
 \cite{presham}
+
 \section{Segmentation of text into words}
-\ref{sec:segtext}
+\label{sec:segtext}
 
@@
 \section{Segmentation of words into characters}
-\ref{sec:segword}
+\label{sec:segword}
+
 The next step in the process is to extract the character images from the given word images.
@@
 \subsection{Natural language processing}
-To cope with some of the high error rates as shown in Section \ref{sec:expres}, there needs to be some kind of correction.
-This correction should allow the system to reproduce possibble written text even when a step of the larger system fails.
+\label{sec:postproc}
+
+To cope with some of the errors from the previous steps, there needs to be some kind of correction.
+This correction should allow the system to reproduce possible written text even when a step of the larger system fails.
 In order to do this we adapt natural language processing techniques as postprocessing steps.
@@
 \section{Experimental results}
-%TODO
+\label{sec:expres}
+
 \subsection{Character Recognition}

### 2017-12-17 src/word_normalizer.py
nearly there

@@ -13,6 +13,9 @@ def normalize_word(word_image):
     difference_percentage = NORMALIZED_HEIGHT / height
     normalized_width = width * difference_percentage
 
+    if (normalized_width < 3):
+        return word_image
+
     normalized_word = cv2.resize(word_image, (int(normalized_width), NORMALIZED_HEIGHT))
 
     #norm_height, norm_width = normalized_word.shape[:2]
@@ -21,4 +24,3 @@ def normalize_word(word_image):
     #container[PADDING:norm_height+PADDING ,PADDING:norm_width+PADDING] = normalized_word
 
     return normalized_word
-

### 2017-12-17 src/experiments/experiment_results.py
redo all the experiments for the 100th time :s

@@ -110,4 +110,5 @@ def visualise_all():
     print(max_i, accuracy[max_i])
 
 
-visualise_experiment_configuration("batchsize", "1024")
+visualise_experiment("learningrate")
+# visualise_experiment_configuration("batchsize", "1024")

### 2017-12-17 src/utils/experiments.py

@@ -221,9 +221,9 @@ def run_experiment(name, iterations=ITERATIONS, start=0, n=EPOCHS, learning_rate
 
 # character_preprocessing.augment_data(confirm=False)
 # character_recognition.train_net(1000)
-# character_preprocessing.reset_data()
-compare_batch_size()
-compare_learning_rate()
-compare_keep_prob()
-compare_weight_decay()
+# -character_preprocessing.reset_data()
+#compare_batch_size()
+#compare_learning_rate()
+#compare_keep_prob()
+#compare_weight_decay()
 compare_preprocessing_params()

### 2017-12-17 Report/paper.tex

--- 2017-12-17 Report/paper.tex
@@
 Then we describe how to recognise characters with a deep neural network approach. 
-Finally we present some postprocessing techniques to correct mispredicted text. 
+Finally we presentf postprocessing techniques to correct mispredicted text. 
 \end{abstract}
@@
 
+\begin{figure}
+    \begin{subfigure}{\linewidth}
+    \centering
+    \includegraphics[width=.7\linewidth]{images/vpp}
+    \vspace{-5px}
+    \caption*{Vertical projection profile.}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+    \centering
+    \includegraphics[width=.7\linewidth]{images/rules}
+    \vspace{-5px}
+    \caption*{Rule based correction.}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+    \centering
+    \includegraphics[width=.7\linewidth]{images/nn}
+    \vspace{-5px}
+    \caption*{Correction through neural network.}
+    \end{subfigure}
+    \caption{Word segmentation steps}
+    \label{fig:char_segm}
+\end{figure}
+
 \subsection{Oversegmentation correction}
@@
 \section{Conclusion}
+There are a lot of steps in this paper from which we can draw conclusions.
+The segregation of a text image in word images is  
+Word extraction is a problem
+
+Character segmentation is another ..
+The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect potential segmentation points the neural network has to process.
+
+The neural network
 
@@
 These methods are mostly based on hidden Markov models and neural networks.
+This approach does not suffer from character segmentation artifacts
 This concept is explained and applied by \cite{presham}.

### 2017-12-17 Report/paper.tex
add table instead of figure

@@
 \usepackage{graphicx} 
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 \usepackage{algorithm}
@@
 \usepackage{titlesec}
+\usepackage{csvsimple}
 
@@
 the exact movements of the pen.
-\cite{olcr} present a technology for online recognition of Japanese handwritten text with an accuracy of 94.6\%.
+\citep{olcr} present a technology for online recognition of Japanese handwritten text with an accuracy of 94.6\%.
 
@@
 Theoretically feeding normalized data to the neural network returns the same output as before. 
-In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}.
+In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \citep{NormGoal}.
 We normalize the data by reducing the input dimensions from (1200,900,3) to (64,64,1). A simple threshold is added to improve the contrast in our image. 
@@
 As our dataset only contains 3410 images with handwritten characters, we need more data for the neural network to be able to function properly.
-Next we will describe the data augmentation techniques we used. For these techniques we based us on \cite{DataAug}. A more thorough explanation can be found there.
+Next we will describe the data augmentation techniques we used. For these techniques we based us on \citep{DataAug}. A more thorough explanation can be found there.
 
@@
 In order to deal with different kinds of slants we add sheared versions of the image to the dataset. 
-Shearing displaces each point in a fixed direction, by an amount proportional to its signed distance from a line that is parallel to that direction \cite{Shear}.
+Shearing displaces each point in a fixed direction, by an amount proportional to its signed distance from a line that is parallel to that direction \citep{Shear}.
 We used the following shear mapping to achieve this:
@@
 This overview is visualized in Figure \ref{fig:dnn}.
-A detailed explanation on how deep neural networks function is explained by \cite{nnbook}.
+A detailed explanation on how deep neural networks function is explained by \citep{nnbook}.
 
@@
 Dropout greatly reduces overfitting in the fully connected layers.
-\cite{dropout} We do not use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
+\citep{dropout} We do not use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
 The last fully connected layer has 62 output neurons, which correspond to the 62 possible output classes.
@@
 Each of these methods effectively improve the previously mentionned method. 
-The Adam optimizer, introduced by \cite{adam}, is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
+The Adam optimizer, introduced by \citep{adam}, is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
 This optimizer achieves the highest accuracy in our case and will be used.
@@
 Neural networks with this constraint tend to overfit less because small weights are less sensible to small changes of the input.
-\cite{presham}
+\citep{presham}
 
@@
 This is an adaptation of the Zhang-Suen thinning algorithm, one of the most commonly used and adapted thinning algorithms.
-\cite{zsthinning}
+\citep{zsthinning}
 
@@
 %As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these segmentation points points. 
-%\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
+%\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
 %In this paper we will only discuss the latter, machine learning based procedures, as we used this approach.
@@
 \subsubsection{Neural network}
-The neural network we use to judge the potential segmentation points, is based on \cite{evalsplitpointsnn}.
+The neural network we use to judge the potential segmentation points, is based on \citep{evalsplitpointsnn}.
 The paper explains how to use neural networks to classify the correctness of segmentation points.
@@
 This neural network takes the actual image of the segmentation point as input (a pixel matrix of the pixels surrounding the splitting point) and returns a one-hot encoded vector indicating if the splitting is correct or incorrect.
-This method allowed us to achieve a higher accuracy than an implementation of the method described by \cite{evalsplitpointsnn}.
+In order to create training data we created a small program to iterate over splitting points of handwritten words from the IAM dataset.
+\citep{iam}
+
+This method allowed us to achieve an accuracy of 72 \%, which is still fairly low because of the difficulty of the problem and the small amount of traing data.
+However this is better than our accuracy with an implementation of the method described by \citep{evalsplitpointsnn}.
 Segmentation points can now be filtered if the neural network decides a segmentation point is incorrect.
@@
 The probability is computed as $P(w_i | w_{i-n},...,w_{i-1})$.
-We use the Markov assumption here: the probability of a word in the context can be approximated by the previous $n-1$ words. \cite{markov} This probability is calculated as follows: 
+We use the Markov assumption here: the probability of a word in the context can be approximated by the previous $n-1$ words. \citep{markov} This probability is calculated as follows: 
 \begin{equation}
@@
 An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+n}$.
-\cite{ngrams}
+\citep{ngrams}
 We use the n-grams from the Google dataset for this purpose.
-\cite{google} Querying these datasets locally takes several minutes and this happens for every word.
+\citep{google} Querying these datasets locally takes several minutes and this happens for every word.
 Instead we use a Python library phrasefinder to make online queries to these corpora.
@@
 Comparisons are hard to find as few papers discuss off-line handwritten character recognition.
-One paper we found, \cite{charrec}, achieves an accuracy of 98.5 \%, they classify 38 classes (10 digits, 26 letters, 2 special symbols) and manually extract features for the neural network instead of using convolutional layers.
+One paper we found, \citep{charrec}, achieves an accuracy of 98.5 \%, they classify 38 classes (10 digits, 26 letters, 2 special symbols) and manually extract features for the neural network instead of using convolutional layers.
 
@@
 
-Figure \ref{fig:preprocess} shows that augmentation techniques have great influence on the performance of the neural network. However these results might seem counterintuitive. 
-All preprocessing techniques have a lower accuracy than the base case which does not use preprocessing techniques.
-The amount of decrease in accuracy can be explained by how much the original image and augmented image differ.
-The configuration with erosion achieves the second highest accuracy, as these images resemble the original images most, the shape, position and rotation of the image remains, the density of regions in the image has only changed.
-A dataset with scaled images instead has much lower accuracy.
-As the image is completely different, the pixels are in other positions and the original shape has changed, the shape has a smaller or bigger surface depending on the kind of scaling.
+Table \ref{tab:preprocess} shows that augmentation techniques have great influence on the performance of the neural network. However these results might seem counterintuitive. 
+All preprocessing techniques have a lower accuracy than the base case.
+The amount of decrease in accuracy can be explained by how much the original image and the augmented image differ.
+The configuration with erosion achieves the second highest accuracy, as these images resemble the original images, the shape, position and rotation of the image remains, only the density of regions in the image has changed.
+A dataset with rotated images instead has much lower accuracy.
+As the input pixel matrix is completely different, the pixels are in other positions and the original shape has changed.
 
-In the final dataset we use augment every image once or more with every technique which causes the dataset to significantly increase in size.
-The accuracy of the configuration with augmented images is only higher because of the increase in data which allows more weight adjustment.
-Figure \ref{fig:allprep} shows the comparison of using no preprocessing techniques and all preprocessing techniques once or more.
-
-\begin{figure}
-\centering
-    \includegraphics[width=\linewidth]{../Graphs/preprocess.png}
-    \caption{Impact of various preprocessing techniques.}
-    \label{fig:preprocess}
-\end{figure}
+\begin{table}
+\caption{Influence of augmentation techniques on performance of the neural network.}
+\label{tab:preprocess}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{lcc} 
+\hline 
+\abovespace\belowspace 
+Configuration & Accuracy (\%) & Time (s) \\ 
+\hline 
+\abovespace
+No preprocessing & 88.52 & 236 \\
+Erosion & 80.95 & 410 \\
+Scaling & 66.10 & 271 \\
+Rotation & 64.47 & 228 \\
+Shearing & 74.36 & 367 \\
+Translation & 70.47 & 306 \\
+\hline 
+\end{tabular} 
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in 
+\end{table}
 
@@
 This approach does not suffer from character segmentation artifacts
-This concept is explained and applied by \cite{presham}.
+This concept is explained and applied by \citep{presham}.
 \section{Acknowledgements}

### 2017-12-17 src/definitions.py

@@ -31,3 +31,4 @@ EXPERIMENTS_CHAR_GRAPHS_PATH = PROJECT_PATH + 'Graphs/'
 WORDSET_INFO_PATH = TXT_PATH + 'words.txt'
 WORDSET_PATH = DATA_PATH + 'words/'
 WORD_SPLITTING_PATH = DATA_PATH + 'segmentation/' + 'segmentation.txt'
+CSV_PATH = PROJECT_PATH + 'Report/csv/'

### 2017-12-17 src/experiments/experiment_results.py

@@ -5,6 +5,7 @@ import numpy as np
 from pylab import savefig
 
 from src import definitions
+import csv
 
 # from CharacterRecognition.experiments import ITERATIONS
 experiments = ['base', 'batchsize', 'decay', 'filtersize', 'keepprob', 'learningrate', 'preprocess']
@@ -94,6 +95,19 @@ def visualise_experiment_configuration(experiment_name, conf, save=False):
     visualise(accuracies, times, save=save, name='_'.join([experiment_name, conf]))
 
 
+def experiment_to_csv(experiment_name):
+    confs, accuracies, times = get_experiment_results(experiment_name)
+    out_file = open(definitions.CSV_PATH + experiment_name + '.csv', 'w')
+    writer = csv.writer(out_file, delimiter =',')
+    writer.writerow(["Configuration", "Maximum accuracy", "Execution time"])
+    for (i, conf) in enumerate(confs):
+        accuracy = 100*(np.max(accuracies[i]))
+        time = times[i][-1]
+        print("{} & {:.2f}\% & {:.0f} \\\\".format(conf, accuracy, time))
+        writer.writerow([conf, accuracy, time])
+    out_file.close()
+
+
 #for exp in experiments:
     #visualise_experiment(exp, max=True, save=True)
 #visualise_experiment_configuration("preprocess", conf="all", save=False)
@@ -110,3 +124,4 @@ def visualise_all():
     print(max_i, accuracy[max_i])
 
 
+experiment_to_csv("preprocess")

### 2017-12-17 Report/paper.tex
mhm

--- 2017-12-17 Report/paper.tex
@@
 More rule based filter could be added for more optimal processing of the potential segmentation points.
-However the focus the segmentation correction is the neural network used to remove incorrect segmenatation points.
 
-%As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these segmentation points points. 
-%\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
-%In this paper we will only discuss the latter, machine learning based procedures, as we used this approach.
+A technique was tested using the feedback of the chracter recognition network to remove incorrect segmentation points.
+However this turned out not to be an effective technique, because the neural network tends to provide random probabilities with a high degree of certainty when multiple characters were fed to the network in the same image.
+Since it is impossible to validate these responses from the neural network, this technique is not used in the final version of our oversegmentation correction.
 
@@
 Character segmentation is another ..
-The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect potential segmentation points the neural network has to process.
+The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect segmentation points the neural network has to process.
+Combinations with other approaches could also yield better results
 
@@
 These methods are mostly based on hidden Markov models and neural networks.
-This approach does not suffer from character segmentation artifacts
+Since this approach does not require the segmentation of a word image into character images, it allows us to focus more on optimization of the neural network without being hindered by incorrectly segmented characters.
 This concept is explained and applied by \cite{presham}.

### 2017-12-17 Report/paper.tex
start of conclusion

--- 2017-12-17 Report/paper.tex
@@
 
-\subsection{Natural language processing}
+\section{Natural language processing}
 \label{sec:postproc}
@@
 
-\subsubsection{Vocabularium}
+\subsection{Vocabularium}
 \label{sec:voc}
@@
 This finalises the logic for recognising handwritten words. Next we will discuss a postprocessing technique which is useful when the word is part of a bigger context.
-\subsubsection{Language model}
+\subsection{Language model}
 \label{sec:lm}
@@
 
-\section{Conclusion}
+\section{Future work}
+Our recognition system cannot be used for real-world applications.Every individual technique has a certain error, which increases the difficulty of the next step, e.g.
+if word segmentation contains mistakes, character recognition becomes much harder.
+This results in a system which cannot correctly recognize handwritten text. 
+However many improvements are still possible.
+
+Firstly, given a dataset with a higher amount of training samples, the deep neural network will be able to better recognize new characters.
+The same applies for the neural network of oversegmentation correction.
+
+Next %TODO Ruben
+
+Finally, other approaches can also be considered. 
+State-of-the-art systems currently use segmentation-free approaches, this involves handwriting recognition without explicitly splitting words into characters.
+These methods are mostly based on hidden Markov models and neural networks.
+This concept is explained and applied by \citep{presham}.
+%%% Dit is jouw oude tekst %%% 
 There are a lot of steps in this paper from which we can draw conclusions.
@@
 The neural network 
-
-\section{Future research}
-State-of-the-art systems currently use segmentation-free approaches, this involves handwriting recognition without explicitly splitting words into characters.
-These methods are mostly based on hidden Markov models and neural networks.
-This approach does not suffer from character segmentation artifacts
-This concept is explained and applied by \citep{presham}.
-\section{Acknowledgements}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Acknowledgements}
 We thank Sarah Vluymans for her comments on this paper.

### 2017-12-17 Report/paper.tex

@@
 More rule based filter could be added for more optimal processing of the potential segmentation points.
-However the focus the segmentation correction is the neural network used to remove incorrect segmenatation points.
+
+A technique was tested using the feedback of the chracter recognition network to remove incorrect segmentation points.
+However this turned out not to be an effective technique, because the neural network tends to provide random probabilities with a high degree of certainty when multiple characters were fed to the network in the same image.
+Since it is impossible to validate these responses from the neural network, this technique is not used in the final version of our oversegmentation correction.
 

### 2017-12-17 Report/paper.tex
removed contouring from word extraction file, and added image in paper, with changed intro for char extraciton oversegmentation correction

@@
  \subsubsection{Weight adjustment}
-Finally, in order to actually train the neural network, we need some kind of weight adjustment.
-This will effectively improve the decision making of perceptrons and thus, the classification of images.
+Finally, in order to actually train the neural network, we need some kind of weight adjustment. This will effectively improve the decision making of perceptrons and thus, the classification of images.
 The Adam optimizer effectively fulfills this goal. In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow.
@@
 
+\begin{figure}
+    \begin{subfigure}{\linewidth}
+    \centering
+    \includegraphics[width=.7\linewidth]{images/text}
+    \vspace{-5px}
+    \caption*{Original text image.}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+    \centering
+    \includegraphics[width=.7\linewidth]{images/cont}
+    \vspace{-5px}
+    \caption*{Contouring algorithm.}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+    \centering
+    \includegraphics[width=.7\linewidth]{images/lines}
+    \vspace{-5px}
+    \caption*{Postprocessing and line selection.}
+    \end{subfigure}
+    \vspace{-10px}
+    \caption{Text segmentation steps}
+    \label{fig:text_segm}
+\end{figure}
+
+\vspace{-20px}
 \subsection{line selection}
@@
 \subsection{Oversegmentation correction}
-In this section we provide a method to fix the oversegmentation of the word image caused by the previous steps.
-To achieve this we process the potential segmentation points with a few rule based filters, to remove redundant and incorrect segmentation points.
-The final decision to which segmentation points we retain is made by a neural network.
+In this section we provide echniques to verify the correctness of these segmentation points points. 
+\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
+[6:39]
+In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
 
@@
 More rule based filter could be added for more optimal processing of the potential segmentation points.
+However the focus the segmentation correction is the neural network used to remove incorrect segmenatation points.
 
@@
 Since it is impossible to validate these responses from the neural network, this technique is not used in the final version of our oversegmentation correction.
-
-%As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these segmentation points points. 
-%\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
-%In this paper we will only discuss the latter, machine learning based procedures, as we used this approach.
 
@@
 
-\section{Natural language processing}
+\subsection{Natural language processing}
 \label{sec:postproc}
@@
 
-\subsection{Vocabularium}
+\subsubsection{Vocabularium}
 \label{sec:voc}
@@
 This finalises the logic for recognising handwritten words. Next we will discuss a postprocessing technique which is useful when the word is part of a bigger context.
-\subsection{Language model}
+\subsubsection{Language model}
 \label{sec:lm}
@@
 
-\section{Future work}
-Our recognition system cannot be used for real-world applications.Every individual technique has a certain error, which increases the difficulty of the next step, e.g.
-if word segmentation contains mistakes, character recognition becomes much harder.
-This results in a system which cannot correctly recognize handwritten text. 
-However many improvements are still possible.
-
-Firstly, given a dataset with a higher amount of training samples, the deep neural network will be able to better recognize new characters.
-The same applies for the neural network of oversegmentation correction.
-
-Next %TODO Ruben
-
-Finally, other approaches can also be considered. 
-State-of-the-art systems currently use segmentation-free approaches, this involves handwriting recognition without explicitly splitting words into characters.
-These methods are mostly based on hidden Markov models and neural networks.
-This concept is explained and applied by \citep{presham}.
-%%% Dit is jouw oude tekst %%% 
+\section{Conclusion}
 There are a lot of steps in this paper from which we can draw conclusions.
@@
 The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect segmentation points the neural network has to process.
-Combinations with other approaches could also yield better results
+Combinations with other approaches are also more likely provide better results.
 
-The neural network 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Acknowledgements}
+The neural network ..
+
+Vocabularium ..
+
+\section{Future research}
+State-of-the-art systems currently use segmentation-free approaches, this involves handwriting recognition without explicitly splitting words into characters.
+These methods are mostly based on hidden Markov models and neural networks.
+Since this approach does not require the segmentation of a word image into character images, it allows us to focus more on optimization of the neural network without being hindered by incorrectly segmented characters.
+This concept is explained and applied by \cite{presham}.
+\section{Acknowledgements}
 We thank Sarah Vluymans for her comments on this paper.

### 2017-12-17 src/word_extraction.py

@@ -453,7 +453,7 @@ def preprocess_image(img, file_index = 0):
             (x, y, w, h) = rectangle
             cv2.rectangle(new_img,(x,y),(x+w,y+h),(greyscale,255,0),5)
         greyscale += color_increment
-        cv2.drawContours(new_img, [rectangles_contours[rect] for rect in rectangles_contours], 0, (120,255,0), 3)
+        #cv2.drawContours(new_img, [rectangles_contours[rect] for rect in rectangles_contours], 0, (120,255,0), 3)
 
     # Write threshold image for demonstration pirposes
     write_threshold_image(new_img, file_index)

### 2017-12-17 Report/paper.tex

--- 2017-12-17 Report/paper.tex
@@
 \subsection{Oversegmentation correction}
-In this section we provide a method to fix the oversegmentation of the word image caused by the previous steps.
-To achieve this we process the potential segmentation points with a few rule based filters, to remove redundant and incorrect segmentation points.
-The final decision to which segmentation points we retain is made by a neural network.
+In this section we provide techniques to verify the correctness of these segmentation points. 
+\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
+[6:39]
+In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
 
@@
 The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect segmentation points the neural network has to process.
-Combinations with other approaches could also yield better results
+Combinations with other approaches are also more likely provide better results.
+
+The neural network ..
 

### 2017-12-17 Report/paper.tex
laatste paragraafje

@@
 The algorithm proposed is only to processes text images, and does not take into account images and other non-textual areas found in the text.
-Dependant on the situation, additional techniques for page segmentation, such as the Run-Length Smearing Algorithm, can be utilized.
+Dependant on the situation, additional techniques for page segmentation can be utilized. % , such as the Run-Length Smearing Algorithm, can be utilized.
 % TODO:: Run length Smearing Algorithm reference --> ....
@@
 
-Next %TODO Ruben
+The text to word segmentation process can be improved upon and made more robust. 
+Advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements.
+Algorithms that provide better noise reduction can be added to handle text images with more noise, 
+and experiments with rotations of the text can provide us with a more robust system to tackle text images containing heavily slanted text lines.
+
+The character segmentation process is inadequate as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
+More postprocessing can be used through rule based filters and a more robust neural network to correct the found segmentation points, but this will never reach an optimal solution through rule based filters alone.
 

### 2017-12-17 Report/paper.tex
improve abstract

--- 2017-12-17 Report/paper.tex
@@
 Then we describe how to recognise characters with a deep neural network approach. 
-Finally we presentf postprocessing techniques to correct mispredicted text.
+Finally we present postprocessing techniques to correct mispredicted text. 
+Our approach does not outperform state-of-the-art systems but hopes to provide a thorough elaboration for every step in handwriting recognition. 
 \end{abstract}
@@
  \subsubsection{Weight adjustment}
-Finally, in order to actually train the neural network, we need some kind of weight adjustment. This will effectively improve the decision making of perceptrons and thus, the classification of images.
+Finally, in order to actually train the neural network, we need some kind of weight adjustment.
+This will effectively improve the decision making of perceptrons and thus, the classification of images.
 The Adam optimizer effectively fulfills this goal. In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow. 
@@
 
-\begin{figure}
-    \begin{subfigure}{\linewidth}
-    \centering
-    \includegraphics[width=.7\linewidth]{images/text}
-    \vspace{-5px}
-    \caption*{Original text image.}
-    \end{subfigure}
-    \begin{subfigure}{\linewidth}
-    \centering
-    \includegraphics[width=.7\linewidth]{images/cont}
-    \vspace{-5px}
-    \caption*{Contouring algorithm.}
-    \end{subfigure}
-    \begin{subfigure}{\linewidth}
-    \centering
-    \includegraphics[width=.7\linewidth]{images/lines}
-    \vspace{-5px}
-    \caption*{Postprocessing and line selection.}
-    \end{subfigure}
-    \vspace{-10px}
-    \caption{Text segmentation steps}
-    \label{fig:text_segm}
-\end{figure}
-
-\vspace{-20px}
 \subsection{line selection}
@@
 \subsection{Oversegmentation correction}
-In this section we provide techniques to verify the correctness of these segmentation points. 
-\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
-[6:39]
-In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
+In this section we provide a method to fix the oversegmentation of the word image caused by the previous steps.
+To achieve this we process the potential segmentation points with a few rule based filters, to remove redundant and incorrect segmentation points.
+The final decision to which segmentation points we retain is made by a neural network.
 
@@
 More rule based filter could be added for more optimal processing of the potential segmentation points.
-However the focus the segmentation correction is the neural network used to remove incorrect segmenatation points.
 
@@
 Since it is impossible to validate these responses from the neural network, this technique is not used in the final version of our oversegmentation correction.
+
+%As our algorithm for segmentation of handwritten words splits the word into to many characters, we can use techniques to verify the correctness of these segmentation points points. 
+%\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
+%In this paper we will only discuss the latter, machine learning based procedures, as we used this approach.
 
@@
 The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect segmentation points the neural network has to process.
-Combinations with other approaches are also more likely provide better results.
-
-The neural network ..
+Combinations with other approaches could also yield better results
 

### 2017-12-17 Report/paper.tex
small changes

--- 2017-12-17 Report/paper.tex
@@
 ]
-% Notes:
-%  redo the preprocessing experiment and explain why some techniques increase the accuracy more than others.
 \begin{abstract}
@@
 the exact movements of the pen.
-\citep{olcr} present a technology for online recognition of Japanese handwritten text with an accuracy of 94.6\%.
+\cite{olcr} present a technology for online recognition of Japanese handwritten text with an accuracy of 94.6\%.
 
@@
 Theoretically feeding normalized data to the neural network returns the same output as before.
-In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \citep{NormGoal}.
+In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}.
 We normalize the data by reducing the input dimensions from (1200,900,3) to (64,64,1). A simple threshold is added to improve the contrast in our image.
@@
 As our dataset only contains 3410 images with handwritten characters, we need more data for the neural network to be able to function properly.
-Next we will describe the data augmentation techniques we used. For these techniques we based us on \citep{DataAug}. A more thorough explanation can be found there.
+Next we will describe the data augmentation techniques we used. For these techniques we based us on \cite{DataAug}. A more thorough explanation can be found there.
 
@@
 In order to deal with different kinds of slants we add sheared versions of the image to the dataset.
-Shearing displaces each point in a fixed direction, by an amount proportional to its signed distance from a line that is parallel to that direction \citep{Shear}.
+Shearing displaces each point in a fixed direction, by an amount proportional to its signed distance from a line that is parallel to that direction \cite{Shear}.
 We used the following shear mapping to achieve this:
@@
 This overview is visualized in Figure \ref{fig:dnn}.
-A detailed explanation on how deep neural networks function is explained by \citep{nnbook}.
+A detailed explanation on how deep neural networks function is explained by \cite{nnbook}.
 
@@
 Dropout greatly reduces overfitting in the fully connected layers.
-\citep{dropout} We do not use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
+\cite{dropout} We do not use this after the convolutional layers as they are very resistant to overfitting \cite{nnbook}.
 The last fully connected layer has 62 output neurons, which correspond to the 62 possible output classes.
@@
 Each of these methods effectively improve the previously mentionned method.
-The Adam optimizer, introduced by \citep{adam}, is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
+The Adam optimizer, introduced by \cite{adam}, is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
 This optimizer achieves the highest accuracy in our case and will be used.
@@
 Neural networks with this constraint tend to overfit less because small weights are less sensible to small changes of the input.
-\citep{presham}
+\cite{presham}
 
@@
 This is an adaptation of the Zhang-Suen thinning algorithm, one of the most commonly used and adapted thinning algorithms.
-\citep{zsthinning}
+\cite{zsthinning}
 
@@
 In this section we provide techniques to verify the correctness of these segmentation points. 
-\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
+\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
 [6:39]
@@
 \subsubsection{Neural network}
-The neural network we use to judge the potential segmentation points, is based on \citep{evalsplitpointsnn}.
+The neural network we use to judge the potential segmentation points, is based on \cite{evalsplitpointsnn}.
 The paper explains how to use neural networks to classify the correctness of segmentation points.
@@
 In order to create training data we created a small program to iterate over splitting points of handwritten words from the IAM dataset.
-\citep{iam}
+\cite{iam}
 
 This method allowed us to achieve an accuracy of 72 \%, which is still fairly low because of the difficulty of the problem and the small amount of traing data.
-However this is better than our accuracy with an implementation of the method described by \citep{evalsplitpointsnn}.
+However this is better than our accuracy with an implementation of the method described by \cite{evalsplitpointsnn}.
 Segmentation points can now be filtered if the neural network decides a segmentation point is incorrect.
@@
 The probability is computed as $P(w_i | w_{i-n},...,w_{i-1})$.
-We use the Markov assumption here: the probability of a word in the context can be approximated by the previous $n-1$ words. \citep{markov} This probability is calculated as follows:
+We use the Markov assumption here: the probability of a word in the context can be approximated by the previous $n-1$ words. \cite{markov} This probability is calculated as follows:
 \begin{equation}
@@
 An n-gram is a sequence of n words from a given sequence of text, denoted as $w_i,...,w_{i+n}$.
-\citep{ngrams}
+\cite{ngrams}
 We use the n-grams from the Google dataset for this purpose.
-\citep{google} Querying these datasets locally takes several minutes and this happens for every word.
+\cite{google} Querying these datasets locally takes several minutes and this happens for every word.
 Instead we use a Python library phrasefinder to make online queries to these corpora.
@@
 Comparisons are hard to find as few papers discuss off-line handwritten character recognition.
-One paper we found, \citep{charrec}, achieves an accuracy of 98.5 \%, they classify 38 classes (10 digits, 26 letters, 2 special symbols) and manually extract features for the neural network instead of using convolutional layers.
+One paper we found, \cite{charrec}, achieves an accuracy of 98.5 \%, they classify 38 classes (10 digits, 26 letters, 2 special symbols) and manually extract features for the neural network instead of using convolutional layers.
 
@@
 These methods are mostly based on hidden Markov models and neural networks.
-This concept is explained and applied by \citep{presham}.
-%%% Dit is jouw oude tekst %%%
-There are a lot of steps in this paper from which we can draw conclusions.
-The segregation of a text image in word images is
-Word extraction is a problem
-
-Character segmentation is another ..
-The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect segmentation points the neural network has to process.
-Combinations with other approaches are also more likely provide better results.
-
-The neural network ..
-
-The neural network
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This concept is explained and applied by \cite{presham}.
 \section*{Acknowledgements}

### 2017-12-17 Report/paper.tex
change sizes of images

@@
   \centering
-  \includegraphics[width=\linewidth]{images/bad_char1}
+  \includegraphics[height=2cm]{images/bad_char1}
 \end{subfigure}
@@
   \centering
-  \includegraphics[width=\linewidth]{images/bad_char2}
+  \includegraphics[height=2cm]{images/bad_char2}
 \end{subfigure}
-\caption{Examples of written 'z' characters, which are hard to recognise or ambiguous.}
+\caption{Examples of written 'z' characters, which are hard to recognise or are ambiguous.}
 \label{fig:char}
@@
 
-\begin{figure}
-    \begin{subfigure}{\linewidth}
+\begin{figure*}
+    \begin{subfigure}{0.33\textwidth}
     \centering
-    \includegraphics[width=.7\linewidth]{images/vpp}
-    \vspace{-5px}
+    \includegraphics[width=\textwidth]{images/vpp}
     \caption*{Vertical projection profile.}
     \end{subfigure}
-    \begin{subfigure}{\linewidth}
+    \begin{subfigure}{0.33\textwidth}
     \centering
-    \includegraphics[width=.7\linewidth]{images/rules}
-    \vspace{-5px}
+    \includegraphics[width=\textwidth]{images/rules}
     \caption*{Rule based correction.}
     \end{subfigure}
-    \begin{subfigure}{\linewidth}
+    \begin{subfigure}{0.33\textwidth}
     \centering
-    \includegraphics[width=.7\linewidth]{images/nn}
-    \vspace{-5px}
+    \includegraphics[width=\textwidth]{images/nn}
     \caption*{Correction through neural network.}
@@
     \label{fig:char_segm}
-\end{figure}
+\end{figure*}
 
@@
             ]
-        \Tree [.{(T, 0.1)}
-        [.{(Te , 0.08)} {(Tes, 0.048)} {(Te5 , 0.032)} ]
-        [.{(Tl , 0.02)} {(Tls, 0.012)} {(Tl5 , 0.008)} ]
-            ]
-\caption{A list of the trees for the word 'Yes' with branching factor 2.
+\caption{The first element of a possible list of trees for the word 'Yes' with branching factor 2.
 The characters have the following probabilities: $P(Y)=0.9, P(T)=0.1, P(e)=0.8, P(l)=0.2, P(s)=0.6, P(5)=0.4$.

### 2017-12-17 Report/paper.tex
small fixes

@@
 In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}.
-We normalize the data by reducing the input dimensions from (1200,900,3) to (64,64,1). A simple threshold is added to improve the contrast in our image.
+We normalize the data by reducing the input dimensions to (64,64,1). A simple threshold is added to improve the contrast in our image.
 This allows the foreground to be more distinguishable from the background and allows the neural network to train and recognise more easily.
@@
 
-\subsection{binarization}
+\subsection{Binarization}
 The first preprocessing step is binarizing the text image.
@@
 
-\subsection{contouring}
+\subsection{Contouring}
 Our heuristic text segmentation algorithm is based on finding the contours in the text image, so we can group these contours in postprocessing stages to recover the original words from the text image.
@@
 
-\subsection{postprocessing}
+\subsection{Postprocessing}
 In the postprocessing step, we eliminate the redundant extracted contours, and reconnect the contours of words that consist of multiple parts.
 
-\subsubsection{contour removal}
+\subsubsection{Contour removal}
 Since the contouring algorithm produces contours that are not contours of the words in the text, these contours need to be filtered out and removed.
@@
 
-\subsubsection{contour combination}
+\subsubsection{Contour combination}
 With the excess contours removed, we can begin to reconstruct the original words from the text by combining multiple contours into one where needed.
@@
 \vspace{-20px}
-\subsection{line selection}
+\subsection{Line selection}
 Now that the words are extracted from the text image, the next goal is to try and reconstruct the original text by reordering these words in their original lines.
@@
 \begin{algorithm}[tb]
-   \caption{Line reconstruction algorithm}
+   \caption{Algorithm for line reconstruction}
    \label{alg:linereconstr}
@@
 \cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures. 
-[6:39]
 In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
 
-\subsubsection{rule based}
+\subsubsection{Rule based techniques}
 The main oversegmentation correction is handled by our neural network.
@@
 
-\subsubsection{Neural network}
-The neural network we use to judge the potential segmentation points, is based on \cite{evalsplitpointsnn}.
+\subsubsection{Machine learning approach}
+The neural network we use to judge the correctness of segmentation points, is based on \cite{evalsplitpointsnn}.
 The paper explains how to use neural networks to classify the correctness of segmentation points.
@@
 \section{Future work}
-Our recognition system cannot be used for real-world applications.Every individual technique has a certain error, which increases the difficulty of the next step, e.g.
+Our recognition system cannot be used for real-world applications.
+Every individual technique has a certain error, which increases the difficulty of the next step, e.g.
 if word segmentation contains mistakes, character recognition becomes much harder.
 This results in a system which cannot correctly recognize handwritten text.
-However many improvements are still possible.
+Many improvements are still possible.
 
@@
 The character segmentation process is inadequate as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
-More postprocessing can be used through rule based filters and a more robust neural network to correct the found segmentation points, but this will never reach an optimal solution through rule based filters alone.
+More postprocessing can be used through rule based filters and a more robust neural network to correct the found segmentation points, but this will never reach an optimal solution through rule based filters alone because of the ambiguity of this problem.
 

### 2017-12-17 Report/paper.tex
last reference

--- 2017-12-17 Report/paper.tex
@@
 ]
+% Notes:
+%  redo the preprocessing experiment and explain why some techniques increase the accuracy more than others.
 \begin{abstract}
@@
 Then we describe how to recognise characters with a deep neural network approach.
-Finally we present postprocessing techniques to correct mispredicted text. 
-Our approach does not outperform state-of-the-art systems but hopes to provide a thorough elaboration for every step in handwriting recognition. 
+Finally we presentf postprocessing techniques to correct mispredicted text.
 \end{abstract}
@@
 Theoretically feeding normalized data to the neural network returns the same output as before.
-In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \cite{NormGoal}.
+In reality however, it is better to normalize data to avoid getting stuck in local optima and to increase training speed of the model \citep{NormGoal}.
 We normalize the data by reducing the input dimensions from (1200,900,3) to (64,64,1). A simple threshold is added to improve the contrast in our image.
@@
 In order to deal with different kinds of slants we add sheared versions of the image to the dataset.
-Shearing displaces each point in a fixed direction, by an amount proportional to its signed distance from a line that is parallel to that direction \cite{Shear}.
+Shearing displaces each point in a fixed direction, by an amount proportional to its signed distance from a line that is parallel to that direction \citep{Shear}.
 We used the following shear mapping to achieve this:
@@
 Each of these methods effectively improve the previously mentionned method.
-The Adam optimizer, introduced by \cite{adam}, is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
+The Adam optimizer, introduced by \citep{adam}, is an optimization of AdaDelta, which uses a distinct learning rate and momentum for every parameter.
 This optimizer achieves the highest accuracy in our case and will be used.
@@
 To be able to recognise the characters in a given text, we will first need to extract the words from the text before we can start evaluating the characters.
-This word extraction has many possible approaches, but in this paper we will focus solely on the word contouring approach.
+This word extraction has many possible approaches, but in this paper we will focus solely on the word contouring approach, as pointed out by following paper \cite{WordSegm} as being a good word extraction approach.
 This method uses a heuristic algorithm to extract the contours of the words from a given text, and through postprocessing we can rearrange these extracted words into lines.
@@
 Again, there is the choise between several different approaches.
-In this paper we focus solely on a word oversegmentation approach to character extraction.
+In this paper we focus solely on a word oversegmentation approach to character extraction, as can be found described here \cite{CharSegm} and \cite{CharSegmOld}.
 The purpose is to segment the words into the individual characters.
@@
 In this section we provide techniques to verify the correctness of these segmentation points. 
-\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
-[6:39]
+\citep{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
 In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
@@
 In order to create training data we created a small program to iterate over splitting points of handwritten words from the IAM dataset.
-\cite{iam}
+\citep{iam}
 
 This method allowed us to achieve an accuracy of 72 \%, which is still fairly low because of the difficulty of the problem and the small amount of traing data.
-However this is better than our accuracy with an implementation of the method described by \cite{evalsplitpointsnn}.
+However this is better than our accuracy with an implementation of the method described by \citep{evalsplitpointsnn}.
 Segmentation points can now be filtered if the neural network decides a segmentation point is incorrect.
@@
 The probability is computed as $P(w_i | w_{i-n},...,w_{i-1})$.
-We use the Markov assumption here: the probability of a word in the context can be approximated by the previous $n-1$ words. \cite{markov} This probability is calculated as follows:
+We use the Markov assumption here: the probability of a word in the context can be approximated by the previous $n-1$ words. \citep{markov} This probability is calculated as follows:
 \begin{equation}
@@
 These methods are mostly based on hidden Markov models and neural networks.
-This concept is explained and applied by \cite{presham}.
+This concept is explained and applied by \citep{presham}.
+%%% Dit is jouw oude tekst %%%
+There are a lot of steps in this paper from which we can draw conclusions.
+The segregation of a text image in word images is
+Word extraction is a problem
+
+Character segmentation is another ..
+The proposed oversegmentation serves the intended purpose, however more rule-based postprocessing steps are needed to reduce the amount of incorrect segmentation points the neural network has to process.
+Combinations with other approaches are also more likely provide better results.
+
+The neural network ..
+
+The neural network
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Acknowledgements}

### 2017-12-17 Report/paper.tex
replaced vspace, my bad
reference fixing, paper small change

@@
 
-To be able to recognise the characters in a given text, we will first need to extract the words from the text before we can start evaluating the characters.
+To be able to recognise the characters in a given text, we first need to extract the words from the text before we can start evaluating the characters.
 This word extraction has many possible approaches, but in this paper we will focus solely on the word contouring approach, as pointed out by following paper \cite{WordSegm} as being a good word extraction approach.
@@
     \label{fig:text_segm}
+    \vspace{-20px}
 \end{figure}
 
-\vspace{-20px}
 \subsection{Line selection}

### 2017-12-17 Report/paper.tex
start rereading

@@
 To be able to recognise the characters in a given text, we first need to extract the words from the text before we can start evaluating the characters.
-This word extraction has many possible approaches, but in this paper we will focus solely on the word contouring approach, as pointed out by following paper \cite{WordSegm} as being a good word extraction approach.
-This method uses a heuristic algorithm to extract the contours of the words from a given text, and through postprocessing we can rearrange these extracted words into lines.
-The algorithm proposed is only to processes text images, and does not take into account images and other non-textual areas found in the text.
-Dependant on the situation, additional techniques for page segmentation can be utilized. % , such as the Run-Length Smearing Algorithm, can be utilized.
+Word extraction from text images has many possible approaches, but in this paper we will focus solely on the word contouring approach, pointed out by following paper \cite{WordSegm} as being a good word extraction approach.
+
+This heuristic approach will extract the contours of the words from a given text image, and through postprocessing we can rearrange these extracted words into lines.
+The proposed approach focusses solely text images, and does not take into account images and other non-textual areas found in the text.
+Dependent on the situation, additional techniques for page segmentation can be utilized. % , such as the Run-Length Smearing Algorithm, can be utilized.
 % TODO:: Run length Smearing Algorithm reference --> ....

### 2017-12-17 Report/paper.tex
update abstract

@@
 The aim of this paper is to present a complete system for off-line handwriting recognition.
-We discuss how to segment texts with contouring. Next we present our method of segmentating words into characters.
-Then we describe how to recognise characters with a deep neural network approach.
+We discuss how to segment handwritten text through contouring. Next we present our method of segmenting words into characters.
+Then we propose a method to recognize characters with a deep neural network.
 Finally we present postprocessing techniques to correct mispredicted text.
-Our approach does not outperform state-of-the-art systems but hopes to provide a thorough elaboration for every step in handwriting recognition.
+Our approach does not outperform state-of-the-art systems but hopes to provide a thorough elaboration for every step in the process.
 \end{abstract}

### 2017-12-17 Report/paper.tex
improve introduction

@@
 This paper presents an off-line handwriting recognition system.
-This system digitizes handwritten texts, which allows modern data processing, storage a transmission techniques to be applied to the information stored in the handwritten documents.
+This system digitizes handwritten texts, which allows modern data processing, storage and transmission techniques to be applied to the information stored in the handwritten documents.
 We only cover off-line handwriting recognition in this paper, where we have the full handwritten text in advance.

### 2017-12-17 Report/paper.tex
correction of errors in section 2

@@
 The smallest distinguishable token in a text is a character. These handwritten characters are uniquely
-written by each individual and variations even occur when written by the same individual. These variations include the inclination of characters, i.e., the slant, orientation and size.
+written by each individual and variations even occur when written by the same individual. These variations include the inclination of characters (i.e. the slant) orientation and size.
 A way of dealing with these includes removing variations from the input image using preprocessing techniques.
@@
 The architecture of this deep neural network is described in Section \ref{sec:dnn}.
-In order to make the deep neural network recognise characters in images with these variations, we augmented the dataset to include a bigger variety of images.
+In order to make the deep neural network recognize characters in images with these variations, we augmented the dataset to include a bigger variety of images.
 We present these augmentation techniques in Section \ref{sec:preproc}.
@@
 Before exposing the neural network to new characters, it needs to be familiar with many training samples to be able to classify any new character correctly.
-This is a why a good dataset is very important. A dataset should have many examples, much variation, and an even balance between classes. A dataset with these properties
-allow the neural network to train effectively and recognise any new character given to the network.
+This is why a good dataset is very important. A dataset should have many examples, much variation, and an even balance between classes. A dataset with these properties
+allows the neural network to train effectively and recognize any new character given to the network.
 
@@
 We normalize the data by reducing the input dimensions to (64,64,1). A simple threshold is added to improve the contrast in our image.
-This allows the foreground to be more distinguishable from the background and allows the neural network to train and recognise more easily.
+This allows the foreground to be more distinguishable from the background and allows the neural network to train and recognize characters more easily.
 
@@
 \label{par:aug}
-Data augmentation serves two purposes. It makes the dataset more robust to different kinds of handwriting as we add variations of the original data to the input data.
-The second purpose is to increase the size of our dataset.
-As our dataset only contains 3410 images with handwritten characters, we need more data for the neural network to be able to function properly.
+Data augmentation serves two purposes. It makes the dataset more robust to different kinds of handwriting as we add variations of the original data to the training samples.
+The second purpose is to increase the size of our dataset. As our dataset only contains 3410 images of handwritten characters, we need more data for the neural network to be able to function properly.
 Next we will describe the data augmentation techniques we used. For these techniques we based us on \cite{DataAug}. A more thorough explanation can be found there.
@@
 \paragraph{Translations of data}
-The dataset is also extended with variations in padding of characters in images. For this purpose we multiply every coordinate of the original image with the following transformation matrix:
+The dataset is extended with variations in padding of characters in images. For this purpose we multiply every coordinate of the original image with the following transformation matrix:
 \begin{equation}
@@
 where $t_x$ is the horizontal shift and $t_y$ is the vertical shift.
-The values for chosen for $t_x$ are -8, 0 and 8, the values chosen for $t_y$ are -4, 0 and 4.
+The values for chosen for $t_x$ are -16, 0 and 16, the values chosen for $t_y$ are -8, 0 and 8.
 All combinations are used except $t_x=0$ and $t_y=0$.
-The translation only occurs if it does not displace the character outside of the image.
+The translation only occurs if it does not move the character outside of the image.
 \paragraph{Rotations of data}
-Rotations of the image are added to the dataset to make the network more robust to handwritings with different orientations. This matrix is equal to:
+Rotations of the image are added to the dataset to make the network more robust to handwritings with different orientations. The transformation matrix is equal to:
 \begin{equation}
@@
 This causes bright regions to get thinner and dark regions to get bigger in the image.
+
 All of the previously discussed preprocessing techniques are also applied to the eroded image.
 The final dataset contains translated, orientated, sheared and scaled versions of both the original image and the eroded image.
-A visualitation of the augmentation of an augmented picture can be seen in Figure \ref{fig:augmented}.
+A visualitation of the augmentation of a picture can be seen in Figure \ref{fig:augmented}.
 
@@
 \label{sec:dnn}
-% TODO: make a drawing, see http://www.ais.uni-bonn.de/deep_learning/images/Convolutional_NN.jpg
 Humans have millions of connected neurons in the visual cortices which are capable of image processing.
-We can recognise most characters without any thought because of this complex neural network inside our brain.
+We can recognize most characters without any thought because of this complex neural network inside our brain.
 For computers, character recognition is a more difficult task.
@@
 We use a neural network, inspired by how the brain processes images.
-These deep neural networks contain several layers of connected perceptrons, where each connection is defined by weight and each perceptron is defined by a bias.
+These deep neural networks contain several layers of connected perceptrons, where each connection is defined by a weight and each perceptron is defined by a bias.
 The perceptrons take an input and calculate an output using an activation function. This allows perceptrons to make decisions.
@@
   \caption{Visualization of the deep neural network for character recognition.
-The first row shows the shape a layer.
+The first row shows the shape of a layer.
 The second row shows the visualization of that layer, the third row shows information about the layer.
@@
  The input is a normalized image with one color channel.
-This is fed to a convolutational layer.
+This is fed to a convolutional layer.
 Convolutional layers allow us to extract features by scanning the image with filters.
@@
  We can apply several of those filters using one convolutional layer to extract several types of features.
- This works because images have a spatial structure: close pixels have shared information.
+ This works because images have a spatial structure; i.e.
+close pixels have shared information.
 By applying a filter, pixels in the same region will use shared biases and weights, which allows the filter to extract particular features.
@@
 This one-dimensional array can now be fed to a fully-connected layer, which is a layer of connected perceptrons as previously discussed.
-There are three fully-connected layers each connected with a dropout layer.
+There are three fully-connected layers of which the first two are connected with a dropout layer and the last one is the output layer.
 Dropout randomly drops units and their connections from the network during training.
@@
  The last layer in the neural network is a softmax layer with a cross-entropy cost function.
-The softmax layer compute a probability distribution (a set of 62 numbers which sum up to 1) using the 62 output neurons from the previous layer.
+The softmax layer computes a probability distribution (i.e.
+a set of 62 numbers which sum up to 1) using the 62 output neurons from the previous layer.
 An element in the probability distribution estimates the probability of the input image to classify as a certain character.
  The cross-entropy cost function uses the actual labels to calculate the error between the actual labels and the predicted labels (the previously discussed probabilities).
-The closer the cross-entropy function is to zero, the better it is at classifying handwritten characters.
+The closer the cross-entropy value is to zero, the better the neural network is at classifying handwritten characters.
  \subsubsection{Weight adjustment}
@@
 The Adam optimizer effectively fulfills this goal. In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow.
-First of all, stochastic gradient descent (SGD) is an iterative method which tries to find minima.
-This function takes a step size, which is known as the learning rate.
+First of all, stochastic gradient descent (SGD) is an iterative method that tries to find minima.
+This function takes a step size which is known as the learning rate.
 SGD can be used to minimize the cross-entropy function iteratively by adjusting the weights.

### 2017-12-17 Report/paper.tex
progress

@@
 \label{sec:segtext}
+In this section we explain the first step in the process of extracting individual characters from a text image: i.e. word extraction.
+Word extraction from text images has many possible approaches, but in this paper we focus solely on the word contouring approach, pointed out by following paper \cite{WordSegm} to be a successfull word extraction approach.
 
-To be able to recognise the characters in a given text, we first need to extract the words from the text before we can start evaluating the characters.
-Word extraction from text images has many possible approaches, but in this paper we will focus solely on the word contouring approach, pointed out by following paper \cite{WordSegm} as being a good word extraction approach.
-
-This heuristic approach will extract the contours of the words from a given text image, and through postprocessing we can rearrange these extracted words into lines.
-The proposed approach focusses solely text images, and does not take into account images and other non-textual areas found in the text.
-Dependent on the situation, additional techniques for page segmentation can be utilized. % , such as the Run-Length Smearing Algorithm, can be utilized.
-% TODO:: Run length Smearing Algorithm reference --> ....
+With this approach we extract the contours of the words in a given text image. Through postprocessing techniques we can rearrange these extracted words into lines.
+The proposed approach focusses solely on text images, and does not take into account non-textual areas found in the input image.
+Additional techniques for page segmentation can be utilized depending on the situation.
 
 \subsection{Binarization}
-The first preprocessing step is binarizing the text image.
-If the image is not yet greyscaled, this is done at the start of the preprocessing phase.
-The greyscaled image is converted from greyscaled into a binary image.
-This process will allows us to easily separate the text from the background.
+We start with binarizing the text image.
 
-To achieve this there are a lot of thresholding algorithms to choose from, but the one we will use throughout the paper is a combination of executing Gaussian blur in combination with OTSU's thresholding algorithm on the binary text image.
-The algorithms used are as implemented in the python opencv library. %TODO:: where do i put this????
+To achieve this there are a lot of thresholding algorithms to choose from, but the one we use is a combination of Gaussian blur in combination with Otsu's thresholding algorithm.
 
-The Gaussian blur will reduce the noise in the image, and smoothen the edges of the text.
-The smoothening of the background noise will result in less noise picked up by the chosen thresholding algorithm, so it is less likely to confuse background noise with the words from the text image.
-
-The tresholding algorithm used in this paper is OTSU's thresholding algorithm.
-This algorithm will deliver adequate result for noisy text documents.
-The choise of tresholding algorithm depends largely on intended purposes of the algorithm.
-If very noisy documents will need to be processed, additional noise removal techniques and other preprocessing techniques may be required to reduce the excess noise picked up by the tresholding algorithm.
+The Gaussian blur will smoothen the noise in the image and smoothen the edges of the text.
+%TODO:: Otsu reference
+Otsu's tresholding algorithm will deliver adequate result for noisy text documents.
+The choise of tresholding algorithm depends mostly on the intended purposes of the algorithm.
+Additional noise removal techniques and other preprocessing techniques may be required for a better result when processing very noise-heavy images.
 
 \subsection{Contouring}
-Our heuristic text segmentation algorithm is based on finding the contours in the text image, so we can group these contours in postprocessing stages to recover the original words from the text image.
+Our text segmentation approach is based on finding contours in the text image.
+We can find these contours using %TODO:: reference paper contouring suzuki85
+These contours in postprocessing stages to recover the original words from the text image.
 The contouring algorithm used is the one implemented in the opencv library for python.
@@
 Before thresholding the image, Gaussian blur is executed to reduce the noise of the word image, and to smoothen the edges of the characters.
-Afterwards we use OTSU's thresholding algorithm again to binarize the blurred word image.
+Afterwards we use Otsu's thresholding algorithm again to binarize the blurred word image.
 This thresholding step is essential for out skeletonization algorithm in the next step, as it requires a binarized image as input.

### 2017-12-17 Report/paper.tex
go

@@
 We can find these contours using %TODO:: reference paper contouring suzuki85
-These contours in postprocessing stages to recover the original words from the text image.
-The contouring algorithm used is the one implemented in the opencv library for python.
-This algorithm generates a lot of unneeded contours, which we will filter through postprocessing.
-Also since words in handwritten text are not always connected into one piece, we will need to connect these contours in the postprocessing stages as well.
+Through postprocessing we can recover the original words from these extracted contours.
 
 \subsection{Postprocessing}
-In the postprocessing step, we eliminate the redundant extracted contours, and reconnect the contours of words that consist of multiple parts.
+We can remove the incorrect contours generated by the contouring algorithm through postprocessing.
+Secondly we will reconnect words that were split because they are not written in a single stroke.
 
 \subsubsection{Contour removal}
-Since the contouring algorithm produces contours that are not contours of the words in the text, these contours need to be filtered out and removed.
-To accomplish this we make use of the bounding box we can draw around a contour.
+The contouring algorthm can create incorrect contours because of left-over noise in the image.
+The tool we will use to remove these incorrect contours is the bounding box we can draw around a contour.
 This bounding box is the smallest rectangle containing said contour, and it will allow us to easily compare contour positions in the text image.
 
-The first filter is removing the contours of which the bounding box is contained by the bounding box of another contour.
-This is allowed because the contents of this inner bounding box will be regarded when processing the contents of the outer bounding box, and so we can remove the contours belonging to the inner bounding boxes.
+Firstly we remove the contours of which the bounding box is contained by the bounding box of a second contour.
+We can do this because the text segmentation process will return the word in the image contained by its bounding box, so we will not change  % TODO:: NOTATION
 
-The second filter is removing the contours that are too small to contain parts of a word.
-Since the contouring algorithm extracts the contours for any size, and because our character recognition network does not process punctuation,
-we need to eliminate these from our results.
+Secondly we remove contours that are too small to contain parts of a word.
+The contouring algorithm will generate contours of noise and punctuation.
+These contours we do not want to process, so we will filter them out.
 
-We achieve this by taking the average height of all the found contours in the text.
-From these contours we calculate the average height, after which we remove the contours with a height lesser than the average height multiplied by a chosen parameter $minimalHeightMultiplier$.
-The optimal results were achieved when we initialize with $minimalHeightMultiplier = 2/3$, so that we retain only contours higher than the average height $ * 2/3$.
+This is done by taking the average height of all the found contours in the text.
+Iteratively we remove remove the contours with a height lesser than the a predetermined threshold,
+\vspace{-5px}
+\begin{equation}
+    threshold = averageHeight * removalMultiplier
+\end{equation}
+with the $removalMultiplier$ initialized on $2 / 3$ as chosen value.
 
 \subsubsection{Contour combination}
-With the excess contours removed, we can begin to reconstruct the original words from the text by combining multiple contours into one where needed.
-The algorithm used for this is a straightforward overlapping algorithm, that combines word segments based on the overlap between the bounding box of their contours.
-The parameters for this algorithm are the $alowedWordGap$ and the $allowedVerticalOverlap$ paramters.
-These parameters are initialized as $interWordSegmentDistance = $ the average width of the found words $ / 16 $ and $allowedVerticalOverlap = 1/2$ to give the best results in the our testset, however further experimentation with these parameters and the algorithm is required to find a more generalized optimal solution.
+With the excess contours removed, we begin to reconstruct the original words from the text.
+We do this by combining multiple contours into one where a word was separated into multiple parts.
+We have created a straightforward overlapping algorithm, that combines word segments based on the overlap between the bounding box of their contours.
+
+We have a parameter for the allowed gap between parts of the same word, and one for the minimal y-axis overlap of both their boundingboxes.
+
+\begin{equation} %TODO:: FIX INDENTATION
+    if \\ 
+        wordXaxisGap < \\
+            averageWordWidth * (1 / (averageCharacterCount * 2) )\\
+        and \\
+        wordYaxisOverlap > \\
+        min(word1.height, word2.height) * requiredYaxisOverlap\\
+    then \\
+        combineWords
+\end{equation}
+The optimal parameters we found are:
+
+\noindent 
+$averageCharacterCount = 8$ and
+
+\noindent
+$requiredYaxisOverlap = 1/2$
+
+With these conditions fulfilled, we combine both contours into one single contour, containing the found word.
+
+\begin{algorithm}[tb]
+   \caption{Algorithm for line reconstruction}
+   \label{alg:linereconstr}
+\begin{algorithmic}
+   \STATE {\bfseries Input:} $extractedWords$, $textlines$
+   \REPEAT
+        \STATE $highestWord = findHighestWord $
+        \STATE $newline = [highestWord]$
+        \STATE Initialize $noChange = true$.
+        \STATE $avgY= newline$.findAverageYpos
+        \REPEAT
+            \FOR{$word$ {\bfseries in} $extractedWords$}
+            \STATE $boundRect = word.contour.boundingRect$
+            \IF{ $avgY > word.y$ {\bfseries and} $avgY < word.y + word.h$ }
+            \STATE $line$.add $word$
+            \STATE $noChange = false$
+            \ENDIF
+        \ENDFOR
+        \UNTIL{$noChange = true$}
+        \STATE $textlines$.add $newline$
+   \UNTIL{$words$.empty $= true$}
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Line selection}
+With the words extracted from the text, the next goal is to reconstruct the original text.
+We will need to order these words into lines 
+For the line selection process, an iterative algorithm, Algorithm \ref{alg:linereconstr}, is used.
+
+This algorithm reorders the found words in lines, and return these lines as ordered in the original text image.
 
@@
 \end{figure}
-
-\subsection{Line selection}
-Now that the words are extracted from the text image, the next goal is to try and reconstruct the original text by reordering these words in their original lines.
-For the line selection process, an iterative algorithm, Algorithm \ref{alg:linereconstr}, is used.
-\begin{algorithm}[tb]
-   \caption{Algorithm for line reconstruction}
-   \label{alg:linereconstr}
-\begin{algorithmic}
-   \STATE {\bfseries Input:} $extractedWords$, $textlines$
-   \REPEAT
-        \STATE $highestWord = findHighestWord $
-        \STATE $newline = [highestWord]$
-        \STATE Initialize $noChange = true$.
-        \STATE $avgY= newline$.findAverageYpos
-        \REPEAT
-            \FOR{$word$ {\bfseries in} $extractedWords$}
-            \STATE $boundRect = word.contour.boundingRect$
-            \IF{ $avgY > word.y$ {\bfseries and} $avgY < word.y + word.h$ }
-            \STATE $line$.add $word$
-            \STATE $noChange = false$
-            \ENDIF
-        \ENDFOR
-        \UNTIL{$noChange = true$}
-        \STATE $textlines$.add $newline$
-   \UNTIL{$words$.empty $= true$}
-\end{algorithmic}
-\end{algorithm}
-
-This algorithm reorders the found words in lines, and return these lines as ordered in the original text image.
-
-This algorithm fares well processing text images, where the text consists of parrallel text lines.
-Angeled parallel text lines can be compensated for by rotating and padding the thresholded text image with different angles, and comparing the results.
-
-Combinations of parrallel and slanted text lines however are not handled well by the algorithm, and will require further adaptations or more advanced algorithms to map the words to text lines correctly.
 

### 2017-12-17 Report/paper.tex
until 4.5

@@
 
-\subsection{Line selection}
+\subsection{Line reconstruction}
 With the words extracted from the text, the next goal is to reconstruct the original text.
@@
 
-The next step in the process is to extract the character images from the given word images.
-Again, there is the choise between several different approaches.
-In this paper we focus solely on a word oversegmentation approach to character extraction, as can be found described here \cite{CharSegm} and \cite{CharSegmOld}.
-The purpose is to segment the words into the individual characters.
-This will be done through initial oversegmentation of the word image, followed by postprocessing steps to remove the excess separation points to recover the original character images.
+The next step in the process is to extract the character images from the found word images.
+Again, there is a choice between several different approaches.
+In this paper we focus solely on a word oversegmentation approach to character extraction, through skeletonization and using a vertical projection profile.
+The reasoning begind this can be found in the following papers \cite{CharSegm} and \cite{CharSegmOld}.
+
+After first oversegmenting the word, we will recover the original character through postprocessing steps combining different word segments.
 
 \subsection{Normalization} % TODO:: necessary? never taken advantage of? maybe for split points count etc???
-The first action is normalizing the word image.
-This normalization step enables us to reason about words written in different pixel scalings the same way.
-The process of normalizing the word image consists of resizing the image to a fixed height of $100px$.
+The first action is again normalizing the word image.
+We will resuze the image to a fixed height of $100px$.
+This will result in more consistency in rule-based post processing steps.
 
 \subsection{Binarization}
-Next we will again need to execute a binarization algorithm on the word image.
-Before thresholding the image, Gaussian blur is executed to reduce the noise of the word image, and to smoothen the edges of the characters.
-Afterwards we use Otsu's thresholding algorithm again to binarize the blurred word image.
-This thresholding step is essential for out skeletonization algorithm in the next step, as it requires a binarized image as input.
+The word image is binarized.
+This is again done with a combination of Gaussian blur and Otsu's thresholding algorithm.
+The binarization is needed for the next step, because we can only apply a skeletonization algorithm to a binary image.
 
 \subsection{Skeletonization}
-The skeletonization of the thresholded word image will allow us to ignore the thickness of the lines in further processing.
-This normalization is needed to reason about potential segmentation points between characters in the word image.
-After comparing different skeletonization algorithms, the algorithm of choise is the skeletonize function found in the scipy python library.
-This is an adaptation of the Zhang-Suen thinning algorithm, one of the most commonly used and adapted thinning algorithms.
-\cite{zsthinning}
+Now we will skeletonize our word image.
+This will normalize the thickness of the strokes to a single pixel.
+This allows us to execute our Verical projection profile and interpret the results easily.
+The skeletonization algorithm used is the scipy adaptation of the Zhang Suen thinning algorithm \cite{zsthinning}.
 
 \subsection{Vertical projection histogram}
-Now that we have the skeletonized word image, we can proceed with determining the potential segmentation points.
-We do this by taking the vertical projection histogram of this skeletonized word image.
-With the vertical projection executed, we now have the summation of every collumn of the word image.
-This is done in a way that black pixels are represented by a $1$ value, and white pixels are represented by a $0$ value.
-Now we can regard the collumns with a total pixel count of zero or one pixel as potential segmentation points, as we can interpret columns containing only a single pixel to represent a connection between different characters, and collumns without pixels to represent spacing between characters.
-With these points as our potential segmentation points, we need to separate the correct segmentation points from the incorrect ones.
-For this we will use oversegmentation correction techniques.
+Using the skeletonized word image, we now create a vertical projection histogram.
+This vertical projection histogram takes the sum of every collumn in the image, counting black pixels as a $1$ value, and white pixels as a $0$ value.
+
+We now regard the collumns with a total pixel count of zero or one pixel as potential segmentation points.
+We interpret these collumn to be the whitespace in between characters or the line connecting two characters to each other.
+This will inevitably cause oversegmentation in our word image.
+The next step will be to filter the potential segmentation points, and remove the incorrect ones through postprocessing.
 

### 2017-12-17 Report/paper.tex
finished rework

@@
 Firstly we remove the contours of which the bounding box is contained by the bounding box of a second contour.
-We can do this because the text segmentation process will return the word in the image contained by its bounding box, so we will not change  % TODO:: NOTATION
+We can do this because the text segmentation process will return the word in the image contained by its biggest bounding box, so we will not change the returned word image through this step  % TODO:: NOTATION
 
@@
 
-\subsection{Normalization} % TODO:: necessary? never taken advantage of? maybe for split points count etc???
+\subsection{Normalization}
 The first action is again normalizing the word image.
@@
 \subsection{Oversegmentation correction}
-In this section we provide techniques to verify the correctness of these segmentation points.
+In this section we provide techniques to verify the correctness of segmentation points.
 \cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
@@
 The main oversegmentation correction is handled by our neural network.
-However some simple solutions can be added to remove unwanted potential segmentation points throught deterministic processing.
+However some simple preprocessing is added to remove obvious unwanted segmentation points.
 
-The first filter we utilize here, is one that removes the segmentation points at the start and the end of the word image.
-These segmentation points are an artifact of the used segmentation approach, and can be removed without information loss, as the alphabet does not contain letters that convert to a single pixel line after skeletonization.
+We first remove removes the segmentation points found at the start and the end of the word image.
+These segmentation points are an artifact of the used segmentation approach, as can be observed in Figure Figure \ref{fig:char_segm}.
+These points can be removed without information loss, as the alphabet does not contain letters that entirely convert to a single pixel line after skeletonization.
 
-A second filter groups consequent potential segmentation points.
-The occurence of these consequent points is another artifact from our segmentation approach.
-From these grouped points, the collumns beloning to these x-coordinates are checked to be empty or containing a single pixel in the vertical projection histogram.
-if the group contains collumns that are empty on the histogram, only the middle segmentation point from this group is retained as potential segmentation point.
-If none of the potential segmentation points has an empty collumn on the histogram, the retained segmentation point is the middle points from the group.
+Subsequently we group consequent segmentation points.
+The occurence of these consequent segmentation points is another artifact from our segmentation approach, as can be seen in Figure \ref{fig:char_segm}.
 
-A third filter is that groups of consequent potential segmentation points, separated by collumns containing less pixels than a certain threshold value $maxSumThresh$, are grouped together.
-The value chosen for this threshold is 3.
-This filter remedies the segmentation of slated lines between the characters.
+From these groups of points, we take the center point as the new potential segmentation points.
+However if one or more of the segmentation points in the group stems from empty collumns, the new segmentation point is chosen as the center point of the points representing the empty collumns in the word image.
 
-This value for this threshold is pretty conservative, and could be further investigated for a more optimal rule based approach.
-More rule based filter could be added for more optimal processing of the potential segmentation points.
-However the focus the segmentation correction is the neural network used to remove incorrect segmenatation points.
+Finally we aggregate groups of consequent potential segmentation points, separated by collumns containing less pixels than a certain threshold value $maxSumThresh$.
+We have chosen a conservative value of $3px$ for this threshold.
+This remedies the segmentation of slated lines inside characters, which can be an artifact of the skeletonization algirithm used.
 
 A technique was tested using the feedback of the chracter recognition network to remove incorrect segmentation points.
-However this turned out not to be an effective technique, because the neural network tends to provide random probabilities with a high degree of certainty when multiple characters were fed to the network in the same image.
-Since it is impossible to validate these responses from the neural network, this technique is not used in the final version of our oversegmentation correction.
+However this turned out not to be ineffective, because the neural network tends to provide random probabilities with a high degree of certainty when images of multiple characters together were fed to the nerual network.
+Since it is impossible to validate these responses from the neural network, this approach was abandoned and is not used in the final version.
+
+More rule based filter could be added for more optimal processing of the potential segmentation points.
+However the focus of our segmentation correction is the neural network used to remove incorrect segmenatation points.
 

### 2017-12-17 Report/paper.tex
proofreading of section 3

--- 2017-12-17 Report/paper.tex
@@
 In this section we explain the first step in the process of extracting individual characters from a text image: i.e. word extraction.
-Word extraction from text images has many possible approaches, but in this paper we focus solely on the word contouring approach, pointed out by following paper \cite{WordSegm} to be a successfull word extraction approach.
-
-With this approach we extract the contours of the words in a given text image. Through postprocessing techniques we can rearrange these extracted words into lines.
+Word extraction from text images has many possible approaches, but in this paper we focus solely on the word contouring approach, pointed out by following paper \cite{WordSegm} to be a successful word extraction method.
+With this method we extract the contours of the words in a given text image. Through postprocessing techniques we can rearrange these extracted words into lines.
 The proposed approach focusses solely on text images, and does not take into account non-textual areas found in the input image.
@@
 We start with binarizing the text image.
-
 To achieve this there are a lot of thresholding algorithms to choose from, but the one we use is a combination of Gaussian blur in combination with Otsu's thresholding algorithm.
-
+\cite{Otsu79}
 The Gaussian blur will smoothen the noise in the image and smoothen the edges of the text.
-%TODO:: Otsu reference
 Otsu's tresholding algorithm will deliver adequate result for noisy text documents.
 The choise of tresholding algorithm depends mostly on the intended purposes of the algorithm.
-Additional noise removal techniques and other preprocessing techniques may be required for a better result when processing very noise-heavy images.
+Additional noise removal techniques and other preprocessing techniques may be required for a better result when processing images with much noise.
 
@@
 Our text segmentation approach is based on finding contours in the text image.
-We can find these contours using %TODO:: reference paper contouring suzuki85
+We can find these contours using the algorithm \cite{Suzuki85}.
 Through postprocessing we can recover the original words from these extracted contours.
@@
 \subsection{Postprocessing}
-We can remove the incorrect contours generated by the contouring algorithm through postprocessing.
-Secondly we will reconnect words that were split because they are not written in a single stroke.
+We remove the incorrect contours generated by the contouring algorithm through postprocessing.
+Next we reconnect words that are split because they are not written in a single stroke.
 
 \subsubsection{Contour removal}
-The contouring algorthm can create incorrect contours because of left-over noise in the image.
-The tool we will use to remove these incorrect contours is the bounding box we can draw around a contour.
-This bounding box is the smallest rectangle containing said contour, and it will allow us to easily compare contour positions in the text image.
+The contouring algorithm can create incorrect contours because of left-over noise in the image.
+The tool we will use to remove these contours is a bounding box drawn around the contour.
+This bounding box is the smallest rectangle containing the contour, it allows us to easily compare positions of contours in the text image.
 
-Firstly we remove the contours of which the bounding box is contained by the bounding box of a second contour.
-We can do this because the text segmentation process will return the word in the image contained by its bounding box, so we will not change  % TODO:: NOTATION
-
-Secondly we remove contours that are too small to contain parts of a word.
-The contouring algorithm will generate contours of noise and punctuation.
-These contours we do not want to process, so we will filter them out.
-
-This is done by taking the average height of all the found contours in the text.
-Iteratively we remove remove the contours with a height lesser than the a predetermined threshold,
-\vspace{-5px}
-\begin{equation}
-    threshold = averageHeight * removalMultiplier
-\end{equation}
-with the $removalMultiplier$ initialized on $2 / 3$ as chosen value.
+Firstly we remove the contours of which the bounding box is inside the box of another contour.
+This is possible as text segmentation returns the word in the image contained by its biggest bounding box. 
+Secondly we remove contours that are too small to contain possible word parts. These contours consist of noise and punctuation.
+These contours are neglected by taking the average height of all the found contours in the text.
+We remove all of the contours with a height smaller than $average height * (2/3)$.
 
 \subsubsection{Contour combination}
-With the excess contours removed, we begin to reconstruct the original words from the text.
-We do this by combining multiple contours into one where a word was separated into multiple parts.
-We have created a straightforward overlapping algorithm, that combines word segments based on the overlap between the bounding box of their contours.
+The leftover contours are used to to reconstruct the original words from the text.
+This is done by combining multiple contours which consist of parts of words belonging to the same word. 
+We have created an algorithm that combines word segments. It is uses the overlap between the bounding box of their contours to find possible combinations.
 
+% Hoe heet die parameter? Wat is de waarde van de parameter. Plaats de namen, de waardes en hun uitleg samen, plaats hier geen algoritme/formule tussen. 
 We have a parameter for the allowed gap between parts of the same word, and one for the minimal y-axis overlap of both their boundingboxes.
 
-\begin{equation} %TODO:: FIX INDENTATION
+\begin{equation} %TODO:: FIX INDENTATION, Plaats zoiets in algorithmic
     if \\ 
@@
 \end{equation}
-The optimal parameters we found are:
-
-\noindent 
-$averageCharacterCount = 8$ and
-
-\noindent
-$requiredYaxisOverlap = 1/2$
-
-With these conditions fulfilled, we combine both contours into one single contour, containing the found word.
+The optimal parameters we found are $averageCharacterCount=8$ and $requiredYaxisOverlap=1/2$
+Given that these conditions are fulfilled, we can combine both contours into one single contour, containing the found word.
 
@@
 \subsection{Line selection}
-With the words extracted from the text, the next goal is to reconstruct the original text.
-We will need to order these words into lines 
-For the line selection process, an iterative algorithm, Algorithm \ref{alg:linereconstr}, is used.
-
-This algorithm reorders the found words in lines, and return these lines as ordered in the original text image.
+As the ordening of the words % TODO, verklaar waarom die ordening weggat? 
+we place the words in lines, and orden these lines. 
+Algorithm \ref{alg:linereconstr} shows the pseudocode of our line selection process.
+The algorithm reorders the words in lines and orders these lines as ordered in the original text.
 

### 2017-12-17 Report/paper.tex
r

--- 2017-12-17 Report/paper.tex
@@
 
-To achieve this there are a lot of thresholding algorithms to choose from, but the one we use is a combination of Gaussian blur in combination with Otsu's thresholding algorithm.
+To achieve this there are a lot of thresholding algorithms to choose from, but the one we use is a combination of Gaussian blur in combination with Otsu's method.
+More information about Otsu's method can be found in the following paper \cite{otsu}
 
 The Gaussian blur will smoothen the noise in the image and smoothen the edges of the text.
-%TODO:: Otsu reference
-Otsu's tresholding algorithm will deliver adequate result for noisy text documents.
+Otsu's method will deliver adequate result for noisy text documents.
 The choise of tresholding algorithm depends mostly on the intended purposes of the algorithm.
@@
 The word image is binarized.
-This is again done with a combination of Gaussian blur and Otsu's thresholding algorithm.
+This is again done with a combination of Gaussian blur and Otsu's method.
 The binarization is needed for the next step, because we can only apply a skeletonization algorithm to a binary image.

### 2017-12-17 Report/paper.tex

--- 2017-12-17 Report/paper.tex
@@
 \subsection{Postprocessing}
-We can remove the incorrect contours generated by the contouring algorithm through postprocessing.
-Secondly we will reconnect words that were split because they are not written in a single stroke.
+We remove the incorrect contours generated by the contouring algorithm through postprocessing.
+Next we reconnect words that are split because they are not written in a single stroke.
 
 \subsubsection{Contour removal}
-The contouring algorthm can create incorrect contours because of left-over noise in the image.
-The tool we will use to remove these incorrect contours is the bounding box we can draw around a contour.
-This bounding box is the smallest rectangle containing said contour, and it will allow us to easily compare contour positions in the text image.
+The contouring algorithm can create incorrect contours because of left-over noise in the image.
+The tool we will use to remove these contours is a bounding box drawn around the contour.
+This bounding box is the smallest rectangle containing the contour, it allows us to easily compare positions of contours in the text image.
 
-Firstly we remove the contours of which the bounding box is contained by the bounding box of a second contour.
-We can do this because the text segmentation process will return the word in the image contained by its biggest bounding box, so we will not change the returned word image through this step  % TODO:: NOTATION
-
-Secondly we remove contours that are too small to contain parts of a word.
-The contouring algorithm will generate contours of noise and punctuation.
-These contours we do not want to process, so we will filter them out.
-
-This is done by taking the average height of all the found contours in the text.
-Iteratively we remove remove the contours with a height lesser than the a predetermined threshold,
-\vspace{-5px}
-\begin{equation}
-    threshold = averageHeight * removalMultiplier
-\end{equation}
-with the $removalMultiplier$ initialized on $2 / 3$ as chosen value.
+Firstly we remove the contours of which the bounding box is inside the box of another contour.
+This is possible as text segmentation returns the word in the image contained by its biggest bounding box. 
+Secondly we remove contours that are too small to contain possible word parts. These contours consist of noise and punctuation.
+These contours are neglected by taking the average height of all the found contours in the text.
+We remove all of the contours with a height smaller than $average height * (2/3)$.
 
 \subsubsection{Contour combination}
-With the excess contours removed, we begin to reconstruct the original words from the text.
-We do this by combining multiple contours into one where a word was separated into multiple parts.
-We have created a straightforward overlapping algorithm, that combines word segments based on the overlap between the bounding box of their contours.
+The leftover contours are used to to reconstruct the original words from the text.
+This is done by combining multiple contours which consist of parts of words belonging to the same word. 
+We have created an algorithm that combines word segments. It is uses the overlap between the bounding box of their contours to find possible combinations.
 
+% Hoe heet die parameter? Wat is de waarde van de parameter. Plaats de namen, de waardes en hun uitleg samen, plaats hier geen algoritme/formule tussen. 
 We have a parameter for the allowed gap between parts of the same word, and one for the minimal y-axis overlap of both their boundingboxes.
 
-\begin{equation} %TODO:: FIX INDENTATION
+\begin{equation} %TODO:: FIX INDENTATION, Plaats zoiets in algorithmic
     if \\ 
@@
 
-\subsection{Line reconstruction}
-With the words extracted from the text, the next goal is to reconstruct the original text.
-We will need to order these words into lines 
-For the line selection process, an iterative algorithm, Algorithm \ref{alg:linereconstr}, is used.
-
-This algorithm reorders the found words in lines, and return these lines as ordered in the original text image.
+\subsection{Line selection}
+As the ordening of the words % TODO, verklaar waarom die ordening weggat? 
+we place the words in lines, and orden these lines. 
+Algorithm \ref{alg:linereconstr} shows the pseudocode of our line selection process.
+The algorithm reorders the words in lines and orders these lines as ordered in the original text.
 

### 2017-12-17 Report/paper.tex
If statement probke

@@
 
+\begin{algorithm}[tb]
+    \caption{Contour combination algorithm}
+    \label{alg:overlap}
+\begin{algorithmic}
+    \STATE {\bfseries Input:} $contour1$, $contour2$
+    \STATE $xGap$ = calculateXaxisDistance $contour1$ $contour2$
+    \STATE $yOverlap$ = calculateYoverlap $contour1$ $contour2$
+    %TODO ::\IF {$xGap < averageWordWidth * (1 / (averageCharacterCount * 2)) $ {\bfseries and} $yOverlap > min(word1.height, word2.height) * requiredYaxisOverlap$}
+    \STATE combineContours $contour1$ $contour2$
+\end{algorithmic}
+The chosen paramters are:
+\\
+$averageCharacterCount = 8$ 
+\\
+$requiredYaxisOverlap = 1/2$
+\end{algorithm}
+
 \subsubsection{Contour combination}
@@
 This is done by combining multiple contours which consist of parts of words belonging to the same word. 
-We have created an algorithm that combines word segments. It is uses the overlap between the bounding box of their contours to find possible combinations.
-
-% Hoe heet die parameter? Wat is de waarde van de parameter. Plaats de namen, de waardes en hun uitleg samen, plaats hier geen algoritme/formule tussen. 
-We have a parameter for the allowed gap between parts of the same word, and one for the minimal y-axis overlap of both their boundingboxes.
-
-\begin{equation} %TODO:: FIX INDENTATION, Plaats zoiets in algorithmic
-    if \\ 
-        wordXaxisGap < \\
-            averageWordWidth * (1 / (averageCharacterCount * 2) )\\
-        and \\
-        wordYaxisOverlap > \\
-        min(word1.height, word2.height) * requiredYaxisOverlap\\
-    then \\
-        combineWords
-\end{equation}
-The optimal parameters we found are $averageCharacterCount=8$ and $requiredYaxisOverlap=1/2$
-Given that these conditions are fulfilled, we can combine both contours into one single contour, containing the found word.
+We have created an algorithm \ref{alg:overlap} that decides when to combine two word segments into a single word.
 

### 2017-12-17 Report/paper.tex
a

@@
 \subsection{Line selection}
-As the ordening of the words % TODO, verklaar waarom die ordening weggat? 
-we place the words in lines, and orden these lines. 
+Because the contour finding algorithm has no particular order in which it saves the found contours, we have to reorder the found words in their original lines.
 Algorithm \ref{alg:linereconstr} shows the pseudocode of our line selection process.

### 2017-12-17 Report/paper.tex
fix some issues in 4

--- 2017-12-17 Report/paper.tex
@@
 \subsection{Binarization}
+\label{sec:binartext}
+
 We start with binarizing the text image.
@@
 The next step in the process is to extract the character images from the found word images.
-Again, there is a choice between several different approaches.
-In this paper we focus solely on a word oversegmentation approach to character extraction, through skeletonization and using a vertical projection profile.
-The reasoning begind this can be found in the following papers \cite{CharSegm} and \cite{CharSegmOld}.
+Possible approaches and their advantages are discussed by \cite{CharSegm} and \cite{CharSegmOld}.
+We choose the word oversegmentation approach for character extraction.
+Word oversegmentation is achieved through skeletonization and using a vertical projection profile which is discussed in consequently \ref{sec:skel} and \ref{sec:vertproj}.
+After oversegmenting the word, we find the actual splitpoints by removing unnecessary splitpoints as explained in \ref{sec:postprocesscharseg}
 
-After first oversegmenting the word, we will recover the original character through postprocessing steps combining different word segments.
-
-\subsection{Normalization}
-The first action is again normalizing the word image.
-We will resuze the image to a fixed height of $100px$.
-This will result in more consistency in rule-based post processing steps.
-
-\subsection{Binarization}
-The word image is binarized.
-This is again done with a combination of Gaussian blur and Otsu's thresholding algorithm.
-The binarization is needed for the next step, because we can only apply a skeletonization algorithm to a binary image.
+\subsection{Preprocessing}
+We resize the image to a fixed height of 100 pixels. 
+Next we binarize the image using a combination of Gaussian blur and Otsu's method. 
+The binarization method is identical to \ref{sec:binartext} but included to allow segmentation of words which are not part of a text. Binarization is essential as it is necessary for skeletonization.
 
 \subsection{Skeletonization}
-Now we will skeletonize our word image.
-This will normalize the thickness of the strokes to a single pixel.
-This allows us to execute our Verical projection profile and interpret the results easily.
-The skeletonization algorithm used is the scipy adaptation of the Zhang Suen thinning algorithm \cite{zsthinning}.
+\label{sec:skel}
+Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
+allows us to create a vertical projection profile. The skeletonization algorithm used is the scipy adaptation of \cite{zsthinning}. % TODO reference scipy
 
 \subsection{Vertical projection histogram}
+\label{sec:vertproj}
+
 Using the skeletonized word image, we now create a vertical projection histogram.
@@
 \subsection{Oversegmentation correction}
+\label{sec:postprocesscharseg}
+
 In this section we provide techniques to verify the correctness of segmentation points.

### 2017-12-17 Report/paper.tex
a
fix
additional fixes

@@
 
-Using the skeletonized word image, we now create a vertical projection histogram.
-This vertical projection histogram takes the sum of every collumn in the image, counting black pixels as a $1$ value, and white pixels as a $0$ value.
-
-We now regard the collumns with a total pixel count of zero or one pixel as potential segmentation points.
-We interpret these collumn to be the whitespace in between characters or the line connecting two characters to each other.
+Using the skeletonized word image, we create a vertical projection histogram.
+This vertical projection histogram counts the black pixels of every column in the image.
+We regard the columns with a total pixel count of zero or one as potential segmentation points.
+We interpret these columns as whitespace in between characters or as a line connecting two characters to each other.
 This will inevitably cause oversegmentation in our word image.
-The next step will be to filter the potential segmentation points, and remove the incorrect ones through postprocessing.
+The next step filters incorrect segmentation points caused by oversegmentation.
 
@@
 \subsubsection{Rule based techniques}
-The main oversegmentation correction is handled by our neural network.
+Most of the oversegmentation correction is done by our neural network.
 However some simple preprocessing is added to remove obvious unwanted segmentation points.

### 2017-12-17 Report/paper.tex

@@
     \STATE {\bfseries Input:} $contour1$, $contour2$
-    \STATE $xGap$ = calculateXaxisDistance $contour1$ $contour2$
-    \STATE $yOverlap$ = calculateYoverlap $contour1$ $contour2$
-    %TODO ::\IF {$xGap < averageWordWidth * (1 / (averageCharacterCount * 2)) $ {\bfseries and} $yOverlap > min(word1.height, word2.height) * requiredYaxisOverlap$}
+    \STATE $xGap =$ calculateXaxisDistance $contour1$ $contour2$
+    \STATE $yOverlap =$ calculateYoverlap $contour1$ $contour2$
+    %TODO:\IF {$xGap < averageWordWidth * (1 / (averageCharacterCount * 2)) $ {\bfseries and} $yOverlap > min(word1.height, word2.height) * requiredYaxisOverlap$}
         \STATE combineContours $contour1$ $contour2$
@@
 Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
-allows us to create a vertical projection profile. The skeletonization algorithm used is the scipy adaptation of \cite{zsthinning}. % TODO reference scipy
+allows us to create a vertical projection profile. We use the skeletonization algorithm of \cite{zsthinning}.
 
@@
 In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
+An example of applying these correction techniques to an image is shown in \ref{fig:char_segm}.
 
 \subsubsection{Rule based techniques}
-Most of the oversegmentation correction is done by our neural network.
-However some simple preprocessing is added to remove obvious unwanted segmentation points.
 
-We first remove removes the segmentation points found at the start and the end of the word image.
-These segmentation points are an artifact of the used segmentation approach, as can be observed in Figure Figure \ref{fig:char_segm}.
+First we remove segmentation points found at the start and the end of the word image.
 These points can be removed without information loss, as the alphabet does not contain letters that entirely convert to a single pixel line after skeletonization.
+Next consequent segmentation points are grouped. The center of these points is considered as segmentation point.
+Unless if one or more of the segmentation points in the group stems from empty columns, the new segmentation point is chosen as the center of the points representing the empty columns in the word image.
+Finally we aggregate groups of consequent potential segmentation points, separated by columns containing less pixels than a threshold value 3.
+This remedies the segmentation of slanted lines inside characters.
 
-Subsequently we group consequent segmentation points.
-The occurence of these consequent segmentation points is another artifact from our segmentation approach, as can be seen in Figure \ref{fig:char_segm}.
-
-From these groups of points, we take the center point as the new potential segmentation points.
-However if one or more of the segmentation points in the group stems from empty collumns, the new segmentation point is chosen as the center point of the points representing the empty collumns in the word image.
-
-Finally we aggregate groups of consequent potential segmentation points, separated by collumns containing less pixels than a certain threshold value $maxSumThresh$.
-We have chosen a conservative value of $3px$ for this threshold.
-This remedies the segmentation of slated lines inside characters, which can be an artifact of the skeletonization algirithm used.
-
-A technique was tested using the feedback of the chracter recognition network to remove incorrect segmentation points.
-However this turned out not to be ineffective, because the neural network tends to provide random probabilities with a high degree of certainty when images of multiple characters together were fed to the nerual network.
-Since it is impossible to validate these responses from the neural network, this approach was abandoned and is not used in the final version.
-
-More rule based filter could be added for more optimal processing of the potential segmentation points.
-However the focus of our segmentation correction is the neural network used to remove incorrect segmenatation points.
+A technique was tested using the output of the character recognition network to remove incorrect segmentation points.
+This turned out to be ineffective because the neural network tends to provide high probabilities given images containing multiple characters.
+These probabilities are not meaningful and cannot be used as feedback for segmentation correctness.
 
@@
 Our recognition system cannot be used for real-world applications.
-Every individual technique has a certain error, which increases the difficulty of the next step, e.g.
+Every individual technique has a certain error, which increases the difficulty of the next step; e.g.
 if word segmentation contains mistakes, character recognition becomes much harder.
@@
 
-The character segmentation process is inadequate as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
-More postprocessing can be used through rule based filters and a more robust neural network to correct the found segmentation points, but this will never reach an optimal solution through rule based filters alone because of the ambiguity of this problem.
-
 Finally, other approaches can also be considered.
@@
 These methods are mostly based on hidden Markov models and neural networks.
+Due to the ambiguous nature of the word segmentation problem, a segmentation-free approach results in much better results.
 This concept is explained and applied by \cite{presham}.

### 2017-12-17 Report/paper.tex
algorithm

--- 2017-12-17 Report/paper.tex
@@
 \subsection{Binarization}
-\label{sec:binartext}
-
 We start with binarizing the text image.
@@
 \begin{algorithmic}
-    \STATE {\bfseries Input:} $contour1$, $contour2$
-    \STATE $xGap$ = calculateXaxisDistance $contour1$ $contour2$
-    \STATE $yOverlap$ = calculateYoverlap $contour1$ $contour2$
-    %TODO ::\IF {$xGap < averageWordWidth * (1 / (averageCharacterCount * 2)) $ {\bfseries and} $yOverlap > min(word1.height, word2.height) * requiredYaxisOverlap$}
+    \STATE {\bfseries Input:}  $contour1$, $contour2$, $text$
+    \STATE $xGap =$ calculateXaxisDistance $contour1$, $contour2$
+    \STATE $yOverlap =$ calculateYoverlap $contour1$, $contour2$
+    \STATE $avgWidth =$ getAverageWordWith $text$
+    \IF{ 
+    $xGap < avgWidth $ \\ $* (1 / (averageCharacterCount * 2))$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height)$ \\ $*  requiredYaxisOverlap $ 
+    }
     \STATE combineContours $contour1$ $contour2$
+    \ENDIF
 \end{algorithmic}
@@
 The next step in the process is to extract the character images from the found word images.
-Possible approaches and their advantages are discussed by \cite{CharSegm} and \cite{CharSegmOld}.
-We choose the word oversegmentation approach for character extraction.
-Word oversegmentation is achieved through skeletonization and using a vertical projection profile which is discussed in consequently \ref{sec:skel} and \ref{sec:vertproj}.
-After oversegmenting the word, we find the actual splitpoints by removing unnecessary splitpoints as explained in \ref{sec:postprocesscharseg}
+Again, there is a choice between several different approaches.
+In this paper we focus solely on a word oversegmentation approach to character extraction, through skeletonization and using a vertical projection profile.
+The reasoning begind this can be found in the following papers \cite{CharSegm} and \cite{CharSegmOld}.
 
-\subsection{Preprocessing}
-We resize the image to a fixed height of 100 pixels. 
-Next we binarize the image using a combination of Gaussian blur and Otsu's method. 
-The binarization method is identical to \ref{sec:binartext} but included to allow segmentation of words which are not part of a text. Binarization is essential as it is necessary for skeletonization.
+After first oversegmenting the word, we will recover the original character through postprocessing steps combining different word segments.
+
+\subsection{Normalization}
+The first action is again normalizing the word image.
+We will resuze the image to a fixed height of $100px$.
+This will result in more consistency in rule-based post processing steps.
+
+\subsection{Binarization}
+The word image is binarized.
+This is again done with a combination of Gaussian blur and Otsu's method.
+The binarization is needed for the next step, because we can only apply a skeletonization algorithm to a binary image.
 
 \subsection{Skeletonization}
-\label{sec:skel}
-Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
-allows us to create a vertical projection profile. The skeletonization algorithm used is the scipy adaptation of \cite{zsthinning}. % TODO reference scipy
+Now we will skeletonize our word image.
+This will normalize the thickness of the strokes to a single pixel.
+This allows us to execute our Verical projection profile and interpret the results easily.
+The skeletonization algorithm used is the Zhang Suen thinning algorithm \cite{zsthinning}.
 
 \subsection{Vertical projection histogram}
-\label{sec:vertproj}
-
 Using the skeletonized word image, we now create a vertical projection histogram.
@@
 \subsection{Oversegmentation correction}
-\label{sec:postprocesscharseg}
-
 In this section we provide techniques to verify the correctness of segmentation points.

### 2017-12-17 Report/paper.tex
algorithm finish typeset, added last references

@@
     \STATE {\bfseries Input:}  $contour1$, $contour2$, $text$
-    \STATE $xGap =$ calculateXaxisDistance $contour1$, $contour2$
-    \STATE $yOverlap =$ calculateYoverlap $contour1$, $contour2$
-    \STATE $avgWidth =$ getAverageWordWith $text$
+    \STATE $xGap =$ XAxisDistance $contour1$, $contour2$
+    \STATE $yOverlap =$ YAxisOverlap $contour1$, $contour2$
     \IF{ 
-    $xGap < avgWidth $ \\ $* (1 / (averageCharacterCount * 2))$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height)$ \\ $*  requiredYaxisOverlap $ 
+    \\ 
+    \hspace{8px} $xGap < text$.averageWordWith 
+    \\
+    \hspace{8px} $* $ \hspace{2px} $(1 / (averageCharacterCount * 2))$ 
+    \\
+    {\bfseries and} 
+    \\ 
+    \hspace{8px} $yOverlap > min(contour1.height, contour2.height)$ 
+    \\
+    \hspace{8px} $* $ \hspace{2px} $  requiredYaxisOverlap $ 
     }
+    \vspace{4px}
     \STATE combineContours $contour1$ $contour2$
@@
    \REPEAT
-        \STATE $highestWord = findHighestWord $
-        \STATE $newline = [highestWord]$
-        \STATE Initialize $noChange = true$.
-        \STATE $avgY= newline$.findAverageYpos
+        \STATE $newline = [ $wordWithHighestYPosition$ ]$
+        \STATE $avgY= newline$.averageYPosition
         \REPEAT
@@
             \STATE $line$.add $word$
-            \STATE $noChange = false$
             \ENDIF
         \ENDFOR
-        \UNTIL{$noChange = true$}
+        \UNTIL{NotChangedLastIteration $line$}
         \STATE $textlines$.add $newline$
@@
 The text to word segmentation process can be improved upon and made more robust.
-Advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements.
+More advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements. \cite{textsegm}
 Algorithms that provide better noise reduction can be added to handle text images with more noise,
@@
 
-The character segmentation process is inadequate as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
-More postprocessing can be used through rule based filters and a more robust neural network to correct the found segmentation points, but this will never reach an optimal solution through rule based filters alone because of the ambiguity of this problem.
+The character segmentation process is insufficient as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
+More postprocessing can be used through rule based techniques and a more robust neural network to correct the found segmentation points. \cite{segmtechniques}
 

### 2017-12-17 Report/paper.tex
done

@@
 In this section we provide techniques to verify the correctness of segmentation points.
-\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
-In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
+\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule-based techniques, cost function optimization based methods and machine learning based procedures.
+In this paper we discuss machine learning based approaches precedented by a few rule-based techniques.
 
@@
 
-More rule based filter could be added for more optimal processing of the potential segmentation points.
+More rule-based filter could be added for more optimal processing of the potential segmentation points.
 However the focus of our segmentation correction is the neural network used to remove incorrect segmenatation points.
@@
 The text to word segmentation process can be improved upon and made more robust.
-More advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements. \cite{textsegm}
-Algorithms that provide better noise reduction can be added to handle text images with more noise,
-and experiments with rotations of the text can provide us with a more robust system to tackle text images containing heavily slanted text lines.
+More advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements.
+A comparison of different text segmentation techniques is done by \cite{textsegm}.
 
 The character segmentation process is insufficient as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
-More postprocessing can be used through rule based techniques and a more robust neural network to correct the found segmentation points. \cite{segmtechniques}
+More postprocessing can be used through rule-based techniques and a more robust neural network to correct the found segmentation points.
+Several rule-based techniques are proposed by \cite{segmtechniques}.
 

### 2017-12-17 Report/paper.tex
small errors

@@
 
-After first oversegmenting the word, we will recover the original character through postprocessing steps combining different word segments.
+After first oversegmenting the word, we recover the original character through postprocessing steps combining different word segments.
 
@@
 The first action is again normalizing the word image.
-We will resuze the image to a fixed height of $100px$.
-This will result in more consistency in rule-based post processing steps.
+We resize the image to a fixed height of $100px$.
+This result in more consistency in rule-based postprocessing steps.
 
@@
 \subsection{Skeletonization}
-Now we will skeletonize our word image.
-This will normalize the thickness of the strokes to a single pixel.
+Now we skeletonize our word image.
+This normalizes the thickness of the strokes to a single pixel.
 This allows us to execute our Verical projection profile and interpret the results easily.

### 2017-12-18 Report/paper.tex
fix indentation

--- 2017-12-17 Report/paper.tex
@@
     \STATE {\bfseries Input:}  $contour1$, $contour2$, $text$
-    \STATE $xGap =$ XAxisDistance $contour1$, $contour2$
-    \STATE $yOverlap =$ YAxisOverlap $contour1$, $contour2$
+    \STATE $xGap =$ calculateXaxisDistance $contour1$, $contour2$
+    \STATE $yOverlap =$ calculateYoverlap $contour1$, $contour2$
+    \STATE $avgWidth =$ getAverageWordWith $text$
     \IF{ 
-    \\ 
-    \hspace{8px} $xGap < text$.averageWordWith 
-    \\
-    \hspace{8px} $* $ \hspace{2px} $(1 / (averageCharacterCount * 2))$ 
-    \\
-    {\bfseries and} 
-    \\ 
-    \hspace{8px} $yOverlap > min(contour1.height, contour2.height)$ 
-    \\
-    \hspace{8px} $* $ \hspace{2px} $  requiredYaxisOverlap $ 
+    $xGap < avgWidth $ \\ $* (1 / (averageCharacterCount * 2))$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height)$ \\ $*  requiredYaxisOverlap $ 
     }
-    \vspace{4px}
     \STATE combineContours $contour1$ $contour2$
@@
    \REPEAT
-        \STATE $newline = [ $wordWithHighestYPosition$ ]$
-        \STATE $avgY= newline$.averageYPosition
+        \STATE $highestWord = findHighestWord $
+        \STATE $newline = [highestWord]$
+        \STATE Initialize $noChange = true$.
+        \STATE $avgY= newline$.findAverageYpos
         \REPEAT
@@
             \STATE $line$.add $word$
+            \STATE $noChange = false$
             \ENDIF
         \ENDFOR
-        \UNTIL{NotChangedLastIteration $line$}
+        \UNTIL{$noChange = true$}
         \STATE $textlines$.add $newline$
@@
 The next step in the process is to extract the character images from the found word images.
-Again, there is a choice between several different approaches.
-In this paper we focus solely on a word oversegmentation approach to character extraction, through skeletonization and using a vertical projection profile.
-The reasoning begind this can be found in the following papers \cite{CharSegm} and \cite{CharSegmOld}.
+Possible approaches and their advantages are discussed by \cite{CharSegm} and \cite{CharSegmOld}.
+We choose the word oversegmentation approach for character extraction.
+Word oversegmentation is achieved through skeletonization and using a vertical projection profile which is discussed in consequently \ref{sec:skel} and \ref{sec:vertproj}.
+After oversegmenting the word, we find the actual splitpoints by removing unnecessary splitpoints as explained in \ref{sec:postprocesscharseg}
 
-After first oversegmenting the word, we will recover the original character through postprocessing steps combining different word segments.
-
-\subsection{Normalization}
-The first action is again normalizing the word image.
-We will resuze the image to a fixed height of $100px$.
-This will result in more consistency in rule-based post processing steps.
-
-\subsection{Binarization}
-The word image is binarized.
-This is again done with a combination of Gaussian blur and Otsu's method.
-The binarization is needed for the next step, because we can only apply a skeletonization algorithm to a binary image.
+\subsection{Preprocessing}
+We resize the image to a fixed height of 100 pixels. 
+Next we binarize the image using a combination of Gaussian blur and Otsu's method. 
+The binarization method is identical to \ref{sec:binartext} but included to allow segmentation of words which are not part of a text. Binarization is essential as it is necessary for skeletonization.
 
 \subsection{Skeletonization}
-Now we will skeletonize our word image.
-This will normalize the thickness of the strokes to a single pixel.
-This allows us to execute our Verical projection profile and interpret the results easily.
-The skeletonization algorithm used is the Zhang Suen thinning algorithm \cite{zsthinning}.
+\label{sec:skel}
+Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
+allows us to create a vertical projection profile. We use the skeletonization algorithm of \cite{zsthinning}.
 
 \subsection{Vertical projection histogram}
-Using the skeletonized word image, we now create a vertical projection histogram.
-This vertical projection histogram takes the sum of every collumn in the image, counting black pixels as a $1$ value, and white pixels as a $0$ value.
+\label{sec:vertproj}
 
-We now regard the collumns with a total pixel count of zero or one pixel as potential segmentation points.
-We interpret these collumn to be the whitespace in between characters or the line connecting two characters to each other.
+Using the skeletonized word image, we create a vertical projection histogram.
+This vertical projection histogram counts the black pixels of every column in the image.
+We regard the columns with a total pixel count of zero or one as potential segmentation points.
+We interpret these columns as whitespace in between characters or as a line connecting two characters to each other.
 This will inevitably cause oversegmentation in our word image.
-The next step will be to filter the potential segmentation points, and remove the incorrect ones through postprocessing.
+The next step filters incorrect segmentation points caused by oversegmentation.
 
@@
 In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
+An example of applying these correction techniques to an image is shown in \ref{fig:char_segm}.
 
 \subsubsection{Rule based techniques}
-The main oversegmentation correction is handled by our neural network.
-However some simple preprocessing is added to remove obvious unwanted segmentation points.
 
-We first remove removes the segmentation points found at the start and the end of the word image.
-These segmentation points are an artifact of the used segmentation approach, as can be observed in Figure Figure \ref{fig:char_segm}.
+First we remove segmentation points found at the start and the end of the word image.
 These points can be removed without information loss, as the alphabet does not contain letters that entirely convert to a single pixel line after skeletonization.
+Next consequent segmentation points are grouped. The center of these points is considered as segmentation point.
+Unless if one or more of the segmentation points in the group stems from empty columns, the new segmentation point is chosen as the center of the points representing the empty columns in the word image.
+Finally we aggregate groups of consequent potential segmentation points, separated by columns containing less pixels than a threshold value 3.
+This remedies the segmentation of slanted lines inside characters.
 
-Subsequently we group consequent segmentation points.
-The occurence of these consequent segmentation points is another artifact from our segmentation approach, as can be seen in Figure \ref{fig:char_segm}.
-
-From these groups of points, we take the center point as the new potential segmentation points.
-However if one or more of the segmentation points in the group stems from empty collumns, the new segmentation point is chosen as the center point of the points representing the empty collumns in the word image.
-
-Finally we aggregate groups of consequent potential segmentation points, separated by collumns containing less pixels than a certain threshold value $maxSumThresh$.
-We have chosen a conservative value of $3px$ for this threshold.
-This remedies the segmentation of slated lines inside characters, which can be an artifact of the skeletonization algirithm used.
-
-A technique was tested using the feedback of the chracter recognition network to remove incorrect segmentation points.
-However this turned out not to be ineffective, because the neural network tends to provide random probabilities with a high degree of certainty when images of multiple characters together were fed to the nerual network.
-Since it is impossible to validate these responses from the neural network, this approach was abandoned and is not used in the final version.
-
-More rule based filter could be added for more optimal processing of the potential segmentation points.
-However the focus of our segmentation correction is the neural network used to remove incorrect segmenatation points.
+A technique was tested using the output of the character recognition network to remove incorrect segmentation points.
+This turned out to be ineffective because the neural network tends to provide high probabilities given images containing multiple characters.
+These probabilities are not meaningful and cannot be used as feedback for segmentation correctness.
 
@@
 The text to word segmentation process can be improved upon and made more robust.
-More advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements. \cite{textsegm}
+Advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements.
 Algorithms that provide better noise reduction can be added to handle text images with more noise,
 and experiments with rotations of the text can provide us with a more robust system to tackle text images containing heavily slanted text lines.
-
-The character segmentation process is insufficient as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
-More postprocessing can be used through rule based techniques and a more robust neural network to correct the found segmentation points. \cite{segmtechniques}
 

### 2017-12-18 Report/paper.tex
finally

@@
     \IF{ 
-    $xGap < avgWidth $ \\ $* (1 / (averageCharacterCount * 2))$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height)$ \\ $*  requiredYaxisOverlap $ 
+    $xGap < avgWidth * (1 / 8 * 2)$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height) / 2 $ 
     }
@@
 \end{algorithmic}
-The chosen paramters are:
-\\
-$averageCharacterCount = 8$ 
-\\
-$requiredYaxisOverlap = 1/2$
 \end{algorithm}
@@
 Advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements.
-Algorithms that provide better noise reduction can be added to handle text images with more noise,
-and experiments with rotations of the text can provide us with a more robust system to tackle text images containing heavily slanted text lines.
+A comparison of different text segmentation techniques is done by \cite{textsegm}.
+
+The character segmentation process can be improved upon by increasing the accuracy of the used neural network, and by adding more techniques to remove incorrect segmentations.
+Several such rule-based techniques are explored in \cite{segmtechniques}.
 

### 2017-12-18 Report/paper.tex
last hands

@@
 The contouring algorithm can create incorrect contours because of left-over noise in the image.
-The tool we will use to remove these contours is a bounding box drawn around the contour.
-This bounding box is the smallest rectangle containing the contour, it allows us to easily compare positions of contours in the text image.
+The tool we will use to remove these contours is the bounding box drawn around the contours.
+This bounding box is the smallest rectangle containing the contour. It allows us to easily compare positions of contours in the text image.
 
@@
 This is possible as text segmentation returns the word in the image contained by its biggest bounding box.
+
 Secondly we remove contours that are too small to contain possible word parts. These contours consist of noise and punctuation.
@@
     \STATE $avgWidth =$ getAverageWordWith $text$
+    \STATE $avgWordCharCount = 8$
     \IF{ 
-    $xGap < avgWidth * (1 / 8 * 2)$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height) / 2 $ 
+    $xGap < avgWidth * (1 / avgWordCharCount * 2)$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height) / 2 $ 
     }
+
     \STATE combineContours $contour1$ $contour2$
@@
 This is done by combining multiple contours which consist of parts of words belonging to the same word. 
-We have created an algorithm \ref{alg:overlap} that decides when to combine two word segments into a single word.
+For this task we have created Algorithm \ref{alg:overlap}, which decides when to combine two word segments into a single word.
 
@@
 Next we binarize the image using a combination of Gaussian blur and Otsu's method. 
-The binarization method is identical to \ref{sec:binartext} but included to allow segmentation of words which are not part of a text. Binarization is essential as it is necessary for skeletonization.
+The binarization method is identical to \ref{sec:binartext} but included to allow segmentation of words which are not part of a text.
+Binarization of the word image is essential because the skeletonization algorithm requires a binary image as input.
 
@@
 In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
-An example of applying these correction techniques to an image is shown in \ref{fig:char_segm}.
+An example of applying these correction techniques to an image is shown in Figure \ref{fig:char_segm}.
 
@@
 Unless if one or more of the segmentation points in the group stems from empty columns, the new segmentation point is chosen as the center of the points representing the empty columns in the word image.
-Finally we aggregate groups of consequent potential segmentation points, separated by columns containing less pixels than a threshold value 3.
+Finally we aggregate groups of consequent potential segmentation points, separated by columns containing less pixels than a threshold value of 3.
 This remedies the segmentation of slanted lines inside characters.
@@
 Segmentation points can now be filtered if the neural network decides a segmentation point is incorrect.
-
-\subsection{Normalization}
-The character normalization step consists of two parts.
-Firstly the character is padded untill the height equals the width.
-A precondition for the padding step is that the image is already binarized, so that we can use white borders as padding, and this will not distort the image should binarization take place in a later stage.
-The image is then resized to $44px x 44px$
-Secondly padding of $10px$ is added on all sides of the character image to create whitespace around the letter.
-We end up with a $64px x 64px$ character image.
-The extra $10px$ padding is to make sure the letters do not stick to the edges, since the Chars74K dataset we work with utilizes borders around the provided characters.
-Should these borders be removed in the training process however, this should be reflected in this step by ignoring this new padding.
 
@@
 Advanced text segmentation approaches can be utilized to extract the text segments from documents containing images and other unwanted elements.
-A comparison of different text segmentation techniques is done by \cite{textsegm}.
+fA comparison of different text segmentation techniques is done by \cite{textsegm}.
 

### 2017-12-18 Report/paper.tex
remove spelling errors

--- 2017-12-18 Report/paper.tex
@@
 Then we propose a method to recognize characters with a deep neural network.
-Finally we present postprocessing techniques to correct mispredicted text.
+Finally we present post processing techniques to correct miss predicted text.
 Our approach does not outperform state-of-the-art systems but hopes to provide a thorough elaboration for every step in the process.
@@
 \section{Introduction}
-Despite the expontially increasing amount of digital information in the contemporary world, a lot of new documents are still handwritten.
+Despite the exponentially increasing amount of digital information in the contemporary world, a lot of new documents are still handwritten.
 Handwritten documents are very inefficient as they need to be read manually in order to extract information.
@@
 First we explain how to recognise characters in Section \ref{sec:charrec}. This explanation involves the used dataset, preprocessing techniques and the deep neural network for classifying characters.
-In Section \ref{sec:segtext} we describe how to segment a text into words using a contouring algorithm with consequent postprocessing to remove unwanted noise.
-After this we present our approach to seperating the characters from words in \ref{sec:segword}.
-We use vertical projections combined with postprocessing techniques including a neural network for segmentation point evaluation.
-Postprocessing steps are presented in Section \ref{sec:postproc}, we try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
+In Section \ref{sec:segtext} we describe how to segment a text into words using a contouring algorithm with consequent post processing to remove unwanted noise.
+After this we present our approach to separating the characters from words in \ref{sec:segword}.
+We use vertical projections combined with post processing techniques including a neural network for segmentation point evaluation.
+Post processing steps are presented in Section \ref{sec:postproc}, we try to correct word recognition errors using an English dictionary and language models, n-gram models in particular.
 We conclude this paper with some experimental results, the problems we encountered combined with possible improvements.
@@
 The final dataset contains translated, orientated, sheared and scaled versions of both the original image and the eroded image.
-A visualitation of the augmentation of a picture can be seen in Figure \ref{fig:augmented}.
+A visualization of the augmentation of a picture can be seen in Figure \ref{fig:augmented}.
 
@@
 \subsubsection{Convolutional layers}
- The input is a normalized image with one color channel.
+ The input is a normalized image with one colour channel.
 This is fed to a convolutional layer.
@@
 A rectified linear unit or rectified linear neuron is a unit which uses a rectifier as linear unit.
-It's output is given by
+Its output is given by
 \begin{equation}
@@
 Finally, in order to actually train the neural network, we need some kind of weight adjustment. This will effectively improve the decision making of perceptrons and thus, the classification of images.
-The Adam optimizer effectively fulfills this goal. In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow.
+The Adam optimizer effectively fulfils this goal. In order to explain the Adam optimizer, a few other methods have to be known, a very brief explanation of these methods follow.
 First of all, stochastic gradient descent (SGD) is an iterative method that tries to find minima.
@@
 The Adam optimizer minimizes the error calculated by the cross-entropy cost function adjusted with a weight decay factor.
-Weight decay regulizes the neural network by penalizing big weights and promoting smaller weights.
+Weight decay regularizes the neural network by penalizing big weights and promoting smaller weights.
 Neural networks with this constraint tend to overfit less because small weights are less sensible to small changes of the input.
@@
 Word extraction from text images has many possible approaches, but in this paper we focus solely on the word contouring approach, pointed out by following paper \cite{WordSegm} to be a successful word extraction method.
-With this method we extract the contours of the words in a given text image. Through postprocessing techniques we can rearrange these extracted words into lines.
+With this method we extract the contours of the words in a given text image. Through post processing techniques we can rearrange these extracted words into lines.
 The proposed approach focusses solely on text images, and does not take into account non-textual areas found in the input image.
@@
 \cite{Otsu79}
-The Gaussian blur will smoothen the noise in the image and smoothen the edges of the text.
-Otsu's tresholding algorithm will deliver adequate result for noisy text documents.
-The choise of tresholding algorithm depends mostly on the intended purposes of the algorithm.
+The Gaussian blur will smoothen the noise in the image and the edges of the text.
+Otsu's thresholding algorithm will deliver adequate result for noisy text documents.
+The choise of thresholding algorithm depends mostly on the intended purposes of the algorithm.
 Additional noise removal techniques and other preprocessing techniques may be required for a better result when processing images with much noise.
@@
 We can find these contours using the algorithm \cite{Suzuki85}.
-Through postprocessing we can recover the original words from these extracted contours.
+Through post processing we can recover the original words from these extracted contours.
 
-\subsection{Postprocessing}
-We remove the incorrect contours generated by the contouring algorithm through postprocessing.
+\subsection{Post processing}
+We remove the incorrect contours generated by the contouring algorithm through post processing.
 Next we reconnect words that are split because they are not written in a single stroke.
@@
     \vspace{-5px}
-    \caption*{Postprocessing and line selection.}
+    \caption*{Post processing and line selection.}
     \end{subfigure}
@@
 Word oversegmentation is achieved through skeletonization and using a vertical projection profile which is discussed in consequently \ref{sec:skel} and \ref{sec:vertproj}.
-After oversegmenting the word, we find the actual splitpoints by removing unnecessary splitpoints as explained in \ref{sec:postprocesscharseg}
+After oversegmenting the word, we remove all unnecessary segmentation points as explained in \ref{sec:postprocesscharseg}
 
@@
 We regard the columns with a total pixel count of zero or one as potential segmentation points.
-We interpret these columns as whitespace in between characters or as a line connecting two characters to each other.
+We interpret these columns as white space in between characters or as a line connecting two characters to each other.
 This will inevitably cause oversegmentation in our word image.
@@
 We take a slightly different approach and consider a deep neural network to tackle this problem.
-This allows us to use the power of convolutational layers to extract features, instead of manually creating features (features which show the density in this case).
+This allows us to use the power of convolutional layers to extract features, instead of manually creating features (features which show the density in this case).
 We use one convolutional layer and three fully connected layers in our neural network.
@@
 
-This method allowed us to achieve an accuracy of 72 \%, which is still fairly low because of the difficulty of the problem and the small amount of traing data.
+This method allowed us to achieve an accuracy of 72 \%, which is still fairly low because of the difficulty of the problem and the small amount of training data.
 However this is better than our accuracy with an implementation of the method described by \cite{evalsplitpointsnn}.
 Segmentation points can now be filtered if the neural network decides a segmentation point is incorrect.
-
-\subsection{Normalization}
-The character normalization step consists of two parts.
-Firstly the character is padded untill the height equals the width.
-A precondition for the padding step is that the image is already binarized, so that we can use white borders as padding, and this will not distort the image should binarization take place in a later stage.
-The image is then resized to $44px x 44px$
-Secondly padding of $10px$ is added on all sides of the character image to create whitespace around the letter.
-We end up with a $64px x 64px$ character image.
-The extra $10px$ padding is to make sure the letters do not stick to the edges, since the Chars74K dataset we work with utilizes borders around the provided characters.
-Should these borders be removed in the training process however, this should be reflected in this step by ignoring this new padding.
 
@@
 This correction should allow the system to reproduce possible written text even when a step of the larger system fails.
-In order to do this we adapt natural language processing techniques as postprocessing steps.
+In order to do this we adapt natural language processing techniques as post processing steps.
 Natural language techniques can often be found in speech recognition, language generation and other systems.
@@
 
-\subsection{Vocabularium}
+\subsection{Vocabulary}
 \label{sec:voc}
 The neural network does not only produce which character is recognised, but rather a list of probabilities indicating the likelihood of an image to be a certain character.
-We can use all of this information in this postprocessing step instead of throwing that information away and naively returning the most likely character as the actual character.
+We can use all of this information in this post processing step instead of throwing that information away and naively returning the most likely character as the actual character.
 \begin{figure}
@@
 The characters have the following probabilities: $P(Y)=0.9, P(T)=0.1, P(e)=0.8, P(l)=0.2, P(s)=0.6, P(5)=0.4$.
-Other characters have probabilities which are neglible.}
+Other characters have probabilities which are negligible.}
 \label{fig:wordtree}
@@
 If we only want to convert the image of one word into text, we can just return the word with the highest score.
-This finalises the logic for recognising handwritten words. Next we will discuss a postprocessing technique which is useful when the word is part of a bigger context.
+This finalizes the logic for recognising handwritten words. Next we will discuss a post processing technique which is useful when the word is part of a bigger context.
 \subsection{Language model}
@@
 \end{equation}
-where $w_i$ is the word of which we want to calculate the probability and $count(w_i,...w_{i+n})$ indicates the amount of occurences of $w_i,...,w_{i+n}$.
+where $w_i$ is the word of which we want to calculate the probability and $count(w_i,...w_{i+n})$ indicates the amount of occurrences of $w_i,...,w_{i+n}$.
 That amount is the same as counting the amount of n-grams with that sequence of words.
@@
 
-We're using the following setup for this experiment.
+We use the following setup for this experiment.
 Every experiment uses the original data in addition with the augmented data using one particular preprocessing technique.
@@
 
-Table \ref{tab:preprocess} shows that augmentation techniques have great influence on the performance of the neural network. However these results might seem counterintuitive.
+Table \ref{tab:preprocess} shows that augmentation techniques have great influence on the performance of the neural network. However these results might seem counter-intuitive.
 All preprocessing techniques have a lower accuracy than the base case.
@@
 This results in a system which cannot correctly recognize handwritten text.
-Many improvements are still possible.
+Many improvements are still possible to our approach.
 
@@
 
-The character segmentation process can be improved upon by increasing the accuracy of the used neural network, and by adding more techniques to remove incorrect segmentations.
-Several such rule-based techniques are explored in \cite{segmtechniques}.
+The character segmentation process can be improved upon by increasing the accuracy of the used neural network and by adding additional techniques to remove incorrect segmentation points.
+Several rule based techniques are explored in \cite{segmtechniques}.
 

### 2017-12-18 Report/paper.tex
last review paper

@@
 In Section \ref{sec:segtext} we describe how to segment a text into words using a contouring algorithm with consequent post processing to remove unwanted noise.
-After this we present our approach to separating the characters from words in \ref{sec:segword}.
+After this we present our approach to separating the characters from words in Section \ref{sec:segword}.
 We use vertical projections combined with post processing techniques including a neural network for segmentation point evaluation.
@@
 The proposed approach focusses solely on text images, and does not take into account non-textual areas found in the input image.
-Additional techniques for page segmentation can be utilized depending on the situation.
+Additional techniques for page segmentation can be utilized depending on the situation, of which several are listed in \cite{textsegm}.
 
@@
 \cite{Otsu79}
-The Gaussian blur will smoothen the noise in the image and the edges of the text.
-Otsu's thresholding algorithm will deliver adequate result for noisy text documents.
-The choise of thresholding algorithm depends mostly on the intended purposes of the algorithm.
-Additional noise removal techniques and other preprocessing techniques may be required for a better result when processing images with much noise.
+The application of Gaussian blur to the image smoothens the noise and the edges of the text.
+Applying Otsu's thresholding algorithm on this image delivers an adequate result for noisy text documents, 
+although for very noise-heavy text images this binarization approach alone may not be sufficient.
 
@@
 The contouring algorithm can create incorrect contours because of left-over noise in the image.
-The tool we will use to remove these contours is the bounding box drawn around the contours.
+The tool we use to remove these contours is the bounding box drawn around the contours.
 This bounding box is the smallest rectangle containing the contour. It allows us to easily compare positions of contours in the text image.
@@
 These contours are neglected by taking the average height of all the found contours in the text.
-We remove all of the contours with a height smaller than $average height * (2/3)$.
+We remove all of the contours with a height smaller than $averageHeight * (2/3)$.
 
@@
 The leftover contours are used to to reconstruct the original words from the text.
-This is done by combining multiple contours which consist of parts of words belonging to the same word. 
+This is done by combining multiple contours that consist of parts of words belonging to the same word. 
 For this task we have created Algorithm \ref{alg:overlap}, which decides when to combine two word segments into a single word.
@@
         \STATE $newline = [highestWord]$
-        \STATE Initialize $noChange = true$.
         \STATE $avgY= newline$.findAverageYpos
         \REPEAT
+            \STATE $wordAdded = false$
             \FOR{$word$ {\bfseries in} $extractedWords$}
@@
             \IF{ $avgY > word.y$ {\bfseries and} $avgY < word.y + word.h$ }
-            \STATE $line$.add $word$
-            \STATE $noChange = false$
+            \STATE $line$.add $word$ {\bfseries and} $wordAdded = true$
             \ENDIF
             \ENDFOR
-        \UNTIL{$noChange = true$}
+        \UNTIL{$wordAdded == true$}
         \STATE $textlines$.add $newline$
@@
 
-\subsection{Line selection}
+\subsection{Line reconstruction}
 Because the contour finding algorithm has no particular order in which it saves the found contours, we have to reorder the found words in their original lines.
-Algorithm \ref{alg:linereconstr} shows the pseudocode of our line selection process.
+Algorithm \ref{alg:linereconstr} shows the pseudocode of our line reconstruction process.
 The algorithm reorders the words in lines and orders these lines as ordered in the original text.
@@
     \centering
+    \caption*{Original text image.}
+    \vspace{-7px}
     \includegraphics[width=.7\linewidth]{images/text}
-    \vspace{-5px}
-    \caption*{Original text image.}
     \end{subfigure}
@@
     \centering
+    \caption*{Contouring algorithm.}
+    \vspace{-7px}
     \includegraphics[width=.7\linewidth]{images/cont}
-    \vspace{-5px}
-    \caption*{Contouring algorithm.}
     \end{subfigure}
@@
     \centering
+    \caption*{Post processing and line reconstruction.}
+    \vspace{-7px}
     \includegraphics[width=.7\linewidth]{images/lines}
+    \end{subfigure}
     \vspace{-5px}
-    \caption*{Post processing and line selection.}
-    \end{subfigure}
-    \vspace{-10px}
     \caption{Text segmentation steps}
     \label{fig:text_segm}
-    \vspace{-20px}
 \end{figure}
@@
 The next step in the process is to extract the character images from the found word images.
-Possible approaches and their advantages are discussed by \cite{CharSegm} and \cite{CharSegmOld}.
-We choose the word oversegmentation approach for character extraction.
-Word oversegmentation is achieved through skeletonization and using a vertical projection profile which is discussed in consequently \ref{sec:skel} and \ref{sec:vertproj}.
-After oversegmenting the word, we remove all unnecessary segmentation points as explained in \ref{sec:postprocesscharseg}
+In this paper we use the word oversegmentation approach for character extraction, as worked out in \cite{CharSegm} and \cite{CharSegmOld}.
+Word oversegmentation is achieved through skeletonization and the use of a vertical projection profile, as discussed in Sections \ref{sec:skel} and \ref{sec:vertproj}.
+After oversegmenting the word, all unnecessary segmentation points are removed as explained in Section \ref{sec:postprocesscharseg}.
 
 \subsection{Preprocessing}
-We resize the image to a fixed height of 100 pixels. 
+We resize the word image to a fixed height of 100 pixels. 
 Next we binarize the image using a combination of Gaussian blur and Otsu's method. 
-The binarization method is identical to \ref{sec:binartext} but included to allow segmentation of words which are not part of a text.
-Binarization of the word image is essential because the skeletonization algorithm requires a binary image as input.
+The binarization method is identical to the one in Section \ref{sec:binartext}. Binarization of the word image is essential because the skeletonization algorithm requires a binary image as input.
 
@@
 \label{sec:skel}
-Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
-allows us to create a vertical projection profile. We use the skeletonization algorithm of \cite{zsthinning}.
+Next we skeletonize the word image.
+This normalizes the thickness of the strokes to a single pixel.
+This allows us interpret the results from the vertical projection histogram we make of the word image in Section \ref{sec:vertproj}.
+The skeletonization algorithm we use is proposed in \cite{zsthinning}.
 
@@
 
-Using the skeletonized word image, we create a vertical projection histogram.
+Using the skeletonized word image, we create the vertical projection histogram.
 This vertical projection histogram counts the black pixels of every column in the image.
@@
 We interpret these columns as white space in between characters or as a line connecting two characters to each other.
-This will inevitably cause oversegmentation in our word image.
-The next step filters incorrect segmentation points caused by oversegmentation.
+This will inevitably cause oversegmentation in our word image, meaning that this approach is likely to also generate incorrect segmentation points.
+In the next section (Section \ref{sec:vertproj}) we present our solution to remove these incorrect segmentation points.
 
@@
 
-In order to limit the time these experiments run we added some small constraints. We only run every experiment 200 epochs, running these experiments longer might result in different conclusions.
+In order to limit the time these experiments run we add some small constraints. We only run every experiment 200 epochs, running these experiments longer might result in different conclusions.
 Next it is important to note that a lot of stochastic behaviour is part of these neural networks because of the dropout layers.
@@
 \section{Future work}
-Our recognition system cannot be used for real-world applications.
-Every individual technique has a certain error, which increases the difficulty of the next step; e.g.
+In its current state our recognition system is not ready for real-world applications.
+In every individual step errors can be added, which increases the difficulty of the next step; e.g.
 if word segmentation contains mistakes, character recognition becomes much harder.
-This results in a system which cannot correctly recognize handwritten text.
+The result is a system that cannot correctly recognize handwritten text.
 Many improvements are still possible to our approach.
